### 2012-05-15 manuscript/textmine.Rmd
started the manuscript, just a template and table of text mining API sources

# rOpenSci: tools for textmining open source science literature

## Abstract

1.  ABSTRACT PARAGRAPH 1

2.  ABSTRACT PARAGRAPH 2

3.  ABSTRACT PARAGRAPH 3

4.  ABSTRACT PARAGRAPH 4

#### Keywords

R, software, API, PLoS, Springer, Pensoft, database, textmining, opensource

Introduction
============

Explosion of digital context online. But we need tools to facilitate textmining (i.e., rOpenSci)

Why text mine it.

How to text mine it.

Here, we introduce and give examples for textmining in R using rOpenSci packages, and related packages.

Text mining sources
===================

[Describe open source data for text mining that we link to in R, and those we don't link to in R.]

There is increasing open source scientific literature content available online. However, only a small proportion of scientific journals provide access to their full content; whereas, many publishers provide open access to their metadata only (Table 1).

Data Provider                      | Content Type            | rOpenSci? | API Documentation
---------------------------------- | ---------------------   | --------- | -----------------
Public Library of Science (PLoS)   | Full text/altmetrics    | rplos     | http://api.plos.org/
Springer                           | Full text on OA content | rspringer | http://dev.springer.com/
Pensoft                            | Full text/altmetrics    | rpensoft  | http://bit.ly/KYP0Zi
Nature Publishing Group            | Metadata only           | No        | http://developers.nature.com/
Mendeley                           | Metadata only           | rmendeley | http://dev.mendeley.com/
DataCite                           | Metadata only           | rdatacite | http://oai.datacite.org/
Biodiversity Heritage Library      | Full content            | rbhl      | http://bit.ly/KYQ1Rd
Scopus (Elsevier)                  | Metadata only ????      | No        | http://bit.ly/J9S616

How to text mine from R: Three case studies
===========================================

Case study 1
------------

Case study 2
------------

Case study 3
------------

Conclusion
==========

Acknowledgements
================

# References

Appendix
========

### 2012-06-25 manuscript/textmine.Rmd
added some table text

@@
 
+#### Table 1. Sources of scientific literature, their content type provided via web services, whether rOpenSci has an R packages for the service, and where to find the API documentation.
+
 Data Provider                      | Content Type            | rOpenSci? | API Documentation
@@
 Scopus (Elsevier)                  | Metadata only ????      | No        | http://bit.ly/J9S616
+
+#### Table 2. Tools for text mining science literature.
+
+Tool           |  Platform            | Services              | URL
+-------------- | -------------------- | --------------------- | ---------------------
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+rOpenSci       | R                    | See Table 1           | http://ropensci.org/
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
 
@@
 ========
+
+### To read:
++ http://lintool.github.com/MapReduceAlgorithms/index.html
++

### 2017-09-12 manuscript.Rmd
reworked setup to follow what we did for the gbif manuscript

--- 2012-06-25 manuscript/textmine.Rmd
@@
-# rOpenSci: tools for textmining open source science literature
+---
+layout: review, 11pt
+linenumbers: true
+title: "rOpenSci tools for textmining open source science literature"
+author:
+  - name: Scott Chamberlain
+    affiliation: cstar
+    email: scott(at)ropensci.org
+    footnote: Corresponding author
+address:
+  - code: cstar
+    address: |
+      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
+abstract: |
+      Corresponding Author:
 
-## Abstract
+      Scott Chamberlain
 
-1.  ABSTRACT PARAGRAPH 1
+      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
 
-2.  ABSTRACT PARAGRAPH 2
+      Email address: [scott@ropensci.org](mailto:scott@ropensci.org)
 
-3.  ABSTRACT PARAGRAPH 3
+      \newpage
 
-4.  ABSTRACT PARAGRAPH 4
+      Background. xxxx.
 
-#### Keywords
+      Methods. xxxx.
 
-R, software, API, PLoS, Springer, Pensoft, database, textmining, opensource
+      Results. xxxx.
 
-Introduction
-============
+      Discussion. xxxx.
+
+bibliography: components/references.bib
+csl: components/peerj.csl
+documentclass: components/elsarticle
+
+output:
+  pdf_document:
+    template: components/elsarticle.latex
+    keep_tex: true
+    fig_caption: true
+---
+
+```{r compile-settings, include=FALSE}
+library("methods")
+library("knitr")
+opts_chunk$set(
+  tidy = FALSE,
+  warning = FALSE,
+  message = FALSE,
+  cache = 1,
+  comment = '#>',
+  collapse = TRUE,
+  verbose = TRUE
+)
+
+basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
+opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
+               cache.path = paste("components/cache/", basename, "/", sep=""))
+
+# tibble options
+options(tibble.max_extra_cols = 10)
+```
+
+\newpage
+
+# Introduction
 
@@
 
-Text mining sources
-===================
+# Text mining sources
 
@@
 
-How to text mine from R: Three case studies
-===========================================
+# How to text mine from R: Three case studies
 
-Case study 1
-------------
+### Case study 1
 
-Case study 2
-------------
+```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
+```
 
-Case study 3
-------------
+### Case study 2
 
-Conclusion
-==========
+```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
+```
 
-Acknowledgements
-================
+### Case study 3
+
+```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
+```
+
+# Conclusions and future directions
+
+# Acknowledgments
+
+# Data Accessibility
+
+All scripts and data used in this paper can be found in the permanent
+data archive Zenodo under the digital object identifier (DOI). This DOI
+corresponds to a snapshot of the GitHub repository at
+<https://github.com/ropensci/textmine>. Software can be found at
+<https://github.com/ropensci/xxx>, xxxx, all under MIT licenses.
 
 # References
-
-Appendix
-========
-
-### To read:
-+ http://lintool.github.com/MapReduceAlgorithms/index.html
-+

### 2017-09-12 manuscript.tex
reworked setup to follow what we did for the gbif manuscript

\documentclass[author-year, review, 11pt]{components/elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
 \def\tightlist{}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rOpenSci tools for textmining open source science literature},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{rOpenSci tools for textmining open source science literature}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{scott(at)ropensci.org} 
   \cortext[c1]{Corresponding author}
      \address[cstar]{rOpenSci, Museum of Paleontology, University of California, Berkeley,
CA, USA}

  \begin{abstract}
  Corresponding Author:

  Scott Chamberlain

  rOpenSci, Museum of Paleontology, University of California, Berkeley,
  CA, USA

  Email address:
  \href{mailto:scott@ropensci.org}{\nolinkurl{scott@ropensci.org}}

  \newpage

  Background. xxxx.

  Methods. xxxx.

  Results. xxxx.

  Discussion. xxxx.
  \end{abstract}

 \end{frontmatter}

\newpage

\section{Introduction}\label{introduction}

Explosion of digital context online. But we need tools to facilitate
textmining (i.e., rOpenSci)

Why text mine it.

How to text mine it.

Here, we introduce and give examples for textmining in R using rOpenSci
packages, and related packages.

\section{Text mining sources}\label{text-mining-sources}

{[}Describe open source data for text mining that we link to in R, and
those we don't link to in R.{]}

There is increasing open source scientific literature content available
online. However, only a small proportion of scientific journals provide
access to their full content; whereas, many publishers provide open
access to their metadata only (Table 1).

\paragraph{Table 1. Sources of scientific literature, their content type
provided via web services, whether rOpenSci has an R packages for the
service, and where to find the API
documentation.}\label{table-1.-sources-of-scientific-literature-their-content-type-provided-via-web-services-whether-ropensci-has-an-r-packages-for-the-service-and-where-to-find-the-api-documentation.}

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.36\columnwidth}\raggedright\strut
Data Provider\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
Content Type\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
rOpenSci?\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
API Documentation\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
Public Library of Science (PLoS)\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Full text/altmetrics\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
rplos\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://api.plos.org/}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
Springer\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Full text on OA content\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
rspringer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://dev.springer.com/}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
Pensoft\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Full text/altmetrics\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
rpensoft\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://bit.ly/KYP0Zi}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
Nature Publishing Group\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Metadata only\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
No\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://developers.nature.com/}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
Mendeley\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Metadata only\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
rmendeley\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://dev.mendeley.com/}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
DataCite\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Metadata only\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
rdatacite\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://oai.datacite.org/}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
Biodiversity Heritage Library\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Full content\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
rbhl\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://bit.ly/KYQ1Rd}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
Scopus (Elsevier)\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Metadata only ????\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
No\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\url{http://bit.ly/J9S616}\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\paragraph{Table 2. Tools for text mining science
literature.}\label{table-2.-tools-for-text-mining-science-literature.}

\begin{longtable}[]{@{}llll@{}}
\toprule
Tool & Platform & Services & URL\tabularnewline
\midrule
\endhead
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
rOpenSci & R & See Table 1 & \url{http://ropensci.org/}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
\bottomrule
\end{longtable}

\section{How to text mine from R: Three case
studies}\label{how-to-text-mine-from-r-three-case-studies}

\subsubsection{Case study 1}\label{case-study-1}

\subsubsection{Case study 2}\label{case-study-2}

\subsubsection{Case study 3}\label{case-study-3}

\section{Conclusions and future
directions}\label{conclusions-and-future-directions}

\section{Acknowledgments}\label{acknowledgments}

\section{Data Accessibility}\label{data-accessibility}

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
\url{https://github.com/ropensci/textmine}. Software can be found at
\url{https://github.com/ropensci/xxx}, xxxx, all under MIT licenses.

\section{References}\label{references}

\end{document}

### 2017-09-12 use-cases/niche-modelling.Rmd
reworked setup to follow what we did for the gbif manuscript

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

In this example, we plot actual occurrence data for _Bradypus_ species against a single predictor variable, BIO1 (annual mean temperature). This is only ont step in a species distribution modelling nworkflow.

This example can be done using BISON data as well with our rbison package.

_Load libraries_

```{r}
library("sp")
library("rgbif")
library("dismo")
library("maptools")
library("dplyr")
```

_Raster files_

Make a list of files that are installed with the dismo package, then create a rasterStack from these

```{r}
files <- list.files(paste(system.file(package = "dismo"), "/ex", sep = ""),
                    "grd", full.names = TRUE)
predictors <- stack(files)
```

_Get world boundaries_

```{r}
data(wrld_simpl)
```

_Get GBIF data using the rOpenSci package rgbif_

```{r cache=FALSE}
nn <- name_lookup("bradypus*", rank = "species")
nn <- na.omit(unique(nn$data$nubKey))
df <- occ_search(taxonKey = nn, hasCoordinate = TRUE, limit = 500)
df_data <- df[ sapply(df, function(x) any(class(x$data) %in% "tbl_df")) ]
df_data <- dplyr::bind_rows(lapply(df_data, "[[", "data"))
df2 <- df_data %>% dplyr::select(decimalLongitude, decimalLatitude)
```

_Plot_

(1) Add raster data, (2) Add political boundaries, (3) Add the points (occurrences)

```{r fig.keep='last'}
plot(predictors, 1)
plot(wrld_simpl, add = TRUE)
points(df2, col = "blue")
```

### 2017-09-12 git_deploy.sh
add circle-ci deploy script, attempt deployment on circleci back to github

@@ -1,0 +1,14 @@
+#!/bin/bash
+set -e
+
+{
+  git pull
+} > /dev/null 2>&1
+
+git config --global user.email 'myrmecocystus@gmail.com' && git config --global user.name 'Scott Chamberlain'
+echo $(date) > date.txt
+git add date.txt
+git commit -m 'built and deployed from circle [ci skip]'
+{
+  git push https://sckott:$GH_TOKEN@github.com/sckott/msrgbif gh-pages
+} &> /dev/null 2>&1

### 2017-09-12 git_deploy.sh
fix repo name in deploy script

@@ -10,5 +10,5 @@ echo $(date) > date.txt
 git add date.txt
 git commit -m 'built and deployed from circle [ci skip]'
 {
-  git push https://sckott:$GH_TOKEN@github.com/sckott/msrgbif gh-pages
+  git push https://sckott:$GH_TOKEN@github.com/ropensci/textmine gh-pages
 } &> /dev/null 2>&1

### 2017-09-13 git_deploy.sh
another attempt to fix deploy

@@ -2,7 +2,7 @@
 set -e
 
 {
-  git pull
+  git fetch --unshallow
 } > /dev/null 2>&1
 
 git config --global user.email 'myrmecocystus@gmail.com' && git config --global user.name 'Scott Chamberlain'

### 2017-09-13 manuscript.Rmd
updates to manuscript, added new sections

@@
 
-Explosion of digital context online. But we need tools to facilitate textmining (i.e., rOpenSci)
+There's likely more than 100 million articles published
+(source: Crossref API), representing an enormous amount
+of knowledge.  In addition to simply reading these articles,
+they contain a vast trove of information of interest to
+researchers.
 
-Why text mine it.
+For example, many researchers are interested in statistical
+outcomes of articles: questions about P-values, about effect
+sizes, and more. With regard to effect sizes, these are
+of particular interest, as they are often combined in
+meta-analyses to draw broad conclusions about a particular
+question.
 
-How to text mine it.
+Text-mining is the broad term associated with pulling information
+out of articles. Given the importance of text-mining, good
+text-mining tools are needed to make it easier for researchers
+to do.
 
-Here, we introduce and give examples for textmining in R using rOpenSci packages, and related packages.
+Here, we do an overview of text-mining tools in the R programming
+language. We do not cover analysis tools per se, but rather those
+tools for searching for, acquiring, and "mashing up" text.
 
-# Text mining sources
+# Digital articles: technical aspects
 
-[Describe open source data for text mining that we link to in R, and those we don't link to in R.]
+Of digital articles some of which are available digitally,
+and some of which are not. Those that are digital can be split
+into two groups: easily machine readable and non-machine readable.
 
-There is increasing open source scientific literature content available online.
-However, only a small proportion of scientific journals provide access to their full content; whereas, many publishers provide open access to their metadata only (Table 1).
+The machine readable articles are those in XML, JSON, or plain text
+format. The former two, XML and JSON, are ideal of the machine
+readable types because they are structured data, whereas
+plain text has no structure - it's simply a long set of characters
+with line breaks and spaces in between.
 
-#### Table 1. Sources of scientific literature, their content type provided via web services, whether rOpenSci has an R packages for the service, and where to find the API documentation.
+Of the non-machine readable kind, there's PDFs. These can be broken out
+into two groups: text based PDFs and scanned PDFs. The former are
+converted from digital versions of various kinds (MS Word, OpenOffice,
+markdown, etc.), while the latter are PDFs created by scanning
+in print articles for which there is no digital version.
 
-Data Provider                      | Content Type            | rOpenSci? | API Documentation
----------------------------------- | ---------------------   | --------- | -----------------
-Public Library of Science (PLoS)   | Full text/altmetrics    | rplos     | http://api.plos.org/
-Springer                           | Full text on OA content | rspringer | http://dev.springer.com/
-Pensoft                            | Full text/altmetrics    | rpensoft  | http://bit.ly/KYP0Zi
-Nature Publishing Group            | Metadata only           | No        | http://developers.nature.com/
-Mendeley                           | Metadata only           | rmendeley | http://dev.mendeley.com/
-DataCite                           | Metadata only           | rdatacite | http://oai.datacite.org/
-Biodiversity Heritage Library      | Full content            | rbhl      | http://bit.ly/KYQ1Rd
-Scopus (Elsevier)                  | Metadata only ????      | No        | http://bit.ly/J9S616
+# Digital articles: the access landscape
 
-#### Table 2. Tools for text mining science literature.
+Acces to full-text is the holy grail in text-mining. Some use cases can
+get by with article metadata (authors, title, etc.), some with abstracts,
+but many use cases need full-text.
 
-Tool           |  Platform            | Services              | URL
--------------- | -------------------- | --------------------- | ---------------------
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-rOpenSci       | R                    | See Table 1           | http://ropensci.org/
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
-PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
+The landscape of access to full-text is a extremely hetergeous, with the
+majority of variation along the publisher axis. The major hurdle are
+paywalls. The majority of articles are published by the big three
+publishers - Wiley, Springer, Elsever - and the majority of their articles
+are behind paywalls.
+
+A promising sign is that there's an increasing number of open access
+publishers. xxxx.
+
+# The discovery problem (maybe remove section)
+
+xxx
+
+# Data sources
+
+There is increasing open source scientific literature content available
+online. However, only a small proportion of scientific journals provide
+access to their full content; whereas, most publishers provide open
+access to their metadata only (most often through Crossref; Table 1).
+
+Table 1. Sources of scientific literature, their content type provided via web services,
+whether rOpenSci has an R packages for the service, and where to find the API documentation.
+
+Data Provider                      | Content Type            | rOpenSci Pkg?            | API Documentation
+---------------------------------- | ---------------------   | ------------------------ | -----------------
+Crossref                           | Metadata only           | rcrossref                | [^1]
+DataCite                           | Metadata only           | rdatacite                | [^2]
+Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
+Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
+Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
+arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
+Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
+bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
+PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
+Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^10]
+
+[^1]: http://api.crossref.org
+[^2]: https://support.datacite.org/docs/api
+[^3]: http://bit.ly/KYQ1Rd
+[^4]: http://api.plos.org/solr
+[^5]: http://bit.ly/J9S616
+[^6]: https://arxiv.org/help/api/index
+[^7]: https://dev.springer.com/
+[^8]: http://www.biorxiv.org/
+[^9]: https://www.ncbi.nlm.nih.gov/books/NBK25500
+[^10]: https://azure.microsoft.com/en-us/services/cognitive-services
+
+The following is a synopsis of the major data sources and associated R tools.
+
+## Crossref
+
+Crossref is a non-profit that creates (or "mints") Digital Object Identifiers (DOIs).
+In addition, they maintain metadata associated with each DOI. The metadata ranges
+from simple (including author, title, dates, DOI, type, publisher) to including
+number of citations to the article, as well as references in the article, and even
+abstracts.
+
+Crossref does have a text-mining opt-in program for publishers. The result of this is that some
+publishers deposit URLs for full text content of their articles. The majority of these links
+are pay-walled, while some are open access. Using any of the various tools for working with
+Crossref data, you can filter your search to get only articles with full text links, and further
+to get only articles with full text links that are open access.
+
+The main interface for Crossref in R is [rcrossref][]. Parallel interfaces are available in
+Ruby ([serrano][]) and Python ([habanero][]).
+
+## Pubmed
+
+Pubmed is a corpus/website of NIH funded research ...
 
@@
 
+xxxx
+
 # Acknowledgments
+
+xxxx
 
@@
 # References
+
+[rcrossref]: https://github.com/ropensci/rcrossref
+[serrano]: https://github.com/sckott/serrano
+[habanero]: https://github.com/sckott/habanero

### 2017-09-13 manuscript.tex
updates to manuscript, added new sections

@@
 
-Explosion of digital context online. But we need tools to facilitate
-textmining (i.e., rOpenSci)
+There's likely more than 100 million articles published (source:
+Crossref API), representing an enormous amount of knowledge. In addition
+to simply reading these articles, they contain a vast trove of
+information of interest to researchers.
 
-Why text mine it.
+For example, many researchers are interested in statistical outcomes of
+articles: questions about P-values, about effect sizes, and more. With
+regard to effect sizes, these are of particular interest, as they are
+often combined in meta-analyses to draw broad conclusions about a
+particular question.
 
-How to text mine it.
+Text-mining is the broad term associated with pulling information out of
+articles. Given the importance of text-mining, good text-mining tools
+are needed to make it easier for researchers to do.
 
-Here, we introduce and give examples for textmining in R using rOpenSci
-packages, and related packages.
+Here, we do an overview of text-mining tools in the R programming
+language. We do not cover analysis tools per se, but rather those tools
+for searching for, acquiring, and ``mashing up'' text.
 
-\section{Text mining sources}\label{text-mining-sources}
+\section{Digital articles: technical
+aspects}\label{digital-articles-technical-aspects}
 
-{[}Describe open source data for text mining that we link to in R, and
-those we don't link to in R.{]}
+Of digital articles some of which are available digitally, and some of
+which are not. Those that are digital can be split into two groups:
+easily machine readable and non-machine readable.
+
+The machine readable articles are those in XML, JSON, or plain text
+format. The former two, XML and JSON, are ideal of the machine readable
+types because they are structured data, whereas plain text has no
+structure - it's simply a long set of characters with line breaks and
+spaces in between.
+
+Of the non-machine readable kind, there's PDFs. These can be broken out
+into two groups: text based PDFs and scanned PDFs. The former are
+converted from digital versions of various kinds (MS Word, OpenOffice,
+markdown, etc.), while the latter are PDFs created by scanning in print
+articles for which there is no digital version.
+
+\section{Digital articles: the access
+landscape}\label{digital-articles-the-access-landscape}
+
+Acces to full-text is the holy grail in text-mining. Some use cases can
+get by with article metadata (authors, title, etc.), some with
+abstracts, but many use cases need full-text.
+
+The landscape of access to full-text is a extremely hetergeous, with the
+majority of variation along the publisher axis. The major hurdle are
+paywalls. The majority of articles are published by the big three
+publishers - Wiley, Springer, Elsever - and the majority of their
+articles are behind paywalls.
+
+A promising sign is that there's an increasing number of open access
+publishers. xxxx.
+
+\section{The discovery problem (maybe remove
+section)}\label{the-discovery-problem-maybe-remove-section}
+
+xxx
+
+\section{Data sources}\label{data-sources}
 
@@
 online. However, only a small proportion of scientific journals provide
-access to their full content; whereas, many publishers provide open
-access to their metadata only (Table 1).
+access to their full content; whereas, most publishers provide open
+access to their metadata only (most often through Crossref; Table 1).
 
-\paragraph{Table 1. Sources of scientific literature, their content type
-provided via web services, whether rOpenSci has an R packages for the
-service, and where to find the API
-documentation.}\label{table-1.-sources-of-scientific-literature-their-content-type-provided-via-web-services-whether-ropensci-has-an-r-packages-for-the-service-and-where-to-find-the-api-documentation.}
+Table 1. Sources of scientific literature, their content type provided
+via web services, whether rOpenSci has an R packages for the service,
+and where to find the API documentation.
 
@@
 \toprule
-\begin{minipage}[b]{0.36\columnwidth}\raggedright\strut
-Data Provider\strut
-\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
-Content Type\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-rOpenSci?\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
-API Documentation\strut
-\end{minipage}\tabularnewline
+Data Provider & Content Type & rOpenSci Pkg? & API
+Documentation\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-Public Library of Science (PLoS)\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Full text/altmetrics\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-rplos\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://api.plos.org/}\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-Springer\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Full text on OA content\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-rspringer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://dev.springer.com/}\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-Pensoft\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Full text/altmetrics\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-rpensoft\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://bit.ly/KYP0Zi}\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-Nature Publishing Group\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Metadata only\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-No\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://developers.nature.com/}\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-Mendeley\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Metadata only\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-rmendeley\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://dev.mendeley.com/}\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-DataCite\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Metadata only\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-rdatacite\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://oai.datacite.org/}\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-Biodiversity Heritage Library\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Full content\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-rbhl\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://bit.ly/KYQ1Rd}\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.36\columnwidth}\raggedright\strut
-Scopus (Elsevier)\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Metadata only ????\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-No\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\url{http://bit.ly/J9S616}\strut
-\end{minipage}\tabularnewline
+Crossref & Metadata only & rcrossref & \footnote{\url{http://api.crossref.org}}\tabularnewline
+DataCite & Metadata only & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
+Biodiversity Heritage Library & Full content/Metadata & rbhl &
+\footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
+Public Library of Science (PLoS) & Full text/altmetrics & rplos &
+\footnote{\url{http://api.plos.org/solr}}\tabularnewline
+Scopus (Elsevier) & Full content/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
+arXiv & Full content/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
+Biomed Central (via Springer) & Full content/Metadata & fulltext &
+\footnote{\url{https://dev.springer.com/}}\tabularnewline
+bioRxiv & Full content/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
+PMC/Pubmed (via Entrez) & Full content/Metadata & rentrez & \footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
+Microsoft Academic Search & Metadata & fulltext/microdemic & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
 \bottomrule
@@
 
-\paragraph{Table 2. Tools for text mining science
-literature.}\label{table-2.-tools-for-text-mining-science-literature.}
+The following is a synopsis of the major data sources and associated R
+tools.
 
-\begin{longtable}[]{@{}llll@{}}
-\toprule
-Tool & Platform & Services & URL\tabularnewline
-\midrule
-\endhead
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-rOpenSci & R & See Table 1 & \url{http://ropensci.org/}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-PMC Miner & Python & PubMed & \url{http://bit.ly/L29ekY}\tabularnewline
-\bottomrule
-\end{longtable}
+\subsection{Crossref}\label{crossref}
+
+Crossref is a non-profit that creates (or ``mints'') Digital Object
+Identifiers (DOIs). In addition, they maintain metadata associated with
+each DOI. The metadata ranges from simple (including author, title,
+dates, DOI, type, publisher) to including number of citations to the
+article, as well as references in the article, and even abstracts.
+
+Crossref does have a text-mining opt-in program for publishers. The
+result of this is that some publishers deposit URLs for full text
+content of their articles. The majority of these links are pay-walled,
+while some are open access. Using any of the various tools for working
+with Crossref data, you can filter your search to get only articles with
+full text links, and further to get only articles with full text links
+that are open access.
+
+The main interface for Crossref in R is
+\href{https://github.com/ropensci/rcrossref}{rcrossref}. Parallel
+interfaces are available in Ruby
+(\href{https://github.com/sckott/serrano}{serrano}) and Python
+(\href{https://github.com/sckott/habanero}{habanero}).
+
+\subsection{Pubmed}\label{pubmed}
+
+Pubmed is a corpus/website of NIH funded research \ldots{}
 
@@
 
+xxxx
+
 \section{Acknowledgments}\label{acknowledgments}
+
+xxxx
 

### 2019-04-05 manuscript.Rmd
use cases filled out a bit, updates to data provider sections

@@
 linenumbers: true
-title: "rOpenSci tools for textmining open source science literature"
+title: "rOpenSci tools for accessing science literature for textmining"
 author:
@@
     affiliation: cstar
-    email: scott(at)ropensci.org
+    email: myrmecocystus(at)gmail.com
     footnote: Corresponding author
@@
 
-      Email address: [scott@ropensci.org](mailto:scott@ropensci.org)
+      Email address: [myrmecocystus@gmail.com](mailto:myrmecocystus@gmail.com)
 
@@
 
-There's likely more than 100 million articles published
+There's more than 100 million articles published
 (source: Crossref API), representing an enormous amount
@@
 they contain a vast trove of information of interest to
-researchers.
+researchers for machine aided questions.
 
@@
 out of articles. Given the importance of text-mining, good
-text-mining tools are needed to make it easier for researchers
-to do.
+text-mining tools are needed to make it easier for researchers.
+In particular, the R programming language is used widely throughout
+many academic fields and thus tools in R for text mining are of
+particular importance.
 
-Here, we do an overview of text-mining tools in the R programming
+Here, we present an overview of text-mining tools in the R programming
 language. We do not cover analysis tools per se, but rather those
@@
 
-Of digital articles some of which are available digitally,
-and some of which are not. Those that are digital can be split
-into two groups: easily machine readable and non-machine readable.
+Those articles that are digital can be split into two groups:
+easily machine readable and non-machine readable.
 
 The machine readable articles are those in XML, JSON, or plain text
-format. The former two, XML and JSON, are ideal of the machine
+format. The former two, XML and JSON, are ideal for the machine
 readable types because they are structured data, whereas
-plain text has no structure - it's simply a long set of characters
+plain text has no structure - it's simply a set of characters
 with line breaks and spaces in between.
@@
 
-The landscape of access to full-text is a extremely hetergeous, with the
-majority of variation along the publisher axis. The major hurdle are
+The landscape of access to full-text is extremely hetergeous, with the
+majority of variation along the publisher axis. The major hurdle is
 paywalls. The majority of articles are published by the big three
-publishers - Wiley, Springer, Elsever - and the majority of their articles
+publishers - Wiley, Springer, Elsevier - and the majority of their articles
 are behind paywalls.
 
-A promising sign is that there's an increasing number of open access
-publishers. xxxx.
+A promising sign is an increasing number of open access publishers, yet
+these represent a very small portion of the total articles (XXXXX) (ref.).
 
-# The discovery problem (maybe remove section)
+With respect to paywalled articles, access varies by institution, depending
+on what each institution decided to pay for. In addition, some users may
+not realize access varies with IP address so that access from campus vs.
+from home (if not on a VPN) will drastically differ.
+
+We can not end this section without mentioning SciHub. This is a last
+resort option for many probably, providing access to full text of articles
+that are normally paywalled. No tools in this manuscript provide access
+to SciHub.
+
+# The discovery problem
+
+xxx
+
+#
 
@@
 
-There is increasing open source scientific literature content available
+There is increasing open access scientific literature content available
 online. However, only a small proportion of scientific journals provide
@@
 access to their metadata only (most often through Crossref; Table 1).
+The following is a synopsis of the major data sources and associated R
+tools.
 
-Table 1. Sources of scientific literature, their content type provided via web services,
-whether rOpenSci has an R packages for the service, and where to find the API documentation.
+\newpage
+Table 1. Sources of scientific literature, their content type provided via
+web services, whether rOpenSci has an R packages for the service, and
+where to find the API documentation.
 
-Data Provider                      | Content Type            | rOpenSci Pkg?            | API Documentation
+Data Provider                      | Content Type            | rOpenSci Package         | Documentation
 ---------------------------------- | ---------------------   | ------------------------ | -----------------
-Crossref                           | Metadata only           | rcrossref                | [^1]
+Crossref                           | Metadata only           | rcrossref/crminer        | [^1]
 DataCite                           | Metadata only           | rdatacite                | [^2]
@@
 PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
-Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^10]
+Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
+Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^11]
+Directory of Open Access Journals  | Metadata                | jaod                     | [^12]
+JSTOR Data for Research            | Full content            | jstor                    | [^13]
+ORCID                              | Metadata                | rorcid                   | [^14]
+Wikimedia's Citoid                 | Citations               | rcitoid                  | [^15]
+Open Citation Corpus               | Citations               | citecorp                 | [^16]
 
-[^1]: http://api.crossref.org
+[^1]: https://api.crossref.org
 [^2]: https://support.datacite.org/docs/api
@@
 [^10]: https://azure.microsoft.com/en-us/services/cognitive-services
+[^11]: https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97
+[^12]: https://doaj.org/api/v1/docs
+[^13]: https://www.jstor.org/dfr/
+[^14]: https://pub.orcid.org/
+[^15]: https://en.wikipedia.org/api/rest_v1/#/Citation/getCitation
+[^16]: http://opencitations.net/
 
-The following is a synopsis of the major data sources and associated R tools.
-
-## Crossref
+## Crossref/Datacite
 
@@
 number of citations to the article, as well as references in the article, and even
-abstracts.
+abstracts. At the time of writing they hold 100 million DOIs.
 
-Crossref does have a text-mining opt-in program for publishers. The result of this is that some
-publishers deposit URLs for full text content of their articles. The majority of these links
+One can search by DOI or search citation data to get citations. In addition, Crossref has a
+text-mining opt-in program for publishers. The result of this is that some
+publishers provide URLs for full text content of their articles. The majority of these links
 are pay-walled, while some are open access. Using any of the various tools for working with
@@
 
-The main interface for Crossref in R is [rcrossref][]. Parallel interfaces are available in
-Ruby ([serrano][]) and Python ([habanero][]).
+The main interfaces for Crossref in R are [rcrossref][] and [crminer][]. Similar interfaces
+are available in Ruby ([serrano][]) and Python ([habanero][]).
 
-## Pubmed
+Datacite is similar to Crossref, but focuses on datasets instead of articles. The main
+interface for Datacite in R is [rdatacite][].
 
-Pubmed is a corpus/website of NIH funded research ...
+## Biodiversity Heritage Library
+
+The Biodiversity Heritage Library (BHL) houses scans of biodiversity books, and provides
+web interfaces and APIs to query and fetch those data. They also provide text of the scanned
+pages. The main R interace to BHL is through [rbhl][].
+
+## Public Library of Science
+
+The Public Library of Science (PLOS) is one of the largest open access only publishers. They
+as of this writing have published 2.1 million articles. One of the strongs advantages of PLOS
+is that they provide an API to their Solr instance, which is a very flexible way to
+search their articles. The main R interace to PLOS is through [rplos][].
+
+## Elsevier/Scopus
+
+Elsevier is one of the largest publishers. Most of their articles are not open access. However,
+they have a numbrer of advantages if you have access to their articles: they are one of the few
+publishers to provide machine readable XML (many publishers do have XML versions of articles,
+but do not provide it); they are one of the few (two) publishers part of Crossref's text
+and data mining program. The packages [fulltext][] and [crminer][] can be used to access
+Elsevier articles through Crossref's TDM program. There's an interface to Scopus article
+search within [fulltext][].
+
+## arXiv/bioRxiv
+
+arXiv and bioRxiv are preprint publishers, the former in existence for many years, and the
+latter new on the scene. You can access articles from these publishers through [fulltext][].
+arXiv does provide a web API that we hook into; bioRxiv does not, but we can get you
+articles nonetheless.
+
+## Pubmed/PMC/Europe PMC
+
+Pubmed/PMC is a corpus/website of NIH funded research in the United States; while Europe PMC
+is an equivalent for the European Union. You can access articles from Pubmed/PMC through
+[fulltext][], and for Europe PMC through [europepmc][].
+
+## Microsoft Academic Research
+
+Microsoft Academic Research (MAR) is a search engine for research articles. You can use their
+GUI web interface to search, and they provide APIs for programmatic access. The R interface
+for MAR is [microdemic][]; and [fulltext][] hooks into `microdemic` as well for article search
+and abstract retrieval.
+
+## Directory of Open Access Journals
+
+xxxxx
+
+## JSTOR
+
+xxxxx
+
+## ORCID
+
+xxxxx
+
+## Citoid/Open Citation Corpus
+
+xxx
 
@@
 
-### Case study 1
+### Case study 1: Citation mining
 
-```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
+```{r citation-mining, child="use-cases/citation-mining.Rmd", eval=TRUE}
 ```
 
-### Case study 2
+### Case study 2: Abstract mining
 
-```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
+```{r abstract-mining, child="use-cases/abstract-mining.Rmd", eval=TRUE}
 ```
 
-### Case study 3
+### Case study 3: Full text mining
 
-```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
+```{r full-text-mining, child="use-cases/full-text-mining.Rmd", eval=TRUE}
 ```
 
-# Conclusions and future directions
+# Future directions
 
@@
 [rcrossref]: https://github.com/ropensci/rcrossref
+[rorcid]: https://github.com/ropensci/rorcid
+[fulltext]: https://github.com/ropensci/fulltext
+[microdemic]: https://github.com/ropensci/microdemic
+[arXiv]: https://github.com/ropensci/arXiv
+[rdatacite]: https://github.com/ropensci/rdatacite
+[rbhl]: https://github.com/ropensci/rbhl
+[rentrez]: https://github.com/ropensci/rentrez
+[rplos]: https://github.com/ropensci/rplos
+[crminer]: https://github.com/ropensci/crminer
+[europepmc]: https://github.com/ropensci/europepmc
+[jaod]: https://github.com/ropensci/jaod
+[jstor]: https://github.com/ropensci/jstor
+[rcitoid]: https://github.com/ropenscilabs/rcitoid
+[citecorp]: https://github.com/ropenscilabs/citecorp
 [serrano]: https://github.com/sckott/serrano
 [habanero]: https://github.com/sckott/habanero
+[quanteda]: https://quanteda.io/
+[pubchunks]: https://github.com/ropensci/pubchunks

### 2019-04-05 manuscript.tex
use cases filled out a bit, updates to data provider sections

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable}
@@
             pdfauthor={},
-            pdftitle={rOpenSci tools for textmining open source science literature},
+            pdftitle={rOpenSci tools for accessing science literature for textmining},
             colorlinks=true,
@@
 
-  \title{rOpenSci tools for textmining open source science literature}
+  \title{rOpenSci tools for accessing science literature for textmining}
     \author[cstar]{Scott Chamberlain\corref{c1}}
-   \ead{scott(at)ropensci.org} 
+   \ead{myrmecocystus(at)gmail.com} 
    \cortext[c1]{Corresponding author}
@@
   Email address:
-  \href{mailto:scott@ropensci.org}{\nolinkurl{scott@ropensci.org}}
+  \href{mailto:myrmecocystus@gmail.com}{\nolinkurl{myrmecocystus@gmail.com}}
 
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-There's likely more than 100 million articles published (source:
-Crossref API), representing an enormous amount of knowledge. In addition
-to simply reading these articles, they contain a vast trove of
-information of interest to researchers.
+There's more than 100 million articles published (source: Crossref API),
+representing an enormous amount of knowledge. In addition to simply
+reading these articles, they contain a vast trove of information of
+interest to researchers for machine aided questions.
 
@@
 articles. Given the importance of text-mining, good text-mining tools
-are needed to make it easier for researchers to do.
+are needed to make it easier for researchers. In particular, the R
+programming language is used widely throughout many academic fields and
+thus tools in R for text mining are of particular importance.
 
-Here, we do an overview of text-mining tools in the R programming
+Here, we present an overview of text-mining tools in the R programming
 language. We do not cover analysis tools per se, but rather those tools
@@
 
+\hypertarget{digital-articles-technical-aspects}{%
 \section{Digital articles: technical
-aspects}\label{digital-articles-technical-aspects}
+aspects}\label{digital-articles-technical-aspects}}
 
-Of digital articles some of which are available digitally, and some of
-which are not. Those that are digital can be split into two groups:
-easily machine readable and non-machine readable.
+Those articles that are digital can be split into two groups: easily
+machine readable and non-machine readable.
 
 The machine readable articles are those in XML, JSON, or plain text
-format. The former two, XML and JSON, are ideal of the machine readable
+format. The former two, XML and JSON, are ideal for the machine readable
 types because they are structured data, whereas plain text has no
-structure - it's simply a long set of characters with line breaks and
-spaces in between.
+structure - it's simply a set of characters with line breaks and spaces
+in between.
 
@@
 
+\hypertarget{digital-articles-the-access-landscape}{%
 \section{Digital articles: the access
-landscape}\label{digital-articles-the-access-landscape}
+landscape}\label{digital-articles-the-access-landscape}}
 
@@
 
-The landscape of access to full-text is a extremely hetergeous, with the
-majority of variation along the publisher axis. The major hurdle are
+The landscape of access to full-text is extremely hetergeous, with the
+majority of variation along the publisher axis. The major hurdle is
 paywalls. The majority of articles are published by the big three
-publishers - Wiley, Springer, Elsever - and the majority of their
+publishers - Wiley, Springer, Elsevier - and the majority of their
 articles are behind paywalls.
 
-A promising sign is that there's an increasing number of open access
-publishers. xxxx.
+A promising sign is an increasing number of open access publishers, yet
+these represent a very small portion of the total articles (XXXXX)
+(ref.).
 
-\section{The discovery problem (maybe remove
-section)}\label{the-discovery-problem-maybe-remove-section}
+With respect to paywalled articles, access varies by institution,
+depending on what each institution decided to pay for. In addition, some
+users may not realize access varies with IP address so that access from
+campus vs.~from home (if not on a VPN) will drastically differ.
+
+We can not end this section without mentioning SciHub. This is a last
+resort option for many probably, providing access to full text of
+articles that are normally paywalled. No tools in this manuscript
+provide access to SciHub.
+
+\hypertarget{the-discovery-problem}{%
+\section{The discovery problem}\label{the-discovery-problem}}
 
@@
 
-\section{Data sources}\label{data-sources}
+\hypertarget{section}{%
+\section{}\label{section}}
 
-There is increasing open source scientific literature content available
+xxx
+
+\hypertarget{data-sources}{%
+\section{Data sources}\label{data-sources}}
+
+There is increasing open access scientific literature content available
 online. However, only a small proportion of scientific journals provide
@@
 access to their metadata only (most often through Crossref; Table 1).
+The following is a synopsis of the major data sources and associated R
+tools.
+
+\newpage
 
@@
 \toprule
-Data Provider & Content Type & rOpenSci Pkg? & API
+Data Provider & Content Type & rOpenSci Package &
 Documentation\tabularnewline
@@
 \endhead
-Crossref & Metadata only & rcrossref & \footnote{\url{http://api.crossref.org}}\tabularnewline
+Crossref & Metadata only & rcrossref/crminer & \footnote{\url{https://api.crossref.org}}\tabularnewline
 DataCite & Metadata only & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
@@
 PMC/Pubmed (via Entrez) & Full content/Metadata & rentrez & \footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
-Microsoft Academic Search & Metadata & fulltext/microdemic & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
+Europe PMC & Full content/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
+Microsoft Academic Search & Metadata & fulltext/microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
+Directory of Open Access Journals & Metadata & jaod & \footnote{\url{https://doaj.org/api/v1/docs}}\tabularnewline
+JSTOR Data for Research & Full content & jstor & \footnote{\url{https://www.jstor.org/dfr/}}\tabularnewline
+ORCID & Metadata & rorcid & \footnote{\url{https://pub.orcid.org/}}\tabularnewline
+Wikimedia's Citoid & Citations & rcitoid & \footnote{\url{https://en.wikipedia.org/api/rest_v1/\#/Citation/getCitation}}\tabularnewline
+Open Citation Corpus & Citations & citecorp & \footnote{\url{http://opencitations.net/}}\tabularnewline
 \bottomrule
@@
 
-The following is a synopsis of the major data sources and associated R
-tools.
-
-\subsection{Crossref}\label{crossref}
+\hypertarget{crossrefdatacite}{%
+\subsection{Crossref/Datacite}\label{crossrefdatacite}}
 
@@
 dates, DOI, type, publisher) to including number of citations to the
-article, as well as references in the article, and even abstracts.
+article, as well as references in the article, and even abstracts. At
+the time of writing they hold 100 million DOIs.
 
-Crossref does have a text-mining opt-in program for publishers. The
-result of this is that some publishers deposit URLs for full text
+One can search by DOI or search citation data to get citations. In
+addition, Crossref has a text-mining opt-in program for publishers. The
+result of this is that some publishers provide URLs for full text
 content of their articles. The majority of these links are pay-walled,
@@
 
-The main interface for Crossref in R is
-\href{https://github.com/ropensci/rcrossref}{rcrossref}. Parallel
-interfaces are available in Ruby
+The main interfaces for Crossref in R are
+\href{https://github.com/ropensci/rcrossref}{rcrossref} and
+\href{https://github.com/ropensci/crminer}{crminer}. Similar interfaces
+are available in Ruby
 (\href{https://github.com/sckott/serrano}{serrano}) and Python
@@
 
-\subsection{Pubmed}\label{pubmed}
+Datacite is similar to Crossref, but focuses on datasets instead of
+articles. The main interface for Datacite in R is
+\href{https://github.com/ropensci/rdatacite}{rdatacite}.
 
-Pubmed is a corpus/website of NIH funded research \ldots{}
+\hypertarget{biodiversity-heritage-library}{%
+\subsection{Biodiversity Heritage
+Library}\label{biodiversity-heritage-library}}
 
+The Biodiversity Heritage Library (BHL) houses scans of biodiversity
+books, and provides web interfaces and APIs to query and fetch those
+data. They also provide text of the scanned pages. The main R interace
+to BHL is through \href{https://github.com/ropensci/rbhl}{rbhl}.
+
+\hypertarget{public-library-of-science}{%
+\subsection{Public Library of Science}\label{public-library-of-science}}
+
+The Public Library of Science (PLOS) is one of the largest open access
+only publishers. They as of this writing have published 2.1 million
+articles. One of the strongs advantages of PLOS is that they provide an
+API to their Solr instance, which is a very flexible way to search their
+articles. The main R interace to PLOS is through
+\href{https://github.com/ropensci/rplos}{rplos}.
+
+\hypertarget{elsevierscopus}{%
+\subsection{Elsevier/Scopus}\label{elsevierscopus}}
+
+Elsevier is one of the largest publishers. Most of their articles are
+not open access. However, they have a numbrer of advantages if you have
+access to their articles: they are one of the few publishers to provide
+machine readable XML (many publishers do have XML versions of articles,
+but do not provide it); they are one of the few (two) publishers part of
+Crossref's text and data mining program. The packages
+\href{https://github.com/ropensci/fulltext}{fulltext} and
+\href{https://github.com/ropensci/crminer}{crminer} can be used to
+access Elsevier articles through Crossref's TDM program. There's an
+interface to Scopus article search within
+\href{https://github.com/ropensci/fulltext}{fulltext}.
+
+\hypertarget{arxivbiorxiv}{%
+\subsection{arXiv/bioRxiv}\label{arxivbiorxiv}}
+
+arXiv and bioRxiv are preprint publishers, the former in existence for
+many years, and the latter new on the scene. You can access articles
+from these publishers through
+\href{https://github.com/ropensci/fulltext}{fulltext}. arXiv does
+provide a web API that we hook into; bioRxiv does not, but we can get
+you articles nonetheless.
+
+\hypertarget{pubmedpmceurope-pmc}{%
+\subsection{Pubmed/PMC/Europe PMC}\label{pubmedpmceurope-pmc}}
+
+Pubmed/PMC is a corpus/website of NIH funded research in the United
+States; while Europe PMC is an equivalent for the European Union. You
+can access articles from Pubmed/PMC through
+\href{https://github.com/ropensci/fulltext}{fulltext}, and for Europe
+PMC through \href{https://github.com/ropensci/europepmc}{europepmc}.
+
+\hypertarget{microsoft-academic-research}{%
+\subsection{Microsoft Academic
+Research}\label{microsoft-academic-research}}
+
+Microsoft Academic Research (MAR) is a search engine for research
+articles. You can use their GUI web interface to search, and they
+provide APIs for programmatic access. The R interface for MAR is
+\href{https://github.com/ropensci/microdemic}{microdemic}; and
+\href{https://github.com/ropensci/fulltext}{fulltext} hooks into
+\texttt{microdemic} as well for article search and abstract retrieval.
+
+\hypertarget{directory-of-open-access-journals}{%
+\subsection{Directory of Open Access
+Journals}\label{directory-of-open-access-journals}}
+
+xxxxx
+
+\hypertarget{jstor}{%
+\subsection{JSTOR}\label{jstor}}
+
+xxxxx
+
+\hypertarget{orcid}{%
+\subsection{ORCID}\label{orcid}}
+
+xxxxx
+
+\hypertarget{citoidopen-citation-corpus}{%
+\subsection{Citoid/Open Citation
+Corpus}\label{citoidopen-citation-corpus}}
+
+xxx
+
+\hypertarget{how-to-text-mine-from-r-three-case-studies}{%
 \section{How to text mine from R: Three case
-studies}\label{how-to-text-mine-from-r-three-case-studies}
+studies}\label{how-to-text-mine-from-r-three-case-studies}}
 
-\subsubsection{Case study 1}\label{case-study-1}
+\hypertarget{case-study-1-citation-mining}{%
+\subsubsection{Case study 1: Citation
+mining}\label{case-study-1-citation-mining}}
 
-\subsubsection{Case study 2}\label{case-study-2}
+In this example, xxxx
 
-\subsubsection{Case study 3}\label{case-study-3}
+\emph{Load libraries}
 
-\section{Conclusions and future
-directions}\label{conclusions-and-future-directions}
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"rcrossref"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"rplos"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"rorcid"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"rcitoid"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"citecorp"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{rcrossref}
+
+Using \texttt{rcrossref} for Crossref data:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\KeywordTok{cr_works}\NormalTok{(}\DataTypeTok{query=}\StringTok{"NSF"}\NormalTok{)}
+\KeywordTok{head}\NormalTok{(x}\OperatorTok{$}\NormalTok{data)}
+\CommentTok{#> # A tibble: 6 x 32}
+\CommentTok{#>   alternative.id container.title created deposited published.print doi  }
+\CommentTok{#>   <chr>          <chr>           <chr>   <chr>     <chr>           <chr>}
+\CommentTok{#> 1 S106352031630~ Applied and Co~ 2016-0~ 2019-02-~ 2018-03         10.1~}
+\CommentTok{#> 2 <NA>           Biogeosciences~ 2017-0~ 2017-07-~ <NA>            10.5~}
+\CommentTok{#> 3 <NA>           Global Biogeoc~ 2018-0~ 2019-01-~ 2018-10         10.1~}
+\CommentTok{#> 4 <NA>           IEEE Communica~ 2016-1~ 2017-12-~ 2017            10.1~}
+\CommentTok{#> 5 S002178241400~ Journal de Mat~ 2014-0~ 2018-10-~ 2014-10         10.1~}
+\CommentTok{#> 6 123            Light: Science~ 2019-0~ 2019-01-~ 2019-12         10.1~}
+\CommentTok{#> # ... with 26 more variables: indexed <chr>, issn <chr>, issue <chr>,}
+\CommentTok{#> #   issued <chr>, member <chr>, page <chr>, prefix <chr>, publisher <chr>,}
+\CommentTok{#> #   reference.count <chr>, score <chr>, ...}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{case-study-2-abstract-mining}{%
+\subsubsection{Case study 2: Abstract
+mining}\label{case-study-2-abstract-mining}}
+
+Sometimes you just need abstracts for your research question. The
+benefit of only needing abstracts, and not need full text, is that
+there's many more articles that will have abstracts available than have
+their full text available.
+
+As an example, let's say you xxxx
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"fulltext"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{xxxxx}
+
+Using \texttt{fulltext}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"crossref"}\NormalTok{,}
+  \DataTypeTok{crossrefopts =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{filter =} \KeywordTok{c}\NormalTok{(}\DataTypeTok{has_abstract =} \OtherTok{TRUE}\NormalTok{)))}
+\NormalTok{ids <-}\StringTok{ }\NormalTok{res}\OperatorTok{$}\NormalTok{crossref}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{doi}
+\NormalTok{out <-}\StringTok{ }\KeywordTok{ft_abstract}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ ids, }\DataTypeTok{from =} \StringTok{"crossref"}\NormalTok{)}
+\NormalTok{abstracts <-}\StringTok{ }\KeywordTok{vapply}\NormalTok{(out}\OperatorTok{$}\NormalTok{crossref, }\StringTok{"[["}\NormalTok{, }\StringTok{""}\NormalTok{, }\StringTok{"abstract"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Using \href{https://quanteda.io/}{quanteda}, read the abstracts into a
+corpus
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"quanteda"}\NormalTok{)}
+\NormalTok{corp <-}\StringTok{ }\KeywordTok{corpus}\NormalTok{(abstracts)}
+\KeywordTok{docvars}\NormalTok{(corp) <-}\StringTok{ }\NormalTok{ids}
+\end{Highlighting}
+\end{Shaded}
+
+Get a summary of the abstracts
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{summary}\NormalTok{(corp)}
+\CommentTok{#> Corpus consisting of 10 documents:}
+\CommentTok{#> }
+\CommentTok{#>    Text Types Tokens Sentences                    V1}
+\CommentTok{#>   text1   143    262        10   10.2458/v22i1.21112}
+\CommentTok{#>   text2   117    244         6   10.2458/v17i1.21696}
+\CommentTok{#>   text3    75    118         4   10.2458/v25i1.23119}
+\CommentTok{#>   text4     5      8         1    10.2458/v1i1.21154}
+\CommentTok{#>   text5   105    171         7   10.1155/2011/868426}
+\CommentTok{#>   text6   112    181         6   10.1155/2012/273413}
+\CommentTok{#>   text7   117    240         8 10.5194/we-13-91-2013}
+\CommentTok{#>   text8   140    245         9 10.5194/we-13-95-2013}
+\CommentTok{#>   text9   107    202         7   10.1155/2014/198707}
+\CommentTok{#>  text10   118    224         6   10.5402/2011/897578}
+\CommentTok{#> }
+\CommentTok{#> Source: /Users/sckott/github/ropensci/textmine/use-cases/* on x86_64 by sckott}
+\CommentTok{#> Created: Fri Apr  5 11:36:04 2019}
+\CommentTok{#> Notes:}
+\end{Highlighting}
+\end{Shaded}
+
+Use the \texttt{kwic()} function to see a word in context across the
+abstracts
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{kwic}\NormalTok{(corp, }\DataTypeTok{pattern =} \StringTok{"ecology"}\NormalTok{)}
+\CommentTok{#>                                                                         }
+\CommentTok{#>   [text1, 33] knowledge production within critical political | ecology |}
+\CommentTok{#>   [text1, 50]              in scientific articles on dryland | ecology |}
+\CommentTok{#>  [text1, 204]                 to equilibrium models in range | ecology |}
+\CommentTok{#>  [text1, 246]    communal areas.Keywords: Critical political | ecology |}
+\CommentTok{#>  [text1, 255]                 , scientific models, rangeland | ecology |}
+\CommentTok{#>    [text2, 5]                            < jats:p> Political | ecology |}
+\CommentTok{#>   [text2, 23]        manifestations of political economy and | ecology |}
+\CommentTok{#>   [text2, 45]                      I try to extend political | ecology |}
+\CommentTok{#>  [text2, 149]                   , in dialogue with political | ecology |}
+\CommentTok{#>  [text2, 177]            people and resources that political | ecology |}
+\CommentTok{#>  [text2, 229]      indigeneity scholars.Key words: political | ecology |}
+\CommentTok{#>   [text3, 71]                   an analysis from a political | ecology |}
+\CommentTok{#>  [text3, 114]                system, supermarkets, political | ecology |}
+\CommentTok{#>  [text6, 134]                was observed when allopatry and | ecology |}
+\CommentTok{#>  [text7, 167]             ecosystem should be considered for | ecology |}
+\CommentTok{#>  [text7, 185]                       the" four-color issue of | ecology |}
+\CommentTok{#>  [text7, 201]             step toward advancing knowledge in | ecology |}
+\CommentTok{#>  [text9, 195]         or for theoretical studies integrating | ecology |}
+\CommentTok{#>                                              }
+\CommentTok{#>  . This article is a                         }
+\CommentTok{#>  , and investigates the functions            }
+\CommentTok{#>  , and the fence-line photographs            }
+\CommentTok{#>  , fence-line photography, scientific        }
+\CommentTok{#>  , Southern Africa</                         }
+\CommentTok{#>  has expanded in multiple new                }
+\CommentTok{#>  in the" problem"                            }
+\CommentTok{#>  to engage with ethnic studies               }
+\CommentTok{#>  approaches to better understand the         }
+\CommentTok{#>  focuses on cannot be adequately             }
+\CommentTok{#>  , coloniality, Maidu,                       }
+\CommentTok{#>  standpoint allows a different interpretation}
+\CommentTok{#>  </ jats:p>                                  }
+\CommentTok{#>  act together, leading to                    }
+\CommentTok{#>  "? Here, I                                  }
+\CommentTok{#>  ", and propose that                         }
+\CommentTok{#>  and conservation biology. In                }
+\CommentTok{#>  and biogeography.</}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{case-study-3-full-text-mining}{%
+\subsubsection{Case study 3: Full text
+mining}\label{case-study-3-full-text-mining}}
+
+In this example, xxxx
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"fulltext"}\NormalTok{)}
+\CommentTok{# library("crminer")}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Search for articles}
+
+Search for the term \emph{ecology} in PLOS journals.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{(res1 <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \StringTok{'plos'}\NormalTok{))}
+\CommentTok{#> Query:}
+\CommentTok{#>   [ecology] }
+\CommentTok{#> Found:}
+\CommentTok{#>   [PLoS: 47337; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0] }
+\CommentTok{#> Returned:}
+\CommentTok{#>   [PLoS: 10; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0]}
+\end{Highlighting}
+\end{Shaded}
+
+Each publisher/search-engine has a slot with metadata and data
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res1}\OperatorTok{$}\NormalTok{plos}
+\CommentTok{#> Query: [ecology] }
+\CommentTok{#> Records found, returned: [47337, 10] }
+\CommentTok{#> License: [CC-BY] }
+\CommentTok{#>                              id}
+\CommentTok{#> 1  10.1371/journal.pone.0001248}
+\CommentTok{#> 2  10.1371/journal.pone.0059813}
+\CommentTok{#> 3  10.1371/journal.pone.0155019}
+\CommentTok{#> 4  10.1371/journal.pone.0080763}
+\CommentTok{#> 5  10.1371/journal.pone.0208370}
+\CommentTok{#> 6  10.1371/journal.pone.0150648}
+\CommentTok{#> 7  10.1371/journal.pcbi.1003594}
+\CommentTok{#> 8  10.1371/journal.pone.0102437}
+\CommentTok{#> 9  10.1371/journal.pone.0175014}
+\CommentTok{#> 10 10.1371/journal.pone.0166559}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get full text}
+
+Using the results from \texttt{ft\_search()} we can grab full text of
+some articles
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{(out <-}\StringTok{ }\KeywordTok{ft_get}\NormalTok{(res1))}
+\CommentTok{#> <fulltext text>}
+\CommentTok{#> [Docs] 10 }
+\CommentTok{#> [Source] ext - /Users/sckott/Library/Caches/R/fulltext }
+\CommentTok{#> [IDs] 10.1371/journal.pone.0001248 10.1371/journal.pone.0059813}
+\CommentTok{#>      10.1371/journal.pone.0155019 10.1371/journal.pone.0080763}
+\CommentTok{#>      10.1371/journal.pone.0208370 10.1371/journal.pone.0150648}
+\CommentTok{#>      10.1371/journal.pcbi.1003594 10.1371/journal.pone.0102437}
+\CommentTok{#>      10.1371/journal.pone.0175014 10.1371/journal.pone.0166559 ...}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Extract text from pdfs}
+
+Ideally for text mining you have access to XML or other text based
+formats. However, sometimes you only have access to PDFs. In this case
+you want to extract text from PDFs. \texttt{fulltext} can help with
+that.
+
+You can extract from any pdf from a file path, like:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{path <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"examples"}\NormalTok{, }\StringTok{"example1.pdf"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"fulltext"}\NormalTok{)}
+\KeywordTok{ft_extract}\NormalTok{(path)}
+\CommentTok{#> <document>/Library/Frameworks/R.framework/Versions/3.5/Resources/library/fulltext/examples/example1.pdf}
+\CommentTok{#>   Title: Suffering and mental health among older people living in nursing homes---a mixed-methods study}
+\CommentTok{#>   Producer: pdfTeX-1.40.10}
+\CommentTok{#>   Creation date: 2015-07-17}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Extract text chunks}
+
+Requires the \href{https://github.com/ropensci/pubchunks}{pubchunks}
+library. Here, we'll search for some PLOS articles, then get their full
+text, then extract various parts of each article with
+\texttt{pub\_chunks()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"pubchunks"}\NormalTok{)}
+\NormalTok{res <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"plos"}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{3}\NormalTok{)}
+\NormalTok{x <-}\StringTok{ }\KeywordTok{ft_get}\NormalTok{(res)}
+\NormalTok{x }\OperatorTok{%>%}\StringTok{ }\KeywordTok{ft_collect}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pub_chunks}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"doi"}\NormalTok{, }\StringTok{"history"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pub_tabularize}\NormalTok{()}
+\CommentTok{#> $plos}
+\CommentTok{#> $plos$`10.1371/journal.pone.0001248`}
+\CommentTok{#>                            doi history.received history.accepted}
+\CommentTok{#> 1 10.1371/journal.pone.0001248       2007-07-02       2007-11-06}
+\CommentTok{#>   .publisher}
+\CommentTok{#> 1       plos}
+\CommentTok{#> }
+\CommentTok{#> $plos$`10.1371/journal.pone.0059813`}
+\CommentTok{#>                            doi history.received history.accepted}
+\CommentTok{#> 1 10.1371/journal.pone.0059813       2012-09-16       2013-02-19}
+\CommentTok{#>   .publisher}
+\CommentTok{#> 1       plos}
+\CommentTok{#> }
+\CommentTok{#> $plos$`10.1371/journal.pone.0155019`}
+\CommentTok{#>                            doi history.received history.accepted}
+\CommentTok{#> 1 10.1371/journal.pone.0155019       2015-09-22       2016-04-22}
+\CommentTok{#>   .publisher}
+\CommentTok{#> 1       plos}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{future-directions}{%
+\section{Future directions}\label{future-directions}}
 
@@
 
-\section{Acknowledgments}\label{acknowledgments}
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
 
@@
 
-\section{Data Accessibility}\label{data-accessibility}
+\hypertarget{data-accessibility}{%
+\section{Data Accessibility}\label{data-accessibility}}
 
@@
 
-\section{References}\label{references}
+\hypertarget{references}{%
+\section{References}\label{references}}
 

### 2019-04-05 use-cases/abstract-mining.Rmd
use cases filled out a bit, updates to data provider sections

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

Sometimes you just need abstracts for your research question. The benefit of 
only needing abstracts, and not need full text, is that there's many more
articles that will have abstracts available than have their full text 
available.

As an example, let's say you xxxx

```{r}
library("fulltext")
```

_xxxxx_

Using `fulltext`:

```{r}
res <- ft_search("ecology", from = "crossref",
  crossrefopts = list(filter = c(has_abstract = TRUE)))
ids <- res$crossref$data$doi
out <- ft_abstract(x = ids, from = "crossref")
abstracts <- vapply(out$crossref, "[[", "", "abstract")
```

Using [quanteda][], read the abstracts into a corpus

```{r}
library("quanteda")
corp <- corpus(abstracts)
docvars(corp) <- ids
```

Get a summary of the abstracts

```{r}
summary(corp)
```

Use the `kwic()` function to see a word in context across
the abstracts

```{r}
kwic(corp, pattern = "ecology")
```

### 2019-04-05 use-cases/full-text-mining.Rmd
use cases filled out a bit, updates to data provider sections

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

In this example, xxxx

```{r}
library("fulltext")
# library("crminer")
```

_Search for articles_

Search for the term _ecology_ in PLOS journals.

```{r}
(res1 <- ft_search(query = 'ecology', from = 'plos'))
```

Each publisher/search-engine has a slot with metadata and data

```{r}
res1$plos
```

_Get full text_

Using the results from `ft_search()` we can grab full text of some articles

```{r}
(out <- ft_get(res1))
```

_Extract text from pdfs_

Ideally for text mining you have access to XML or other text based formats. However, 
sometimes you only have access to PDFs. In this case you want to extract text 
from PDFs. `fulltext` can help with that.

You can extract from any pdf from a file path, like:

```{r}
path <- system.file("examples", "example1.pdf", package = "fulltext")
ft_extract(path)
```

_Extract text chunks_

Requires the [pubchunks][] library. Here, we'll search for some PLOS articles, then
get their full text, then extract various parts of each article with `pub_chunks()`.

```{r}
library("pubchunks")
res <- ft_search(query = "ecology", from = "plos", limit = 3)
x <- ft_get(res)
x %>% ft_collect() %>% pub_chunks(c("doi", "history")) %>% pub_tabularize()
```

### 2019-04-08 manuscript.Rmd
paper updates

@@
 linenumbers: true
-title: "rOpenSci tools for accessing science literature for textmining"
+title: "rOpenSci tools for accessing research literature for text mining"
 author:
@@
 
+# fulltext: a swiss army knife for text mining in R
+
+[fulltext][] is a general purpose R package for the data part of text-mining:
+search for articles, get links to articles, get article abstracts, and fetch
+full text of articles. The `fulltext` package is always adding additional
+data sources as time allows (See Table 1). Starting from searching for 
+articles, the outputs of search can be fed into a function to get links 
+to those articles, or to get abstracts for those articles, or to fetch 
+their full text.
+
+The following is a breakdown of the major distinct functional parts of 
+`fulltext`.
+
+## Search
+
+`ft_search()` provides search access to nine different data sources (PLOS, BMC, Crossref,
+Entrez, arXiv, bioRxiv, Europe PMC, Scopus, Microsoft Academic), creating a
+mostly unified interface to all data sources. The parts of each data source
+that are common are mostly factored into the parameters of the `ft_search()` function,
+and we also allow the user to pass on data source specific options as needed.
+
+xxxxx
+
+## Links
+
+`ft_links()` provides two pathways to get links (URLs) for articles, with a choice
+of four different data sources (PLOS, BMC, Crossref, Entrez). First, you can
+use `ft_search()`, then pass the output of that function to `ft_links()`. Second, 
+you can pass DOIs directly to `ft_links()`. Both end up at the same point, links
+for each article, if they could be found for the user selected data source.
+
+The biggest caveat with `ft_links()` is that we can't gaurantee that the links
+will work. Link rot is one way in which the links may not work: link rot is when 
+the URL does not point to the original content anymore, or fails altogether. 
+Additionally, with Crossref, publishers can deposit URLs for articles, but they 
+make change the URLs at some later date but not update the URLs with Crossref.
+
+## Abstracts
+
+`ft_abstract()` provides access to article abstracts from four different 
+data sources (PLOS, Scopus, Microsoft Academic Research, Crossref). The only 
+way to use the function is to pass article identifiers, which are for the most 
+DOIs.
+
+The advantage of abstracts over full text is that abstracts can often be 
+retrieved even for paywalled articles. That is, you can have much broader
+coverage of the articles you're targeting relative to full text.
+
+If you are after abstracts, and you are already getting or already have full text, 
+and if the articles are in XML format, then you can use [pubchunks][] to extract
+out the abstracts.
+
+## Fetch full text
+
+`ft_get()` fetchs full text of articles from many different data sources. From the DOIs
+that are passed in to the function, we detect the publisher, and there are specific plugins
+for certain publishers:
+
+- aaas
+- aip
+- amersocclinoncol
+- amersocmicrobiol
+- arxiv
+- biorxiv
+- bmc
+- copernicus
+- crossref
+- elife
+- elsevier
+- entrez
+- frontiersin
+- ieee
+- informa
+- instinvestfil
+- jama
+- microbiology
+- peerj
+- pensoft
+- plos
+- pnas
+- royalsocchem
+- sciencedirect
+- scientificsocieties
+- wiley
+
+If there's no built-in plugin for the publisher already, we use the FTDOI API 
+(<https://ftdoi.org>) to try to get the link for the full text of the article. If 
+the FTDOI API doesn't bear fruit, we search Crossref for a link to the full text.
+If Crossref doesn't have any full text links, we give up.
+
+Since users can go through a lot of article requests, we cache successfully 
+downloaded articles, and keep that knowledge consistent across R sessions; all 
+subsequent requests for the same article just use the cached version. Additionally,
+all errors in `ft_get()` are collected in a tidy data.frame in the output of the 
+function to help the user quickly determine what went wrong.
+
 # How to text mine from R: Three case studies
 
-### Case study 1: Citation mining
+## Case study 1: Citation mining
 
@@
 
-### Case study 2: Abstract mining
+## Case study 2: Abstract mining
 
@@
 
-### Case study 3: Full text mining
+## Case study 3: Full text mining
 

### 2019-04-08 manuscript.tex
paper updates

@@
             pdfauthor={},
-            pdftitle={rOpenSci tools for accessing science literature for textmining},
+            pdftitle={rOpenSci tools for accessing research literature for text mining},
             colorlinks=true,
@@
 
-  \title{rOpenSci tools for accessing science literature for textmining}
+  \title{rOpenSci tools for accessing research literature for text mining}
     \author[cstar]{Scott Chamberlain\corref{c1}}
@@
 
+\hypertarget{fulltext-a-swiss-army-knife-for-text-mining-in-r}{%
+\section{fulltext: a swiss army knife for text mining in
+R}\label{fulltext-a-swiss-army-knife-for-text-mining-in-r}}
+
+\href{https://github.com/ropensci/fulltext}{fulltext} is a general
+purpose R package for the data part of text-mining: search for articles,
+get links to articles, get article abstracts, and fetch full text of
+articles. The \texttt{fulltext} package is always adding additional data
+sources as time allows (See Table 1). Starting from searching for
+articles, the outputs of search can be fed into a function to get links
+to those articles, or to get abstracts for those articles, or to fetch
+their full text.
+
+The following is a breakdown of the major distinct functional parts of
+\texttt{fulltext}.
+
+\hypertarget{search}{%
+\subsection{Search}\label{search}}
+
+\texttt{ft\_search()} provides search access to nine different data
+sources (PLOS, BMC, Crossref, Entrez, arXiv, bioRxiv, Europe PMC,
+Scopus, Microsoft Academic), creating a mostly unified interface to all
+data sources. The parts of each data source that are common are mostly
+factored into the parameters of the \texttt{ft\_search()} function, and
+we also allow the user to pass on data source specific options as
+needed.
+
+xxxxx
+
+\hypertarget{links}{%
+\subsection{Links}\label{links}}
+
+\texttt{ft\_links()} provides two pathways to get links (URLs) for
+articles, with a choice of four different data sources (PLOS, BMC,
+Crossref, Entrez). First, you can use \texttt{ft\_search()}, then pass
+the output of that function to \texttt{ft\_links()}. Second, you can
+pass DOIs directly to \texttt{ft\_links()}. Both end up at the same
+point, links for each article, if they could be found for the user
+selected data source.
+
+The biggest caveat with \texttt{ft\_links()} is that we can't gaurantee
+that the links will work. Link rot is one way in which the links may not
+work: link rot is when the URL does not point to the original content
+anymore, or fails altogether. Additionally, with Crossref, publishers
+can deposit URLs for articles, but they make change the URLs at some
+later date but not update the URLs with Crossref.
+
+\hypertarget{abstracts}{%
+\subsection{Abstracts}\label{abstracts}}
+
+\texttt{ft\_abstract()} provides access to article abstracts from four
+different data sources (PLOS, Scopus, Microsoft Academic Research,
+Crossref). The only way to use the function is to pass article
+identifiers, which are for the most DOIs.
+
+The advantage of abstracts over full text is that abstracts can often be
+retrieved even for paywalled articles. That is, you can have much
+broader coverage of the articles you're targeting relative to full text.
+
+If you are after abstracts, and you are already getting or already have
+full text, and if the articles are in XML format, then you can use
+\href{https://github.com/ropensci/pubchunks}{pubchunks} to extract out
+the abstracts.
+
+\hypertarget{fetch-full-text}{%
+\subsection{Fetch full text}\label{fetch-full-text}}
+
+\texttt{ft\_get()} fetchs full text of articles from many different data
+sources. From the DOIs that are passed in to the function, we detect the
+publisher, and there are specific plugins for certain publishers:
+
+\begin{itemize}
+\tightlist
+\item
+  aaas
+\item
+  aip
+\item
+  amersocclinoncol
+\item
+  amersocmicrobiol
+\item
+  arxiv
+\item
+  biorxiv
+\item
+  bmc
+\item
+  copernicus
+\item
+  crossref
+\item
+  elife
+\item
+  elsevier
+\item
+  entrez
+\item
+  frontiersin
+\item
+  ieee
+\item
+  informa
+\item
+  instinvestfil
+\item
+  jama
+\item
+  microbiology
+\item
+  peerj
+\item
+  pensoft
+\item
+  plos
+\item
+  pnas
+\item
+  royalsocchem
+\item
+  sciencedirect
+\item
+  scientificsocieties
+\item
+  wiley
+\end{itemize}
+
+If there's no built-in plugin for the publisher already, we use the
+FTDOI API (\url{https://ftdoi.org}) to try to get the link for the full
+text of the article. If the FTDOI API doesn't bear fruit, we search
+Crossref for a link to the full text. If Crossref doesn't have any full
+text links, we give up.
+
+Since users can go through a lot of article requests, we cache
+successfully downloaded articles, and keep that knowledge consistent
+across R sessions; all subsequent requests for the same article just use
+the cached version. Additionally, all errors in \texttt{ft\_get()} are
+collected in a tidy data.frame in the output of the function to help the
+user quickly determine what went wrong.
+
 \hypertarget{how-to-text-mine-from-r-three-case-studies}{%
@@
 \hypertarget{case-study-1-citation-mining}{%
-\subsubsection{Case study 1: Citation
+\subsection{Case study 1: Citation
 mining}\label{case-study-1-citation-mining}}
@@
 \hypertarget{case-study-2-abstract-mining}{%
-\subsubsection{Case study 2: Abstract
+\subsection{Case study 2: Abstract
 mining}\label{case-study-2-abstract-mining}}
@@
 \hypertarget{case-study-3-full-text-mining}{%
-\subsubsection{Case study 3: Full text
+\subsection{Case study 3: Full text
 mining}\label{case-study-3-full-text-mining}}

### 2019-04-12 manuscript.Rmd
work on ms

@@
 linenumbers: true
-title: "rOpenSci tools for accessing research literature for text mining"
+title: "R tools for accessing research literature for text mining"
 author:
@@
 abstract: |
-      Corresponding Author:
-
-      Scott Chamberlain
-
-      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
-
-      Email address: [myrmecocystus@gmail.com](mailto:myrmecocystus@gmail.com)
-
-      \newpage
-
-      Background. xxxx.
-
-      Methods. xxxx.
-
-      Results. xxxx.
-
-      Discussion. xxxx.
-
+  Text-mining is a powerful method for answering research questions. However,
+  getting texts to extract information can be a daunting and complicated task.
+  The primary reason for this is publishers. There are thousands of different
+  publishers, each with their own licenses, URL patterns, access options, and
+  more. Layered on top of that is the varied access each user has based on
+  their institutional affiliation. Here, I introduce a suite of software packages
+  in the R programming language for fetching texts. The tapestry of different
+  publishers, access levels, and other factors requires a patchwork of approaches
+  for getting texts to users. The flagship R package called fulltext attempts
+  to simplify search and retrieval of texts for text-mining. The fulltext package,
+  along with many others, make acquiring texts easier than ever, facilitating
+  answering research questions with text-mining.
 bibliography: components/references.bib
@@
 documentclass: components/elsarticle
-
 output:
@@
 
-There's more than 100 million articles published
-(source: Crossref API), representing an enormous amount
-of knowledge.  In addition to simply reading these articles,
-they contain a vast trove of information of interest to
-researchers for machine aided questions.
+There's more than 100 million research articles published
+(Crossref API: <https://github.com/CrossRef/rest-api-doc>), representing
+an enormous amount of knowledge.  In addition to simply reading these
+articles, the articles contain a vast trove of information of interest to
+researchers for machine aided questions [@Kong2018; @Usai2018].
 
@@
 out of articles. Given the importance of text-mining, good
-text-mining tools are needed to make it easier for researchers.
-In particular, the R programming language is used widely throughout
-many academic fields and thus tools in R for text mining are of
-particular importance.
+text-mining tools are needed to make it easier for researchers to do.
+Graphical user interface (GUI) based text mining tools are available
+[e.g., @ba2016interoperability] and some research papers have used them
+[@Chaix2019], but given the urgent recent call to action for more reproducible
+research [@OSC2015; @Camerer2016; @Camerer2018],
+we must move away from GUI based tools as fast as possible.
 
-Here, we present an overview of text-mining tools in the R programming
-language. We do not cover analysis tools per se, but rather those
-tools for searching for, acquiring, and "mashing up" text.
+In particular, the R programming language is used widely throughout many
+academic fields and thus tools in R for text mining are of particular
+importance because they can be adopted by academics rapidly.
+
+Here, I present an overview of text-mining tools in the R programming
+language, not for text-mining analysis, but rather those
+tools for searching for, acquiring, and extracting particular chunks
+of texts (e.g., title, abstract, authors). Most of the packages are
+part of the rOpenSci suite (https://ropensci.org/).
 
@@
 
-Those articles that are digital can be split into two groups:
-easily machine readable and non-machine readable.
+Those articles that are digital (which in theory includes all articles)
+can be split into two groups: machine readable and non-machine
+readable.
 
-The machine readable articles are those in XML, JSON, or plain text
-format. The former two, XML and JSON, are ideal for the machine
-readable types because they are structured data, whereas
+The machine readable articles are those in XML[^xml], JSON[^json], or
+plain text format. The former two, XML and JSON, are the best machine
+readable types because they are structured data[^sd], whereas
 plain text has no structure - it's simply a set of characters
@@
 
-Of the non-machine readable kind, there's PDFs. These can be broken out
-into two groups: text based PDFs and scanned PDFs. The former are
-converted from digital versions of various kinds (MS Word, OpenOffice,
-markdown, etc.), while the latter are PDFs created by scanning
-in print articles for which there is no digital version.
+Of the non-machine readable types, the most noteable is the Portable
+Document Format (PDF)[^pdf]. These can be broken out into two groups: text based
+PDFs and scanned PDFs. The former are converted from digital versions of
+various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
+latter are created by scanning print articles to a PDF format. Text-based
+PDFs are much better for text-mining purposes as plain text can be extracted
+easily in R with [pdftools][], a binding to [libpoppler][]. However, with
+scanned PDFs, text must be extracted using Optimal Character Recognition
+(OCR; see R package [tesseract][]), which isn't always a clean solution,
+especially compared to true text based PDFs.
+
+The reality in scholarly publishing is all publishers, if they provide
+any access to their articles, only provide PDF format. Very few publishers,
+with some quite large (Elsevier, Pensoft, PLOS), provide XML format. Although
+most publishers most likely have the XML behind each of their articles,
+they for some indefensible reason do not share it - making text-mining more
+difficult. Some provide plain text (Elsevier). I only know of one publisher
+that provides full text as JSON (PLOS). Thus, text-mining, in most cases,
+will require extracting text from PDFs.
+
+[^xml]: https://www.w3.org/TR/xml/
+[^json]: https://tools.ietf.org/html/rfc7159
+[^sd]: https://en.wikipedia.org/wiki/Data_model
+[^pdf]: https://en.wikipedia.org/wiki/PDF
 
@@
 get by with article metadata (authors, title, etc.), some with abstracts,
-but many use cases need full-text.
+but many use cases require full-text.
 
@@
 
-A promising sign is an increasing number of open access publishers, yet
-these represent a very small portion of the total articles (XXXXX) (ref.).
+A promising sign is an increasing number of open access articles, yet
+open access articles represent a small percent of all articles: an estimate in
+2018 said that 28% of the scholarly literature was open access [@Piwowar2018].
 
 With respect to paywalled articles, access varies by institution, depending
-on what each institution decided to pay for. In addition, some users may
-not realize access varies with IP address so that access from campus vs.
-from home (if not on a VPN) will drastically differ.
+on each institution's publisher contracts. MORE ABOUT THIS ...
 
-We can not end this section without mentioning SciHub. This is a last
-resort option for many probably, providing access to full text of articles
-that are normally paywalled. No tools in this manuscript provide access
-to SciHub.
+Some may not realize access to articles varies with IP address so that
+access from campus vs. from home (if not on a VPN) will drastically differ.
+Sometimes a VPN is required, and this can provide a significant technical
+hurdle to users attempting to do text-mining work.
+
+One final hurdle in text-mining comes unsurprisingly from Elsevier. They
+use so-called "fences" for programmatic access. That is, even if a person
+trying to get an article programmatically their institution has access to
+and they have access to, and they are on the correct IP address, they may
+still not get access to an Elsevier article. Elsevier puts in place these
+fences and only if you contact their technical team directly can you
+get these fences removed, and only then on a per institution basis.
+
+I can not end this section without mentioning SciHub. This is a last
+resort option for many probably (or possibly first, depending on your level
+of access), providing access to full text of articles that are normally
+paywalled. No tools in this manuscript provide access to SciHub.
 
@@
 
-xxx
+A text-mining project starts with a question. From that question,
+researchers then attempt to acquire scholarly articles for
+text-mining. Finding those articles is not altogether straight-forward.
 
-#
+There are many places to search for articles; a non-exhaustive list:
+Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect,
+Web of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access Journals,
+Open Knowledge Maps, and more. It's probably difficult to know where the
+best place is to search. Some of these are paywalled (e.g., Web of Science),
+and some are not.
 
-xxx
+The most important aspect about any source for article search with respect to
+reproducible research is being able to use the data source programmatically.
+Of those listed above, the following can  be used programmatically:
+Microsoft Academic Research, Scopus, ScienceDirect, Pubmed/Entrez, Europe PMC,
+and Directory of Open Access Journals. All of these are included in the R
+package [fulltext][], discussed further below.
 
@@
 
-xxxxx
+Directory of Open Access Journals (DOAJ) maintains data on open access journals, as well
+as some portion of the articles in those journals. Thus, you can search for journals as
+well as articles with DOAJ. The R interface for DOAJ is [jaod][].
 
@@
 
-xxxxx
+JSTOR's Data for Research program gives institutions with access to JSTOR, access
+to full text of articles within JSTOR. There is no way however to make the interaction
+with JSTOR completely programmatic, thus making reproducible research very difficult.
+Nonetheless, there is an R package ([jstor][]) for using data from JSTOR's Data
+for Research.
 
@@
 
-xxxxx
+ORCID (https://orcid.org/) is an organization keeping track of identifiers and metadata
+for researchers around the world. Individuals can optionally maintain metadata on their
+scholarly works connected to their account with ORCID. Thus, across all of ORCID, a
+significant cache of metadata is accruing on scholarly works, their funding amounts,
+collaborators, etc., useful for bibliometrics research and more. The R interface for
+ORCID is [rorcid][].
 
@@
 
-xxx
+The Open Citation Corpus (http://opencitations.net/) holds records of which articles
+cite which other articles, allowing for all important research on the scholarly
+web of citation. Citation data has been very closely guarded until recently, but the
+largest publishers are still not contributing to the Open Citation Corpus. The R
+interface to the Open Citation Corpus is [rcitoid][].
 
@@
 
-The following is a breakdown of the major distinct functional parts of 
-`fulltext`.
+The following is a breakdown of the major distinct parts of `fulltext`.
 
@@
 mostly unified interface to all data sources. The parts of each data source
-that are common are mostly factored into the parameters of the `ft_search()` function,
-and we also allow the user to pass on data source specific options as needed.
+that are common are for the most part factored out into the parameters of the
+`ft_search()` function: query term(s), pagination (number of results, result
+number to start at). In addition, we allow the user to pass on data source specific
+options to refine the search per data source.
 
-xxxxx
+With `ft_search()`, you can query any combination of the nine data sources at once.
+The returned object is a list, with access to results of each data source by its name
+(e.g., `$plos`, or `$crossref`). For each data source, the returned object does vary
+because the returned data from each data source widely varies; for the most part
+data.frame's are returned. For those data sources not queried, their slot is empty.
+
+One important aspect of the research result we highlight is the licenses in the
+returned data for each data source.
+
+```{r}
+x <- ft_search(query = 'ecology', from = c("plos", "crossref"))
+```
+
+The results for this PLOS search have all CC-BY licenses
+
+```{r}
+x$plos
+```
+
+Whereas the results for this Crossref search have mixed licenses
+
+```{r}
+x$crossref
+```
+
+You can dig into the license field for each article, with URLs holding information
+on each license
+
+```{r}
+vapply(x$crossref$data$license, function(w) w$URL[1], "")
+```
 
@@
 of four different data sources (PLOS, BMC, Crossref, Entrez). First, you can
-use `ft_search()`, then pass the output of that function to `ft_links()`. Second, 
-you can pass DOIs directly to `ft_links()`. Both end up at the same point, links
+use `ft_search()`, then pass the output of that function to `ft_links()`.
+
+```{r}
+out <- ft_search(query = "ecology", from = "entrez")
+ft_links(out)
+```
+
+Second, you can pass DOIs directly to `ft_links()`. Both end up at the same point, links
 for each article, if they could be found for the user selected data source.
+
+```{r eval=FALSE}
+# FIXME
+ft_links(out$entrez$data$doi)
+```
 
@@
 that are passed in to the function, we detect the publisher, and there are specific plugins
-for certain publishers:
-
-- aaas
-- aip
-- amersocclinoncol
-- amersocmicrobiol
-- arxiv
-- biorxiv
-- bmc
-- copernicus
-- crossref
-- elife
-- elsevier
-- entrez
-- frontiersin
-- ieee
-- informa
-- instinvestfil
-- jama
-- microbiology
-- peerj
-- pensoft
-- plos
-- pnas
-- royalsocchem
-- sciencedirect
-- scientificsocieties
-- wiley
+for certain publishers: AAAS, American Institute of Physics, American Society of
+Clinical Oncology, American Society for Microbiology, arXiv, bioRxiv, BiomedCentral,
+Copernicus, Crossref, Elife, Elsevier, Pubmed/PMC via NCBI's Entrez, Frontiers, IEEE,
+Informa, Instituto de Investigaciones Filologicas, American Medical Association,
+Microbiology Society, PeerJ, Pensoft, PLOS, PNAS, Royal Society of Chemistry,
+ScienceDirect, Scientific Societies, and Wiley.
 
@@
 
-xxxx
+To make text mining easier:
+
+1. publishers should provide XML if they have it
+2. publishers should not change URL patterns so often, or at all
+3. publishers should keep their Crossref metadata up to date
+4. remove all paywalls (Easy, yes?)
+5. ...
 
@@
 [pubchunks]: https://github.com/ropensci/pubchunks
+[pdftools]: https://github.com/ropensci/pdftools
+[libpoppler]: https://poppler.freedesktop.org/
+[tesseract]: https://github.com/ropensci/tesseract

### 2019-04-12 manuscript.tex
work on ms

@@
             pdfauthor={},
-            pdftitle={rOpenSci tools for accessing research literature for text mining},
+            pdftitle={R tools for accessing research literature for text mining},
             colorlinks=true,
@@
 
-  \title{rOpenSci tools for accessing research literature for text mining}
+  \title{R tools for accessing research literature for text mining}
     \author[cstar]{Scott Chamberlain\corref{c1}}
@@
   \begin{abstract}
-  Corresponding Author:
-
-  Scott Chamberlain
-
-  rOpenSci, Museum of Paleontology, University of California, Berkeley,
-  CA, USA
-
-  Email address:
-  \href{mailto:myrmecocystus@gmail.com}{\nolinkurl{myrmecocystus@gmail.com}}
-
-  \newpage
-
-  Background. xxxx.
-
-  Methods. xxxx.
-
-  Results. xxxx.
-
-  Discussion. xxxx.
+  Text-mining is a powerful method for answering research questions.
+  However, getting texts to extract information can be a daunting and
+  complicated task. The primary reason for this is publishers. There are
+  thousands of different publishers, each with their own licenses, URL
+  patterns, access options, and more. Layered on top of that is the varied
+  access each user has based on their institutional affiliation. Here, I
+  introduce a suite of software packages in the R programming language for
+  fetching texts. The tapestry of different publishers, access levels, and
+  other factors requires a patchwork of approaches for getting texts to
+  users. The flagship R package called fulltext attempts to simplify
+  search and retrieval of texts for text-mining. The fulltext package,
+  along with many others, make acquiring texts easier than ever,
+  facilitating answering research questions with text-mining.
   \end{abstract}
@@
 
-There's more than 100 million articles published (source: Crossref API),
-representing an enormous amount of knowledge. In addition to simply
-reading these articles, they contain a vast trove of information of
-interest to researchers for machine aided questions.
+There's more than 100 million research articles published (Crossref API:
+\url{https://github.com/CrossRef/rest-api-doc}), representing an
+enormous amount of knowledge. In addition to simply reading these
+articles, the articles contain a vast trove of information of interest
+to researchers for machine aided questions (Kong \& Gerstein, 2018; Usai
+et al., 2018).
 
@@
 articles. Given the importance of text-mining, good text-mining tools
-are needed to make it easier for researchers. In particular, the R
-programming language is used widely throughout many academic fields and
-thus tools in R for text mining are of particular importance.
+are needed to make it easier for researchers to do. Graphical user
+interface (GUI) based text mining tools are available (e.g., Ba \&
+Bossy, 2016) and some research papers have used them (Chaix et al.,
+2019), but given the urgent recent call to action for more reproducible
+research (Open Science Collaboration, 2015; Camerer et al., 2016, 2018),
+we must move away from GUI based tools as fast as possible.
 
-Here, we present an overview of text-mining tools in the R programming
-language. We do not cover analysis tools per se, but rather those tools
-for searching for, acquiring, and ``mashing up'' text.
+In particular, the R programming language is used widely throughout many
+academic fields and thus tools in R for text mining are of particular
+importance because they can be adopted by academics rapidly.
+
+Here, I present an overview of text-mining tools in the R programming
+language, not for text-mining analysis, but rather those tools for
+searching for, acquiring, and extracting particular chunks of texts
+(e.g., title, abstract, authors). Most of the packages are part of the
+rOpenSci suite (\url{https://ropensci.org/}).
 
@@
 
-Those articles that are digital can be split into two groups: easily
-machine readable and non-machine readable.
+Those articles that are digital (which in theory includes all articles)
+can be split into two groups: machine readable and non-machine readable.
 
-The machine readable articles are those in XML, JSON, or plain text
-format. The former two, XML and JSON, are ideal for the machine readable
-types because they are structured data, whereas plain text has no
-structure - it's simply a set of characters with line breaks and spaces
-in between.
+The machine readable articles are those in XML\footnote{\url{https://www.w3.org/TR/xml/}},
+JSON\footnote{\url{https://tools.ietf.org/html/rfc7159}}, or plain text
+format. The former two, XML and JSON, are the best machine readable
+types because they are structured data\footnote{\url{https://en.wikipedia.org/wiki/Data_model}},
+whereas plain text has no structure - it's simply a set of characters
+with line breaks and spaces in between.
 
-Of the non-machine readable kind, there's PDFs. These can be broken out
-into two groups: text based PDFs and scanned PDFs. The former are
-converted from digital versions of various kinds (MS Word, OpenOffice,
-markdown, etc.), while the latter are PDFs created by scanning in print
-articles for which there is no digital version.
+Of the non-machine readable types, the most noteable is the Portable
+Document Format (PDF)\footnote{\url{https://en.wikipedia.org/wiki/PDF}}.
+These can be broken out into two groups: text based PDFs and scanned
+PDFs. The former are converted from digital versions of various kinds
+(MS Word, OpenOffice, LaTeX, markdown, etc.), while the latter are
+created by scanning print articles to a PDF format. Text-based PDFs are
+much better for text-mining purposes as plain text can be extracted
+easily in R with \href{https://github.com/ropensci/pdftools}{pdftools},
+a binding to \href{https://poppler.freedesktop.org/}{libpoppler}.
+However, with scanned PDFs, text must be extracted using Optimal
+Character Recognition (OCR; see R package
+\href{https://github.com/ropensci/tesseract}{tesseract}), which isn't
+always a clean solution, especially compared to true text based PDFs.
+
+The reality in scholarly publishing is all publishers, if they provide
+any access to their articles, only provide PDF format. Very few
+publishers, with some quite large (Elsevier, Pensoft, PLOS), provide XML
+format. Although most publishers most likely have the XML behind each of
+their articles, they for some indefensible reason do not share it -
+making text-mining more difficult. Some provide plain text (Elsevier). I
+only know of one publisher that provides full text as JSON (PLOS). Thus,
+text-mining, in most cases, will require extracting text from PDFs.
 
@@
 get by with article metadata (authors, title, etc.), some with
-abstracts, but many use cases need full-text.
+abstracts, but many use cases require full-text.
 
@@
 
-A promising sign is an increasing number of open access publishers, yet
-these represent a very small portion of the total articles (XXXXX)
-(ref.).
+A promising sign is an increasing number of open access articles, yet
+open access articles represent a small percent of all articles: an
+estimate in 2018 said that 28\% of the scholarly literature was open
+access (Piwowar et al., 2018).
 
 With respect to paywalled articles, access varies by institution,
-depending on what each institution decided to pay for. In addition, some
-users may not realize access varies with IP address so that access from
-campus vs.~from home (if not on a VPN) will drastically differ.
+depending on each institution's publisher contracts. MORE ABOUT THIS
+\ldots{}
 
-We can not end this section without mentioning SciHub. This is a last
-resort option for many probably, providing access to full text of
-articles that are normally paywalled. No tools in this manuscript
-provide access to SciHub.
+Some may not realize access to articles varies with IP address so that
+access from campus vs.~from home (if not on a VPN) will drastically
+differ. Sometimes a VPN is required, and this can provide a significant
+technical hurdle to users attempting to do text-mining work.
+
+One final hurdle in text-mining comes unsurprisingly from Elsevier. They
+use so-called ``fences'' for programmatic access. That is, even if a
+person trying to get an article programmatically their institution has
+access to and they have access to, and they are on the correct IP
+address, they may still not get access to an Elsevier article. Elsevier
+puts in place these fences and only if you contact their technical team
+directly can you get these fences removed, and only then on a per
+institution basis.
+
+I can not end this section without mentioning SciHub. This is a last
+resort option for many probably (or possibly first, depending on your
+level of access), providing access to full text of articles that are
+normally paywalled. No tools in this manuscript provide access to
+SciHub.
 
@@
 
-xxx
+A text-mining project starts with a question. From that question,
+researchers then attempt to acquire scholarly articles for text-mining.
+Finding those articles is not altogether straight-forward.
 
-\hypertarget{section}{%
-\section{}\label{section}}
+There are many places to search for articles; a non-exhaustive list:
+Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect, Web
+of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access
+Journals, Open Knowledge Maps, and more. It's probably difficult to know
+where the best place is to search. Some of these are paywalled (e.g.,
+Web of Science), and some are not.
 
-xxx
+The most important aspect about any source for article search with
+respect to reproducible research is being able to use the data source
+programmatically. Of those listed above, the following can be used
+programmatically: Microsoft Academic Research, Scopus, ScienceDirect,
+Pubmed/Entrez, Europe PMC, and Directory of Open Access Journals. All of
+these are included in the R package
+\href{https://github.com/ropensci/fulltext}{fulltext}, discussed further
+below.
 
@@
 
-xxxxx
+Directory of Open Access Journals (DOAJ) maintains data on open access
+journals, as well as some portion of the articles in those journals.
+Thus, you can search for journals as well as articles with DOAJ. The R
+interface for DOAJ is \href{https://github.com/ropensci/jaod}{jaod}.
 
@@
 
-xxxxx
+JSTOR's Data for Research program gives institutions with access to
+JSTOR, access to full text of articles within JSTOR. There is no way
+however to make the interaction with JSTOR completely programmatic, thus
+making reproducible research very difficult. Nonetheless, there is an R
+package (\href{https://github.com/ropensci/jstor}{jstor}) for using data
+from JSTOR's Data for Research.
 
@@
 
-xxxxx
+ORCID (\url{https://orcid.org/}) is an organization keeping track of
+identifiers and metadata for researchers around the world. Individuals
+can optionally maintain metadata on their scholarly works connected to
+their account with ORCID. Thus, across all of ORCID, a significant cache
+of metadata is accruing on scholarly works, their funding amounts,
+collaborators, etc., useful for bibliometrics research and more. The R
+interface for ORCID is
+\href{https://github.com/ropensci/rorcid}{rorcid}.
 
@@
 
-xxx
+The Open Citation Corpus (\url{http://opencitations.net/}) holds records
+of which articles cite which other articles, allowing for all important
+research on the scholarly web of citation. Citation data has been very
+closely guarded until recently, but the largest publishers are still not
+contributing to the Open Citation Corpus. The R interface to the Open
+Citation Corpus is
+\href{https://github.com/ropenscilabs/rcitoid}{rcitoid}.
 
@@
 
-The following is a breakdown of the major distinct functional parts of
+The following is a breakdown of the major distinct parts of
 \texttt{fulltext}.
@@
 Scopus, Microsoft Academic), creating a mostly unified interface to all
-data sources. The parts of each data source that are common are mostly
-factored into the parameters of the \texttt{ft\_search()} function, and
-we also allow the user to pass on data source specific options as
-needed.
+data sources. The parts of each data source that are common are for the
+most part factored out into the parameters of the \texttt{ft\_search()}
+function: query term(s), pagination (number of results, result number to
+start at). In addition, we allow the user to pass on data source
+specific options to refine the search per data source.
 
-xxxxx
+With \texttt{ft\_search()}, you can query any combination of the nine
+data sources at once. The returned object is a list, with access to
+results of each data source by its name (e.g., \texttt{\$plos}, or
+\texttt{\$crossref}). For each data source, the returned object does
+vary because the returned data from each data source widely varies; for
+the most part data.frame's are returned. For those data sources not
+queried, their slot is empty.
+
+One important aspect of the research result we highlight is the licenses
+in the returned data for each data source.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \KeywordTok{c}\NormalTok{(}\StringTok{"plos"}\NormalTok{, }\StringTok{"crossref"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+The results for this PLOS search have all CC-BY licenses
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x}\OperatorTok{$}\NormalTok{plos}
+\CommentTok{#> Query: [ecology] }
+\CommentTok{#> Records found, returned: [47257, 10] }
+\CommentTok{#> License: [CC-BY] }
+\CommentTok{#>                              id}
+\CommentTok{#> 1  10.1371/journal.pone.0001248}
+\CommentTok{#> 2  10.1371/journal.pone.0059813}
+\CommentTok{#> 3  10.1371/journal.pone.0080763}
+\CommentTok{#> 4  10.1371/journal.pone.0155019}
+\CommentTok{#> 5  10.1371/journal.pone.0175014}
+\CommentTok{#> 6  10.1371/journal.pone.0150648}
+\CommentTok{#> 7  10.1371/journal.pone.0208370}
+\CommentTok{#> 8  10.1371/journal.pcbi.1003594}
+\CommentTok{#> 9  10.1371/journal.pone.0102437}
+\CommentTok{#> 10 10.1371/journal.pone.0166559}
+\end{Highlighting}
+\end{Shaded}
+
+Whereas the results for this Crossref search have mixed licenses
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x}\OperatorTok{$}\NormalTok{crossref}
+\CommentTok{#> Query: [ecology] }
+\CommentTok{#> Records found, returned: [164657, 10] }
+\CommentTok{#> License: [variable, see individual records] }
+\CommentTok{#>    archive               container.title    created  deposited}
+\CommentTok{#> 1  Portico                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 2  Portico                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 3       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 4       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 5       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 6       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 7       NA                       Ecology 2006-05-09 2018-08-01}
+\CommentTok{#> 8  Portico                       Ecology 2017-04-26 2019-03-08}
+\CommentTok{#> 9       NA Trends in Ecology & Evolution 2002-07-25 2017-06-14}
+\CommentTok{#> 10      NA Journal of Industrial Ecology 2014-11-21 2017-06-23}
+\CommentTok{#> Variables not shown: published.print (chr), published.online (chr), doi}
+\CommentTok{#>      (chr), indexed (chr), issn (chr), issue (chr), issued (chr), member}
+\CommentTok{#>      (chr), page (chr), prefix (chr), publisher (chr), reference.count}
+\CommentTok{#>      (chr), score (chr), source (chr), title (chr), type (chr), url (chr),}
+\CommentTok{#>      volume (chr), author (list), link (list), license (list), subject}
+\CommentTok{#>      (chr), alternative.id (chr), subtitle (chr), reference (list)}
+\end{Highlighting}
+\end{Shaded}
+
+You can dig into the license field for each article, with URLs holding
+information on each license
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{vapply}\NormalTok{(x}\OperatorTok{$}\NormalTok{crossref}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{license, }\ControlFlowTok{function}\NormalTok{(w) w}\OperatorTok{$}\NormalTok{URL[}\DecValTok{1}\NormalTok{], }\StringTok{""}\NormalTok{)}
+\CommentTok{#>  [1] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [2] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [3] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [4] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [5] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [6] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [7] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [8] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [9] "http://www.elsevier.com/tdm/userlicense/1.0/"}
+\CommentTok{#> [10] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\end{Highlighting}
+\end{Shaded}
 
@@
 Crossref, Entrez). First, you can use \texttt{ft\_search()}, then pass
-the output of that function to \texttt{ft\_links()}. Second, you can
-pass DOIs directly to \texttt{ft\_links()}. Both end up at the same
-point, links for each article, if they could be found for the user
-selected data source.
+the output of that function to \texttt{ft\_links()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"entrez"}\NormalTok{)}
+\KeywordTok{ft_links}\NormalTok{(out)}
+\CommentTok{#> <fulltext links>}
+\CommentTok{#> [Found] 6 }
+\CommentTok{#> [IDs] ID_30964001 ID_30962485 ID_30962432 ID_30952928 ID_30674747}
+\CommentTok{#>      ID_30674743 ...}
+\end{Highlighting}
+\end{Shaded}
+
+Second, you can pass DOIs directly to \texttt{ft\_links()}. Both end up
+at the same point, links for each article, if they could be found for
+the user selected data source.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# }\AlertTok{FIXME}
+\KeywordTok{ft_links}\NormalTok{(out}\OperatorTok{$}\NormalTok{entrez}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{doi)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 sources. From the DOIs that are passed in to the function, we detect the
-publisher, and there are specific plugins for certain publishers:
-
-\begin{itemize}
-\tightlist
-\item
-  aaas
-\item
-  aip
-\item
-  amersocclinoncol
-\item
-  amersocmicrobiol
-\item
-  arxiv
-\item
-  biorxiv
-\item
-  bmc
-\item
-  copernicus
-\item
-  crossref
-\item
-  elife
-\item
-  elsevier
-\item
-  entrez
-\item
-  frontiersin
-\item
-  ieee
-\item
-  informa
-\item
-  instinvestfil
-\item
-  jama
-\item
-  microbiology
-\item
-  peerj
-\item
-  pensoft
-\item
-  plos
-\item
-  pnas
-\item
-  royalsocchem
-\item
-  sciencedirect
-\item
-  scientificsocieties
-\item
-  wiley
-\end{itemize}
+publisher, and there are specific plugins for certain publishers: AAAS,
+American Institute of Physics, American Society of Clinical Oncology,
+American Society for Microbiology, arXiv, bioRxiv, BiomedCentral,
+Copernicus, Crossref, Elife, Elsevier, Pubmed/PMC via NCBI's Entrez,
+Frontiers, IEEE, Informa, Instituto de Investigaciones Filologicas,
+American Medical Association, Microbiology Society, PeerJ, Pensoft,
+PLOS, PNAS, Royal Society of Chemistry, ScienceDirect, Scientific
+Societies, and Wiley.
 
@@
 \CommentTok{#> Source: /Users/sckott/github/ropensci/textmine/use-cases/* on x86_64 by sckott}
-\CommentTok{#> Created: Fri Apr  5 11:36:04 2019}
+\CommentTok{#> Created: Thu Apr 11 11:20:19 2019}
 \CommentTok{#> Notes:}
@@
 \CommentTok{#> Found:}
-\CommentTok{#>   [PLoS: 47337; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0] }
+\CommentTok{#>   [PLoS: 47272; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0] }
 \CommentTok{#> Returned:}
@@
 \CommentTok{#> Query: [ecology] }
-\CommentTok{#> Records found, returned: [47337, 10] }
+\CommentTok{#> Records found, returned: [47272, 10] }
 \CommentTok{#> License: [CC-BY] }
@@
 \CommentTok{#> 2  10.1371/journal.pone.0059813}
-\CommentTok{#> 3  10.1371/journal.pone.0155019}
-\CommentTok{#> 4  10.1371/journal.pone.0080763}
-\CommentTok{#> 5  10.1371/journal.pone.0208370}
+\CommentTok{#> 3  10.1371/journal.pone.0080763}
+\CommentTok{#> 4  10.1371/journal.pone.0155019}
+\CommentTok{#> 5  10.1371/journal.pone.0175014}
 \CommentTok{#> 6  10.1371/journal.pone.0150648}
-\CommentTok{#> 7  10.1371/journal.pcbi.1003594}
-\CommentTok{#> 8  10.1371/journal.pone.0102437}
-\CommentTok{#> 9  10.1371/journal.pone.0175014}
+\CommentTok{#> 7  10.1371/journal.pone.0208370}
+\CommentTok{#> 8  10.1371/journal.pcbi.1003594}
+\CommentTok{#> 9  10.1371/journal.pone.0102437}
 \CommentTok{#> 10 10.1371/journal.pone.0166559}
@@
 \CommentTok{#> [IDs] 10.1371/journal.pone.0001248 10.1371/journal.pone.0059813}
-\CommentTok{#>      10.1371/journal.pone.0155019 10.1371/journal.pone.0080763}
-\CommentTok{#>      10.1371/journal.pone.0208370 10.1371/journal.pone.0150648}
-\CommentTok{#>      10.1371/journal.pcbi.1003594 10.1371/journal.pone.0102437}
-\CommentTok{#>      10.1371/journal.pone.0175014 10.1371/journal.pone.0166559 ...}
+\CommentTok{#>      10.1371/journal.pone.0080763 10.1371/journal.pone.0155019}
+\CommentTok{#>      10.1371/journal.pone.0175014 10.1371/journal.pone.0150648}
+\CommentTok{#>      10.1371/journal.pone.0208370 10.1371/journal.pcbi.1003594}
+\CommentTok{#>      10.1371/journal.pone.0102437 10.1371/journal.pone.0166559 ...}
 \end{Highlighting}
@@
 \CommentTok{#> }
-\CommentTok{#> $plos$`10.1371/journal.pone.0155019`}
+\CommentTok{#> $plos$`10.1371/journal.pone.0080763`}
 \CommentTok{#>                            doi history.received history.accepted}
-\CommentTok{#> 1 10.1371/journal.pone.0155019       2015-09-22       2016-04-22}
+\CommentTok{#> 1 10.1371/journal.pone.0080763       2013-08-15       2013-10-16}
 \CommentTok{#>   .publisher}
@@
 
-xxxx
+To make text mining easier:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  publishers should provide XML if they have it
+\item
+  publishers should not change URL patterns so often, or at all
+\item
+  publishers should keep their Crossref metadata up to date
+\item
+  remove all paywalls (Easy, yes?)
+\item
+  \ldots{}
+\end{enumerate}
 
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-ba2016interoperability}{}%
+Ba M., Bossy R. 2016. Interoperability of corpus processing work-flow
+engines: The case of alvisnlp/ml in openminted. In: \emph{Proceedings of
+the workshop on cross-platform text mining and natural language
+processing interoperability (interop 2016) at lrec}. 15--18.
+
+\leavevmode\hypertarget{ref-Camerer2016}{}%
+Camerer CF., Dreber A., Forsell E., Ho T-H., Huber J., Johannesson M.,
+Kirchler M., Almenberg J., Altmejd A., Chan T., Heikensten E.,
+Holzmeister F., Imai T., Isaksson S., Nave G., Pfeiffer T., Razen M., Wu
+H. 2016. Evaluating replicability of laboratory experiments in
+economics. \emph{Science} 351:1433--1436.
+
+\leavevmode\hypertarget{ref-Camerer2018}{}%
+Camerer CF., Dreber A., Holzmeister F., Ho T-H., Huber J., Johannesson
+M., Kirchler M., Nave G., Nosek BA., Pfeiffer T., Altmejd A., Buttrick
+N., Chan T., Chen Y., Forsell E., Gampa A., Heikensten E., Hummer L.,
+Imai T., Isaksson S., Manfredi D., Rose J., Wagenmakers E-J., Wu H.
+2018. Evaluating the replicability of social science experiments in
+nature and science between 2010 and 2015. \emph{Nature Human Behaviour}
+2:637--644.
+
+\leavevmode\hypertarget{ref-Chaix2019}{}%
+Chaix E., Deléger L., Bossy R., Nédellec C. 2019. Text mining tools for
+extracting information about microbial biodiversity in food. \emph{Food
+Microbiology} 81:63--75.
+
+\leavevmode\hypertarget{ref-Kong2018}{}%
+Kong X., Gerstein MB. 2018. Text mining systems biology: Turning the
+microscope back on the observer. \emph{Current Opinion in Systems
+Biology} 11:117--122.
+
+\leavevmode\hypertarget{ref-OSC2015}{}%
+Open Science Collaboration. 2015. Estimating the reproducibility of
+psychological science. \emph{Science} 349:aac4716--aac4716.
+
+\leavevmode\hypertarget{ref-Piwowar2018}{}%
+Piwowar H., Priem J., Larivière V., Alperin JP., Matthias L., Norlander
+B., Farley A., West J., Haustein S. 2018. The state of OA: A large-scale
+analysis of the prevalence and impact of open access articles.
+\emph{PeerJ} 6:e4375.
+
+\leavevmode\hypertarget{ref-Usai2018}{}%
+Usai A., Pironti M., Mital M., Mejri CA. 2018. Knowledge discovery out
+of text data: A systematic review via text mining. \emph{Journal of
+Knowledge Management} 22:1471--1488.
 

### 2019-04-16 manuscript.Rmd
updates

@@
 articles, the articles contain a vast trove of information of interest to
-researchers for machine aided questions [@Kong2018; @Usai2018].
-
-For example, many researchers are interested in statistical
-outcomes of articles: questions about P-values, about effect
-sizes, and more. With regard to effect sizes, these are
-of particular interest, as they are often combined in
-meta-analyses to draw broad conclusions about a particular
-question.
+researchers for machine aided questions [@Kong2018; @Usai2018]. For example,
+many researchers are interested in statistical outcomes of articles that 
+can be extracted from numeric results: P-values, effect sizes, means, and
+more. In addition, researchers are often interested in words in articles,
+their use through time, and the contexts they are found in.
 
@@
 Graphical user interface (GUI) based text mining tools are available
-[e.g., @ba2016interoperability] and some research papers have used them
-[@Chaix2019], but given the urgent recent call to action for more reproducible
-research [@OSC2015; @Camerer2016; @Camerer2018],
-we must move away from GUI based tools as fast as possible.
+[e.g., @ba2016interoperability; @Munoz2019; @Canada2017] and some research papers
+have used them [@Chaix2019], but given the urgent recent call to action for
+more reproducible research [@OSC2015; @Camerer2016; @Camerer2018],
+we must move away from GUI based tools as fast as possible. A number of examples
+of programmatic tools can be found in the literature. For example, @Sinclair2016
+present a tool in Python called seqenv for the domain specific task of
+linking sequences to environments through text mining.
 
-In particular, the R programming language is used widely throughout many
-academic fields and thus tools in R for text mining are of particular
-importance because they can be adopted by academics rapidly.
+Most recent text-mining papers do not use programmatic approaches, highlighting
+the need for more programmatic text-mining tools, and increased
+discussion of those tools to increase awareness. For example, many papers 
+search Web of Science using their web interface, and downloading papers manually 
+[@Ding2018; @McCallen2019]. Many of these papers doing GUI based searching and 
+paper downloads are using R or Python downstream for analysis; replacing 
+GUI based data acquisition with programmatic approaches will improve 
+research.
+
+The R programming language is used widely throughout many academic fields;
+tools in R for text mining are of particular importance because they
+can be adopted by academics rapidly.
 
@@
 language, not for text-mining analysis, but rather those
-tools for searching for, acquiring, and extracting particular chunks
-of texts (e.g., title, abstract, authors). Most of the packages are
+tools for searching for, acquiring, and extracting parts of texts
+(e.g., title, abstract, authors). Most of the packages presented here are
 part of the rOpenSci suite (https://ropensci.org/).
@@
 
-A text-mining project starts with a question. From that question,
-researchers then attempt to acquire scholarly articles for
-text-mining. Finding those articles is not altogether straight-forward.
+A text-mining project starts with a question. From that question, researchers
+then attempt to acquire scholarly articles for text-mining. Finding appropriate
+articles is not altogether straight-forward.
 
-There are many places to search for articles; a non-exhaustive list:
+Some of the discovery difficulty relates to the fact that there are so many
+places to search for articles; a non-exhaustive list:
 Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect,
 Web of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access Journals,
-Open Knowledge Maps, and more. It's probably difficult to know where the
-best place is to search. Some of these are paywalled (e.g., Web of Science),
-and some are not.
+Open Knowledge Maps, CORE, Fatcat, and more. It's probably difficult to know
+where the best place is to search. Some of these are paywalled
+(e.g., Web of Science), and some are not.
 
@@
 package [fulltext][], discussed further below.
+
+On top of the vast array of different data sources is the varied ways that 
+search is implemented in each source. Most sources are probably using Solr
+or Elasticsearch under the hood, though we can't know this for sure as most 
+do not make their software infrastructure public knowledge. Nonetheless, data
+sources differ in how search works from the user perspective. For example, some 
+provide wild card search and some do not. Some sources are searching full text
+of articles, while others only search metadata (i.e., title, authors, abstract).
+In addition, each source has a different set of metadata/full text available.
+In brief, the same search against different sources produces different results.
+Some text-mining research articles perform the same search against many
+different sources (refs), while others choose just one source.
 
@@
 Data Provider                      | Content Type            | rOpenSci Package         | Documentation
----------------------------------- | ---------------------   | ------------------------ | -----------------
-Crossref                           | Metadata only           | rcrossref/crminer        | [^1]
-DataCite                           | Metadata only           | rdatacite                | [^2]
+---------------------------------- | ---------------------   | ------------------------ | -------------
+Crossref                           | Metadata                | rcrossref/crminer        | [^1]
+DataCite                           | Metadata                | rdatacite                | [^2]
 Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
-Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
+Public Library of Science (PLoS)   | Full content/altmetrics | rplos                    | [^4]
 Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
@@
 Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
-Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^11]
+Microsoft Academic Search          | Metadata                | microdemic               | [^11]
 Directory of Open Access Journals  | Metadata                | jaod                     | [^12]
@@
 Open Citation Corpus               | Citations               | citecorp                 | [^16]
+Fatcat                             | Metadata                | none                     | [^17]
+SHERPA/RoMEO                       | Journal Level Metadata  | rromeo                   | [^18]
+CORE                               | Full content/Metadata   | rcoreoa                  | [^19]
+Dissemin                           | Metadata                | dissemr                  | [^20]
 
@@
 [^16]: http://opencitations.net/
+[^17]: https://fatcat.wiki/
+[^18]: http://www.sherpa.ac.uk/romeo/apimanual.php?la=en&fIDnum=|&mode=simple
+[^19]: https://core.ac.uk/
+[^20]: https://dissemin.readthedocs.io/en/latest/api.html
 
@@
 
-# fulltext: a swiss army knife for text mining in R
+## Fatcat
+
+Fatcat is a project from Ben Newbold of the Internet Archive Labs. It is a 
+"versioned, publicly-editable catalog of research publications:  journal articles,
+conference proceedings, pre-prints, blog posts". Fatcat is currently does not have an
+R client, but is used inside of the [fulltext][] package.
+
+## SHERPA/RoMEO
+
+SHERPA/RoMEO (http://sherpa.mimas.ac.uk/romeo/index.php) aggregates and analyses
+publisher open access policies and provides summaries of self-archiving permissions
+and conditions of rights given to authors. The [rromeo][] is an R interface
+to SHERPA/RoMEO.
+
+## CORE
+
+CORE (https://core.ac.uk/) touts itself as the world's largest collection of open
+access research articles, providing metadata on journals and articles, as well
+as access to the full text of articles. The [rcoreoa][] R package interfaces
+with the CORE API.
+
+## Dissemin
+
+Dissemin (https://dissem.in/) detects papers behind pay-walls and invites authors
+to upload them to an open repository. Dissemin provides metadata including links to
+open versions of articles. The [dissemr][] R package interfaces with the
+Dissemin API.
+
+# fulltext: a toolset for text mining in R
 
@@
 to those articles, or to get abstracts for those articles, or to fetch
-their full text.
-
-The following is a breakdown of the major distinct parts of `fulltext`.
+their full text. The following is a breakdown of the major distinct
+parts of `fulltext`.
 
@@
 
-To make text mining easier:
+Text-mining will always be a complex task given all the layers involved:
+often temporal time-span of research questions; varied permissions among 
+researchers and their articles they're trying to access; varied approaches
+to getting full text (xml vs pdf vs plain text); and more.
 
-1. publishers should provide XML if they have it
-2. publishers should not change URL patterns so often, or at all
-3. publishers should keep their Crossref metadata up to date
-4. remove all paywalls (Easy, yes?)
-5. ...
+Programmatic text-mining is a first step towards making text-mining easier.
+The R ecosystem is an especially good place to do text-mining because 
+there are many packages for text-mining analysis, and endless packages for
+any required statistical analyses. In addition, rOpenSci and others are 
+building up a set of packages in R for searching for and acquiring full
+text programatically to help make the research workflow as reproducible
+as possible.
+
+Future work for `fulltext` includes:
+
+1. Adding more publisher plugins
+2. Fine tuned user control over publishers 
+3. Improve VPN/proxy controls
+4. Incorporate more search engines to help resolve URLs for fulltext versions
+5. Improve documentation
+
+With respect to what publishers can do to make text-mining easier, publishers
+should:
+
+1. provide XML if they have it
+2. not change URL patterns so often, or at all
+3. maintain consistent URL patterns among journals, years, etc.
+4. keep their Crossref metadata up to date
+5. open up their citation data
 
@@
 [tesseract]: https://github.com/ropensci/tesseract
+[rcoreoa]: https://github.com/ropensci/rcoreoa

### 2019-04-16 manuscript.tex
updates

@@
 to researchers for machine aided questions (Kong \& Gerstein, 2018; Usai
-et al., 2018).
-
-For example, many researchers are interested in statistical outcomes of
-articles: questions about P-values, about effect sizes, and more. With
-regard to effect sizes, these are of particular interest, as they are
-often combined in meta-analyses to draw broad conclusions about a
-particular question.
+et al., 2018). For example, many researchers are interested in
+statistical outcomes of articles that can be extracted from numeric
+results: P-values, effect sizes, means, and more. In addition,
+researchers are often interested in words in articles, their use through
+time, and the contexts they are found in.
 
@@
 interface (GUI) based text mining tools are available (e.g., Ba \&
-Bossy, 2016) and some research papers have used them (Chaix et al.,
-2019), but given the urgent recent call to action for more reproducible
-research (Open Science Collaboration, 2015; Camerer et al., 2016, 2018),
-we must move away from GUI based tools as fast as possible.
-
-In particular, the R programming language is used widely throughout many
-academic fields and thus tools in R for text mining are of particular
-importance because they can be adopted by academics rapidly.
+Bossy, 2016; Cañada et al., 2017; Muñoz, Kissling \& Loon, 2019) and
+some research papers have used them (Chaix et al., 2019), but given the
+urgent recent call to action for more reproducible research (Open
+Science Collaboration, 2015; Camerer et al., 2016, 2018), we must move
+away from GUI based tools as fast as possible. A number of examples of
+programmatic tools can be found in the literature. For example, Sinclair
+et al. (2016) present a tool in Python called seqenv for the domain
+specific task of linking sequences to environments through text mining.
+
+Most recent text-mining papers do not use programmatic approaches,
+highlighting the need for more programmatic text-mining tools, and
+increased discussion of those tools to increase awareness. For example,
+many papers search Web of Science using their web interface, and
+downloading papers manually (Ding, Li \& Fan, 2018; McCallen et al.,
+2019). Many of these papers doing GUI based searching and paper
+downloads are using R or Python downstream for analysis; replacing GUI
+based data acquisition with programmatic approaches will improve
+research.
+
+The R programming language is used widely throughout many academic
+fields; tools in R for text mining are of particular importance because
+they can be adopted by academics rapidly.
 
@@
 language, not for text-mining analysis, but rather those tools for
-searching for, acquiring, and extracting particular chunks of texts
-(e.g., title, abstract, authors). Most of the packages are part of the
+searching for, acquiring, and extracting parts of texts (e.g., title,
+abstract, authors). Most of the packages presented here are part of the
 rOpenSci suite (\url{https://ropensci.org/}).
@@
 researchers then attempt to acquire scholarly articles for text-mining.
-Finding those articles is not altogether straight-forward.
-
-There are many places to search for articles; a non-exhaustive list:
-Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect, Web
-of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access
-Journals, Open Knowledge Maps, and more. It's probably difficult to know
-where the best place is to search. Some of these are paywalled (e.g.,
-Web of Science), and some are not.
+Finding appropriate articles is not altogether straight-forward.
+
+Some of the discovery difficulty relates to the fact that there are so
+many places to search for articles; a non-exhaustive list: Google
+Scholar, Microsoft Academic Research, Scopus, ScienceDirect, Web of
+Science, Pubmed/Entrez, Europe PMC, Directory of Open Access Journals,
+Open Knowledge Maps, CORE, Fatcat, and more. It's probably difficult to
+know where the best place is to search. Some of these are paywalled
+(e.g., Web of Science), and some are not.
 
@@
 
+On top of the vast array of different data sources is the varied ways
+that search is implemented in each source. Most sources are probably
+using Solr or Elasticsearch under the hood, though we can't know this
+for sure as most do not make their software infrastructure public
+knowledge. Nonetheless, data sources differ in how search works from the
+user perspective. For example, some provide wild card search and some do
+not. Some sources are searching full text of articles, while others only
+search metadata (i.e., title, authors, abstract). In addition, each
+source has a different set of metadata/full text available. In brief,
+the same search against different sources produces different results.
+Some text-mining research articles perform the same search against many
+different sources (refs), while others choose just one source.
+
 \hypertarget{data-sources}{%
@@
 \endhead
-Crossref & Metadata only & rcrossref/crminer & \footnote{\url{https://api.crossref.org}}\tabularnewline
-DataCite & Metadata only & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
+Crossref & Metadata & rcrossref/crminer & \footnote{\url{https://api.crossref.org}}\tabularnewline
+DataCite & Metadata & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
 Biodiversity Heritage Library & Full content/Metadata & rbhl &
 \footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
-Public Library of Science (PLoS) & Full text/altmetrics & rplos &
+Public Library of Science (PLoS) & Full content/altmetrics & rplos &
 \footnote{\url{http://api.plos.org/solr}}\tabularnewline
@@
 Europe PMC & Full content/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
-Microsoft Academic Search & Metadata & fulltext/microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
+Microsoft Academic Search & Metadata & microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
 Directory of Open Access Journals & Metadata & jaod & \footnote{\url{https://doaj.org/api/v1/docs}}\tabularnewline
@@
 Open Citation Corpus & Citations & citecorp & \footnote{\url{http://opencitations.net/}}\tabularnewline
+Fatcat & Metadata & none & \footnote{\url{https://fatcat.wiki/}}\tabularnewline
+SHERPA/RoMEO & Journal Level Metadata & rromeo & \footnote{\url{http://www.sherpa.ac.uk/romeo/apimanual.php?la=en\&fIDnum=\%7C\&mode=simple}}\tabularnewline
+CORE & Full content/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
+Dissemin & Metadata & dissemr & \footnote{\url{https://dissemin.readthedocs.io/en/latest/api.html}}\tabularnewline
 \bottomrule
@@
 
-\hypertarget{fulltext-a-swiss-army-knife-for-text-mining-in-r}{%
-\section{fulltext: a swiss army knife for text mining in
-R}\label{fulltext-a-swiss-army-knife-for-text-mining-in-r}}
+\hypertarget{fatcat}{%
+\subsection{Fatcat}\label{fatcat}}
+
+Fatcat is a project from Ben Newbold of the Internet Archive Labs. It is
+a ``versioned, publicly-editable catalog of research publications:
+journal articles, conference proceedings, pre-prints, blog posts''.
+Fatcat is currently does not have an R client, but is used inside of the
+\href{https://github.com/ropensci/fulltext}{fulltext} package.
+
+\hypertarget{sherparomeo}{%
+\subsection{SHERPA/RoMEO}\label{sherparomeo}}
+
+SHERPA/RoMEO (\url{http://sherpa.mimas.ac.uk/romeo/index.php})
+aggregates and analyses publisher open access policies and provides
+summaries of self-archiving permissions and conditions of rights given
+to authors. The {[}rromeo{]}{[}{]} is an R interface to SHERPA/RoMEO.
+
+\hypertarget{core}{%
+\subsection{CORE}\label{core}}
+
+CORE (\url{https://core.ac.uk/}) touts itself as the world's largest
+collection of open access research articles, providing metadata on
+journals and articles, as well as access to the full text of articles.
+The \href{https://github.com/ropensci/rcoreoa}{rcoreoa} R package
+interfaces with the CORE API.
+
+\hypertarget{dissemin}{%
+\subsection{Dissemin}\label{dissemin}}
+
+Dissemin (\url{https://dissem.in/}) detects papers behind pay-walls and
+invites authors to upload them to an open repository. Dissemin provides
+metadata including links to open versions of articles. The
+{[}dissemr{]}{[}{]} R package interfaces with the Dissemin API.
+
+\hypertarget{fulltext-a-toolset-for-text-mining-in-r}{%
+\section{fulltext: a toolset for text mining in
+R}\label{fulltext-a-toolset-for-text-mining-in-r}}
 
@@
 to those articles, or to get abstracts for those articles, or to fetch
-their full text.
-
-The following is a breakdown of the major distinct parts of
-\texttt{fulltext}.
+their full text. The following is a breakdown of the major distinct
+parts of \texttt{fulltext}.
 
@@
 
-To make text mining easier:
+Text-mining will always be a complex task given all the layers involved:
+often temporal time-span of research questions; varied permissions among
+researchers and their articles they're trying to access; varied
+approaches to getting full text (xml vs pdf vs plain text); and more.
+
+Programmatic text-mining is a first step towards making text-mining
+easier. The R ecosystem is an especially good place to do text-mining
+because there are many packages for text-mining analysis, and endless
+packages for any required statistical analyses. In addition, rOpenSci
+and others are building up a set of packages in R for searching for and
+acquiring full text programatically to help make the research workflow
+as reproducible as possible.
+
+Future work for \texttt{fulltext} includes:
 
@@
 \item
-  publishers should provide XML if they have it
+  Adding more publisher plugins
 \item
-  publishers should not change URL patterns so often, or at all
+  Fine tuned user control over publishers
 \item
-  publishers should keep their Crossref metadata up to date
+  Improve VPN/proxy controls
 \item
-  remove all paywalls (Easy, yes?)
+  Incorporate more search engines to help resolve URLs for fulltext
+  versions
 \item
-  \ldots{}
+  Improve documentation
+\end{enumerate}
+
+With respect to what publishers can do to make text-mining easier,
+publishers should:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  provide XML if they have it
+\item
+  not change URL patterns so often, or at all
+\item
+  maintain consistent URL patterns among journals, years, etc.
+\item
+  keep their Crossref metadata up to date
+\item
+  open up their citation data
 \end{enumerate}
@@
 
+\leavevmode\hypertarget{ref-Canada2017}{}%
+Cañada A., Capella-Gutierrez S., Rabal O., Oyarzabal J., Valencia A.,
+Krallinger M. 2017. LimTox: A web tool for applied text mining of
+adverse event and toxicity associations of compounds, drugs and genes.
+\emph{Nucleic Acids Research} 45:W484--W489.
+
 \leavevmode\hypertarget{ref-Chaix2019}{}%
@@
 
+\leavevmode\hypertarget{ref-Ding2018}{}%
+Ding Z., Li Z., Fan C. 2018. Building energy savings: Analysis of
+research trends based on text mining. \emph{Automation in Construction}
+96:398--410.
+
 \leavevmode\hypertarget{ref-Kong2018}{}%
@@
 Biology} 11:117--122.
+
+\leavevmode\hypertarget{ref-McCallen2019}{}%
+McCallen E., Knott J., Nunez-Mir G., Taylor B., Jo I., Fei S. 2019.
+Trends in ecology: Shifts in ecological research themes over the past
+four decades. \emph{Frontiers in Ecology and the Environment}
+17:109--116.
+
+\leavevmode\hypertarget{ref-Munoz2019}{}%
+Muñoz G., Kissling WD., Loon EE van. 2019. Biodiversity observations
+miner: A web application to unlock primary biodiversity data from
+published literature. \emph{Biodiversity Data Journal} 7.
 
@@
 
+\leavevmode\hypertarget{ref-Sinclair2016}{}%
+Sinclair L., Ijaz UZ., Jensen LJ., Coolen MJ., Gubry-Rangin C.,
+Chroňáková A., Oulas A., Pavloudi C., Schnetzer J., Weimann A., Ijaz A.,
+Eiler A., Quince C., Pafilis E. 2016. \texttt{Seqenv}: Linking sequences
+to environments through text mining. \emph{PeerJ} 4:e2690.
+
 \leavevmode\hypertarget{ref-Usai2018}{}%

### 2019-05-06 manuscript.Rmd
Typo

@@
 abstract: |
-  Text-mining is a powerful method for answering research questions. However,
+  Text mining is a powerful method for answering research questions. However,
   getting texts to extract information can be a daunting and complicated task.
-  The primary reason for this is publishers. There are thousands of different
-  publishers, each with their own licenses, URL patterns, access options, and
-  more. Layered on top of that is the varied access each user has based on
-  their institutional affiliation. Here, I introduce a suite of software packages
-  in the R programming language for fetching texts. The tapestry of different
-  publishers, access levels, and other factors requires a patchwork of approaches
-  for getting texts to users. The flagship R package called fulltext attempts
-  to simplify search and retrieval of texts for text-mining. The fulltext package,
-  along with many others, make acquiring texts easier than ever, facilitating
-  answering research questions with text-mining.
+  The primary reason for this is the diversity of publisher technologies.
+  There are thousands of different publishers, each with their own licenses,
+  URL patterns, access options, and more. Layered on top of that is the varied
+  access each user has based on their institutional affiliation. Here, I
+  introduce a suite of software packages in the R programming language for
+  fetching texts. The tapestry of different publishers, access levels, and
+  other factors requires a patchwork of approaches for getting texts to users.
+  The flagship R package called fulltext attempts to simplify search and
+  retrieval of texts for text mining by serving as an interface to the varied
+  and complex publishers. The fulltext package, along with many others, make
+  acquiring texts easier than ever, facilitating answering research questions
+  with text mining.
 bibliography: components/references.bib
@@
 
-There's more than 100 million research articles published
+There are more than 100 million research articles published
 (Crossref API: <https://github.com/CrossRef/rest-api-doc>), representing
@@
 
-Text-mining is the broad term associated with pulling information
-out of articles. Given the importance of text-mining, good
-text-mining tools are needed to make it easier for researchers to do.
+Text mining is the broad term associated with pulling information
+out of articles. Given the importance of text mining, good
+text mining tools are needed to make it easier for researchers to do.
 Graphical user interface (GUI) based text mining tools are available
@@
 
-Most recent text-mining papers do not use programmatic approaches, highlighting
-the need for more programmatic text-mining tools, and increased
+Using programmatic tools for text mining is important for a variety of reasons.
+First, using programmatic tools makes it easier for yourself. You can re-run code
+to get updated results as data changes, and results can always be re-created.
+Second, code can be shared whereas GUI driving workflows can not; colleagues,
+reviewers and other researchers can reproduce and check your work. (FIXME: 
+re-work this; citations)
+
+Most recent text mining papers do not use programmatic approaches, highlighting
+the need for more programmatic text mining tools, and increased
 discussion of those tools to increase awareness. For example, many papers
@@
 
-The R programming language is used widely throughout many academic fields;
-tools in R for text mining are of particular importance because they
-can be adopted by academics rapidly.
+The R programming language is free of cost, and is used widely throughout
+many academic fields; tools in R for text mining are of particular
+importance because they can be adopted by academics rapidly.
 
-Here, I present an overview of text-mining tools in the R programming
-language, not for text-mining analysis, but rather those
+Here, I present an overview of text mining tools in the R programming
+language, not for text mining analysis, but rather those
 tools for searching for, acquiring, and extracting parts of texts
@@
 Document Format (PDF)[^pdf]. These can be broken out into two groups: text based
-PDFs and scanned PDFs. The former are converted from digital versions of
-various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
+PDFs and scanned PDFs (images of text). The former are converted from digital
+versions of various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
 latter are created by scanning print articles to a PDF format. Text-based
-PDFs are much better for text-mining purposes as plain text can be extracted
+PDFs are much better for text mining purposes as plain text can be extracted
 easily in R with [pdftools][], a binding to [libpoppler][]. However, with
@@
 most publishers most likely have the XML behind each of their articles,
-they for some indefensible reason do not share it - making text-mining more
+they for some indefensible reason do not share it - making text mining more
 difficult. Some provide plain text (Elsevier). I only know of one publisher
-that provides full text as JSON (PLOS). Thus, text-mining, in most cases,
+that provides full text as JSON (PLOS). Thus, text mining, in most cases,
 will require extracting text from PDFs.
@@
 
-Acces to full-text is the holy grail in text-mining. Some use cases can
+Acces to full-text is the holy grail in text mining. Some use cases can
 get by with article metadata (authors, title, etc.), some with abstracts,
@@
 Sometimes a VPN is required, and this can provide a significant technical
-hurdle to users attempting to do text-mining work.
+hurdle to users attempting to do text mining work.
 
-One final hurdle in text-mining comes unsurprisingly from Elsevier. They
+One final hurdle in text mining comes unsurprisingly from Elsevier. They
 use so-called "fences" for programmatic access. That is, even if a person
@@
 
-A text-mining project starts with a question. From that question, researchers
-then attempt to acquire scholarly articles for text-mining. Finding appropriate
+A text mining project starts with a question. From that question, researchers
+then attempt to acquire scholarly articles for text mining. Finding appropriate
 articles is not altogether straight-forward.
@@
 sources differ in how search works from the user perspective. For example, some
-provide wild card search and some do not. Some sources are searching full text
+provide wild card/fuzzy search (i.e., 'appl\*' includes results for 'apple',
+'application', etc.) and some do not. Some sources are searching full text
 of articles, while others only search metadata (i.e., title, authors, abstract).
@@
 In brief, the same search against different sources produces different results.
-Some text-mining research articles perform the same search against many
+Some text mining research articles perform the same search against many
 different sources (refs), while others choose just one source.
@@
 online. However, only a small proportion of scientific journals provide
-access to their full content; whereas, most publishers provide open
+access to their full text; whereas, most publishers provide open
 access to their metadata only (most often through Crossref; Table 1).
@@
 Data Provider                      | Content Type                 | rOpenSci Package   | Documentation
----------------------------------- | ---------------------   | ------------------------ | -------------
+---------------------------------- | ---------------------------- | ------------------ | -------------
 Crossref                           | Metadata                     | rcrossref/crminer  | [^1]
 DataCite                           | Metadata                     | rdatacite          | [^2]
-Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
-Public Library of Science (PLoS)   | Full content/altmetrics | rplos                    | [^4]
-Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
-arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
-Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
-bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
-PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
-Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
+Biodiversity Heritage Library      | Full text/Metadata           | rbhl               | [^3]
+Public Library of Science (PLoS)   | Full text (pdf/xml)/Metadata | rplos              | [^4]
+Scopus (Elsevier)                  | Full text (pdf/xml)/Metadata | fulltext           | [^5]
+arXiv                              | Full text (pdf)/Metadata     | aRxiv              | [^6]
+Biomed Central (via Springer)      | Full text (pdf)/Metadata     | fulltext           | [^7]
+bioRxiv                            | Full text (pdf)/Metadata     | fulltext           | [^8]
+PMC/Pubmed (via Entrez)            | Full text (pdf/xml)/Metadata | rentrez            | [^9]
+Europe PMC                         | Full text (pdf/xml)/Metadata | europepmc          | [^10]
 Microsoft Academic Search          | Metadata                     | microdemic         | [^11]
 Directory of Open Access Journals  | Metadata                     | jaod               | [^12]
-JSTOR Data for Research            | Full content            | jstor                    | [^13]
+JSTOR Data for Research            | Full text                    | jstor              | [^13]
 ORCID                              | Metadata                     | rorcid             | [^14]
@@
 SHERPA/RoMEO                       | Journal Level Metadata       | rromeo             | [^18]
-CORE                               | Full content/Metadata   | rcoreoa                  | [^19]
+CORE                               | Full text (pdf)/Metadata     | rcoreoa            | [^19]
 Dissemin                           | Metadata                     | dissemr            | [^20]
@@
 
-## Crossref/Datacite
+## Crossref and Datacite
 
@@
 One can search by DOI or search citation data to get citations. In addition, Crossref has a
-text-mining opt-in program for publishers. The result of this is that some
+text mining opt-in program for publishers. The result of this is that some
 publishers provide URLs for full text content of their articles. The majority of these links
@@
 
-The main interfaces for Crossref in R are [rcrossref][] and [crminer][]. Similar interfaces
+The main interfaces for Crossref in R are [rcrossref][] and [crminer][]. rcrossref is a
+complete client for the public facing Crossref web services including metadata, whereas crminer
+focuses only on retrieving full text of articles. Similar interfaces to rcrossref
 are available in Ruby ([serrano][]) and Python ([habanero][]).
@@
 
-Elsevier is one of the largest publishers. Most of their articles are not open access. However,
-they have a numbrer of advantages if you have access to their articles: they are one of the few
+Elsevier is one of the largest publishers. Scopus is one of their products, a searchable
+database of articles. Most of their articles are not open access. However,
+they have a number of advantages if you have access to their articles: they are one of the few
 publishers to provide machine readable XML (many publishers do have XML versions of articles,
-but do not provide it); they are one of the few (two) publishers part of Crossref's text
-and data mining program. The packages [fulltext][] and [crminer][] can be used to access
-Elsevier articles through Crossref's TDM program. There's an interface to Scopus article
-search within [fulltext][].
+but do not provide it, as mentioned in the Introduction); they are one of the few (two)
+publishers that take part in Crossref's text and data mining program. The packages
+[fulltext][] and [crminer][] can be used to access Elsevier articles through Crossref's
+TDM program. There's an interface to Scopus article search within [fulltext][].
 
@@
 arXiv and bioRxiv are preprint publishers, the former in existence for many years, and the
-latter new on the scene. You can access articles from these publishers through [fulltext][].
+latter new on the scene. Preprints are scholarly articles that are generally not
+peer-reviewed, but that for the most part will later be published in a different peer-review
+publication. You can access articles from these publishers through [fulltext][].
 arXiv does provide a web API that we hook into; bioRxiv does not, but we can get you
@@
 "versioned, publicly-editable catalog of research publications:  journal articles,
-conference proceedings, pre-prints, blog posts". Fatcat is currently does not have an
+conference proceedings, preprints, blog posts". Fatcat is currently does not have an
 R client, but is used inside of the [fulltext][] package.
@@
 
+## Learn more
+
+To learn how to use the above packages, follow the links to get to documentation,
+and how to report issues. The rOpenSci forum (<https://discuss.ropensci.org/>) is
+a good place to ask questions about these packages.
+
 # fulltext: a toolset for text mining in R
 
-[fulltext][] is a general purpose R package for the data part of text-mining:
+[fulltext][] is a general purpose R package for the data part of text mining:
 search for articles, get links to articles, get article abstracts, and fetch
@@
 subsequent requests for the same article just use the cached version. Additionally,
-all errors in `ft_get()` are collected in a tidy data.frame in the output of the
+all errors in `ft_get()` are collected in a data.frame in the output of the
 function to help the user quickly determine what went wrong.
 
-# How to text mine from R: Three case studies
+# Text mining in R: Use Cases
 
-## Case study 1: Citation mining
+## Use cases in the literature
+
+At least three papers have used the fulltext package. @Bauer2016 explored
+a way to think about the qualitative impact of citations rather than simple
+counts of citations; they did not use fulltext, but pointed it out as a good
+option for collecting papers. @Walsh2018 used fulltext and [roadoi][], an
+rOpenSci R package, to "map the literature" on nutritional interventions
+in cognitive health. @Piper2019 in a review of DNA metabarcoding for
+high-throughput insect surveillance used fulltext, along with [rentrez][]
+(an rOpenSci package), to collect articles for text mining.
+
+## Three case studies
+
+The following three case studies demonstrate in brief how one could use
+fulltext and related packages to do a few different text mining tasks.
+
+### Case study 1: Citation mining
 
@@
 
-## Case study 2: Abstract mining
+### Case study 2: Abstract mining
 
@@
 
-## Case study 3: Full text mining
+### Case study 3: Full text mining
 
@@
 
-Text-mining will always be a complex task given all the layers involved:
+Text mining will always be a complex task given all the layers involved:
 often temporal time-span of research questions; varied permissions among
@@
 
-Programmatic text-mining is a first step towards making text-mining easier.
-The R ecosystem is an especially good place to do text-mining because 
-there are many packages for text-mining analysis, and endless packages for
+Programmatic text mining is a first step towards making text mining easier.
+The R ecosystem is an especially good place to do text mining because
+there are many packages for text mining analysis, and endless packages for
 any required statistical analyses. In addition, rOpenSci and others are
@@
 
-With respect to what publishers can do to make text-mining easier, publishers
+With respect to what publishers can do to make text mining easier, publishers
 should:
@@
 
-xxxx
+This manuscript was greatly improved by comments from Maëlle Salmon.
 
@@
 <https://github.com/ropensci/textmine>. Software can be found at
-<https://github.com/ropensci/xxx>, xxxx, all under MIT licenses.
+<https://github.com/ropensci>, all under MIT licenses.
 
@@
 [rcoreoa]: https://github.com/ropensci/rcoreoa
+[roadoi]: https://github.com/ropensci/roadoi

### 2019-08-23 manuscript.Rmd
Typo
fix #3 text-mining to text minig

--- 2019-04-16 manuscript.Rmd
@@
 abstract: |
-  Text-mining is a powerful method for answering research questions. However,
+  Text mining is a powerful method for answering research questions. However,
   getting texts to extract information can be a daunting and complicated task.
@@
   for getting texts to users. The flagship R package called fulltext attempts
-  to simplify search and retrieval of texts for text-mining. The fulltext package,
+  to simplify search and retrieval of texts for text mining. The fulltext package,
   along with many others, make acquiring texts easier than ever, facilitating
-  answering research questions with text-mining.
+  answering research questions with text mining.
 bibliography: components/references.bib
@@
 
-Most recent text-mining papers do not use programmatic approaches, highlighting
-the need for more programmatic text-mining tools, and increased
+Most recent text mining papers do not use programmatic approaches, highlighting
+the need for more programmatic text mining tools, and increased
 discussion of those tools to increase awareness. For example, many papers 
@@
 
-Here, I present an overview of text-mining tools in the R programming
-language, not for text-mining analysis, but rather those
+Here, I present an overview of text mining tools in the R programming
+language, not for text mining analysis, but rather those
 tools for searching for, acquiring, and extracting parts of texts
@@
 latter are created by scanning print articles to a PDF format. Text-based
-PDFs are much better for text-mining purposes as plain text can be extracted
+PDFs are much better for text mining purposes as plain text can be extracted
 easily in R with [pdftools][], a binding to [libpoppler][]. However, with
@@
 
-[fulltext][] is a general purpose R package for the data part of text-mining:
+[fulltext][] is a general purpose R package for the data part of text mining:
 search for articles, get links to articles, get article abstracts, and fetch
@@
 
-Text-mining will always be a complex task given all the layers involved:
+Text mining will always be a complex task given all the layers involved:
 often temporal time-span of research questions; varied permissions among 
@@
 
-Programmatic text-mining is a first step towards making text-mining easier.
-The R ecosystem is an especially good place to do text-mining because 
-there are many packages for text-mining analysis, and endless packages for
+Programmatic text mining is a first step towards making text mining easier.
+The R ecosystem is an especially good place to do text mining because 
+there are many packages for text mining analysis, and endless packages for
 any required statistical analyses. In addition, rOpenSci and others are 

### 2019-08-23 manuscript.tex
fix #3 text-mining to text minig

@@
   \begin{abstract}
-  Text-mining is a powerful method for answering research questions.
+  Text mining is a powerful method for answering research questions.
   However, getting texts to extract information can be a daunting and
@@
   users. The flagship R package called fulltext attempts to simplify
-  search and retrieval of texts for text-mining. The fulltext package,
+  search and retrieval of texts for text mining. The fulltext package,
   along with many others, make acquiring texts easier than ever,
-  facilitating answering research questions with text-mining.
+  facilitating answering research questions with text mining.
   \end{abstract}
@@
 
-Text-mining is the broad term associated with pulling information out of
-articles. Given the importance of text-mining, good text-mining tools
+Text mining is the broad term associated with pulling information out of
+articles. Given the importance of text mining, good text mining tools
 are needed to make it easier for researchers to do. Graphical user
@@
 
-Most recent text-mining papers do not use programmatic approaches,
-highlighting the need for more programmatic text-mining tools, and
+Most recent text mining papers do not use programmatic approaches,
+highlighting the need for more programmatic text mining tools, and
 increased discussion of those tools to increase awareness. For example,
@@
 
-Here, I present an overview of text-mining tools in the R programming
-language, not for text-mining analysis, but rather those tools for
+Here, I present an overview of text mining tools in the R programming
+language, not for text mining analysis, but rather those tools for
 searching for, acquiring, and extracting parts of texts (e.g., title,
@@
 created by scanning print articles to a PDF format. Text-based PDFs are
-much better for text-mining purposes as plain text can be extracted
+much better for text mining purposes as plain text can be extracted
 easily in R with \href{https://github.com/ropensci/pdftools}{pdftools},
@@
 their articles, they for some indefensible reason do not share it -
-making text-mining more difficult. Some provide plain text (Elsevier). I
+making text mining more difficult. Some provide plain text (Elsevier). I
 only know of one publisher that provides full text as JSON (PLOS). Thus,
-text-mining, in most cases, will require extracting text from PDFs.
+text mining, in most cases, will require extracting text from PDFs.
 
@@
 
-Acces to full-text is the holy grail in text-mining. Some use cases can
+Acces to full-text is the holy grail in text mining. Some use cases can
 get by with article metadata (authors, title, etc.), some with
@@
 differ. Sometimes a VPN is required, and this can provide a significant
-technical hurdle to users attempting to do text-mining work.
-
-One final hurdle in text-mining comes unsurprisingly from Elsevier. They
+technical hurdle to users attempting to do text mining work.
+
+One final hurdle in text mining comes unsurprisingly from Elsevier. They
 use so-called ``fences'' for programmatic access. That is, even if a
@@
 
-A text-mining project starts with a question. From that question,
-researchers then attempt to acquire scholarly articles for text-mining.
+A text mining project starts with a question. From that question,
+researchers then attempt to acquire scholarly articles for text mining.
 Finding appropriate articles is not altogether straight-forward.
@@
 the same search against different sources produces different results.
-Some text-mining research articles perform the same search against many
+Some text mining research articles perform the same search against many
 different sources (refs), while others choose just one source.
@@
 One can search by DOI or search citation data to get citations. In
-addition, Crossref has a text-mining opt-in program for publishers. The
+addition, Crossref has a text mining opt-in program for publishers. The
 result of this is that some publishers provide URLs for full text
@@
 \href{https://github.com/ropensci/fulltext}{fulltext} is a general
-purpose R package for the data part of text-mining: search for articles,
+purpose R package for the data part of text mining: search for articles,
 get links to articles, get article abstracts, and fetch full text of
@@
 
-Programmatic text-mining is a first step towards making text-mining
-easier. The R ecosystem is an especially good place to do text-mining
-because there are many packages for text-mining analysis, and endless
+Programmatic text mining is a first step towards making text mining
+easier. The R ecosystem is an especially good place to do text mining
+because there are many packages for text mining analysis, and endless
 packages for any required statistical analyses. In addition, rOpenSci
@@
 
-With respect to what publishers can do to make text-mining easier,
+With respect to what publishers can do to make text mining easier,
 publishers should:

### 2019-08-23 manuscript.Rmd
fix #4 abstract fix

@@
   getting texts to extract information can be a daunting and complicated task.
-  The primary reason for this is publishers. There are thousands of different
-  publishers, each with their own licenses, URL patterns, access options, and
-  more. Layered on top of that is the varied access each user has based on
-  their institutional affiliation. Here, I introduce a suite of software packages
-  in the R programming language for fetching texts. The tapestry of different
-  publishers, access levels, and other factors requires a patchwork of approaches
-  for getting texts to users. The flagship R package called fulltext attempts
-  to simplify search and retrieval of texts for text mining. The fulltext package,
-  along with many others, make acquiring texts easier than ever, facilitating
-  answering research questions with text mining.
+  The primary reason for this is the diversity of publisher technologies. 
+  There are thousands of different publishers, each with their own licenses,
+  URL patterns, access options, and more. Layered on top of that is the varied
+  access each user has based on their institutional affiliation. Here, I
+  introduce a suite of software packages in the R programming language for
+  fetching texts. The tapestry of different publishers, access levels, and
+  other factors requires a patchwork of approaches for getting texts to users.
+  The flagship R package called fulltext attempts to simplify search and
+  retrieval of texts for text mining by serving as an interface to the varied
+  and complex publishers. The fulltext package, along with many others, make
+  acquiring texts easier than ever, facilitating answering research questions
+  with text mining.
 bibliography: components/references.bib

### 2019-08-23 manuscript.tex
fix #4 abstract fix

@@
   However, getting texts to extract information can be a daunting and
-  complicated task. The primary reason for this is publishers. There are
-  thousands of different publishers, each with their own licenses, URL
-  patterns, access options, and more. Layered on top of that is the varied
-  access each user has based on their institutional affiliation. Here, I
-  introduce a suite of software packages in the R programming language for
-  fetching texts. The tapestry of different publishers, access levels, and
-  other factors requires a patchwork of approaches for getting texts to
-  users. The flagship R package called fulltext attempts to simplify
-  search and retrieval of texts for text mining. The fulltext package,
-  along with many others, make acquiring texts easier than ever,
-  facilitating answering research questions with text mining.
+  complicated task. The primary reason for this is the diversity of
+  publisher technologies. There are thousands of different publishers,
+  each with their own licenses, URL patterns, access options, and more.
+  Layered on top of that is the varied access each user has based on their
+  institutional affiliation. Here, I introduce a suite of software
+  packages in the R programming language for fetching texts. The tapestry
+  of different publishers, access levels, and other factors requires a
+  patchwork of approaches for getting texts to users. The flagship R
+  package called fulltext attempts to simplify search and retrieval of
+  texts for text mining by serving as an interface to the varied and
+  complex publishers. The fulltext package, along with many others, make
+  acquiring texts easier than ever, facilitating answering research
+  questions with text mining.
   \end{abstract}

### 2019-08-23 manuscript.Rmd
intro change

@@
 
-The R programming language is used widely throughout many academic fields;
-tools in R for text mining are of particular importance because they
-can be adopted by academics rapidly.
+The R programming language is free of cost, and is used widely throughout
+many academic fields; tools in R for text mining are of particular
+importance because they can be adopted by academics rapidly.
 

### 2019-08-23 manuscript.tex
intro change

@@
 
-The R programming language is used widely throughout many academic
-fields; tools in R for text mining are of particular importance because
-they can be adopted by academics rapidly.
+The R programming language is free of cost, and is used widely
+throughout many academic fields; tools in R for text mining are of
+particular importance because they can be adopted by academics rapidly.
 

### 2019-08-23 manuscript.Rmd
fix #7 clarify scanned pdf; fix #8 clarify what wild card search is

@@
 Document Format (PDF)[^pdf]. These can be broken out into two groups: text based
-PDFs and scanned PDFs. The former are converted from digital versions of
-various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
+PDFs and scanned PDFs (images of text). The former are converted from digital
+versions of various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
 latter are created by scanning print articles to a PDF format. Text-based
@@
 sources differ in how search works from the user perspective. For example, some 
-provide wild card search and some do not. Some sources are searching full text
+provide wild card/fuzzy search (i.e., 'appl\*' includes results for 'apple',
+'application', etc.) and some do not. Some sources are searching full text
 of articles, while others only search metadata (i.e., title, authors, abstract).

### 2019-08-23 manuscript.tex
fix #7 clarify scanned pdf; fix #8 clarify what wild card search is

@@
 These can be broken out into two groups: text based PDFs and scanned
-PDFs. The former are converted from digital versions of various kinds
-(MS Word, OpenOffice, LaTeX, markdown, etc.), while the latter are
-created by scanning print articles to a PDF format. Text-based PDFs are
-much better for text mining purposes as plain text can be extracted
-easily in R with \href{https://github.com/ropensci/pdftools}{pdftools},
-a binding to \href{https://poppler.freedesktop.org/}{libpoppler}.
-However, with scanned PDFs, text must be extracted using Optimal
-Character Recognition (OCR; see R package
+PDFs (images of text). The former are converted from digital versions of
+various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
+latter are created by scanning print articles to a PDF format.
+Text-based PDFs are much better for text mining purposes as plain text
+can be extracted easily in R with
+\href{https://github.com/ropensci/pdftools}{pdftools}, a binding to
+\href{https://poppler.freedesktop.org/}{libpoppler}. However, with
+scanned PDFs, text must be extracted using Optimal Character Recognition
+(OCR; see R package
 \href{https://github.com/ropensci/tesseract}{tesseract}), which isn't
@@
 knowledge. Nonetheless, data sources differ in how search works from the
-user perspective. For example, some provide wild card search and some do
-not. Some sources are searching full text of articles, while others only
-search metadata (i.e., title, authors, abstract). In addition, each
-source has a different set of metadata/full text available. In brief,
-the same search against different sources produces different results.
-Some text mining research articles perform the same search against many
-different sources (refs), while others choose just one source.
+user perspective. For example, some provide wild card/fuzzy search
+(i.e., `appl*' includes results for `apple', `application', etc.) and
+some do not. Some sources are searching full text of articles, while
+others only search metadata (i.e., title, authors, abstract). In
+addition, each source has a different set of metadata/full text
+available. In brief, the same search against different sources produces
+different results. Some text mining research articles perform the same
+search against many different sources (refs), while others choose just
+one source.
 

### 2019-08-23 manuscript.Rmd
fix #9 change full content to full text

@@
 DataCite                           | Metadata                | rdatacite                | [^2]
-Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
-Public Library of Science (PLoS)   | Full content/altmetrics | rplos                    | [^4]
-Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
-arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
-Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
-bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
-PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
-Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
+Biodiversity Heritage Library      | Full text/Metadata      | rbhl                     | [^3]
+Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
+Scopus (Elsevier)                  | Full text/Metadata      | fulltext                 | [^5]
+arXiv                              | Full text/Metadata      | aRxiv                    | [^6]
+Biomed Central (via Springer)      | Full text/Metadata      | fulltext                 | [^7]
+bioRxiv                            | Full text/Metadata      | fulltext                 | [^8]
+PMC/Pubmed (via Entrez)            | Full text/Metadata      | rentrez                  | [^9]
+Europe PMC                         | Full text/Metadata      | europepmc                | [^10]
 Microsoft Academic Search          | Metadata                | microdemic               | [^11]
 Directory of Open Access Journals  | Metadata                | jaod                     | [^12]
-JSTOR Data for Research            | Full content            | jstor                    | [^13]
+JSTOR Data for Research            | Full text               | jstor                    | [^13]
 ORCID                              | Metadata                | rorcid                   | [^14]
@@
 SHERPA/RoMEO                       | Journal Level Metadata  | rromeo                   | [^18]
-CORE                               | Full content/Metadata   | rcoreoa                  | [^19]
+CORE                               | Full text/Metadata      | rcoreoa                  | [^19]
 Dissemin                           | Metadata                | dissemr                  | [^20]

### 2019-08-23 manuscript.tex
fix #9 change full content to full text

@@
 online. However, only a small proportion of scientific journals provide
-access to their full content; whereas, most publishers provide open
-access to their metadata only (most often through Crossref; Table 1).
-The following is a synopsis of the major data sources and associated R
+access to their full text; whereas, most publishers provide open access
+to their metadata only (most often through Crossref; Table 1). The
+following is a synopsis of the major data sources and associated R
 tools.
@@
 DataCite & Metadata & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
-Biodiversity Heritage Library & Full content/Metadata & rbhl &
-\footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
-Public Library of Science (PLoS) & Full content/altmetrics & rplos &
+Biodiversity Heritage Library & Full text/Metadata & rbhl & \footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
+Public Library of Science (PLoS) & Full text/altmetrics & rplos &
 \footnote{\url{http://api.plos.org/solr}}\tabularnewline
-Scopus (Elsevier) & Full content/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
-arXiv & Full content/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
-Biomed Central (via Springer) & Full content/Metadata & fulltext &
+Scopus (Elsevier) & Full text/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
+arXiv & Full text/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
+Biomed Central (via Springer) & Full text/Metadata & fulltext &
 \footnote{\url{https://dev.springer.com/}}\tabularnewline
-bioRxiv & Full content/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
-PMC/Pubmed (via Entrez) & Full content/Metadata & rentrez & \footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
-Europe PMC & Full content/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
+bioRxiv & Full text/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
+PMC/Pubmed (via Entrez) & Full text/Metadata & rentrez & \footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
+Europe PMC & Full text/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
 Microsoft Academic Search & Metadata & microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
 Directory of Open Access Journals & Metadata & jaod & \footnote{\url{https://doaj.org/api/v1/docs}}\tabularnewline
-JSTOR Data for Research & Full content & jstor & \footnote{\url{https://www.jstor.org/dfr/}}\tabularnewline
+JSTOR Data for Research & Full text & jstor & \footnote{\url{https://www.jstor.org/dfr/}}\tabularnewline
 ORCID & Metadata & rorcid & \footnote{\url{https://pub.orcid.org/}}\tabularnewline
@@
 SHERPA/RoMEO & Journal Level Metadata & rromeo & \footnote{\url{http://www.sherpa.ac.uk/romeo/apimanual.php?la=en\&fIDnum=\%7C\&mode=simple}}\tabularnewline
-CORE & Full content/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
+CORE & Full text/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
 Dissemin & Metadata & dissemr & \footnote{\url{https://dissemin.readthedocs.io/en/latest/api.html}}\tabularnewline

### 2019-08-23 manuscript.tex
fix #11 clarify the various crossref clients

@@
 
-\hypertarget{crossrefdatacite}{%
-\subsection{Crossref/Datacite}\label{crossrefdatacite}}
+\hypertarget{crossref-and-datacite}{%
+\subsection{Crossref and Datacite}\label{crossref-and-datacite}}
 
@@
 \href{https://github.com/ropensci/rcrossref}{rcrossref} and
-\href{https://github.com/ropensci/crminer}{crminer}. Similar interfaces
-are available in Ruby
+\href{https://github.com/ropensci/crminer}{crminer}. rcrossref is a
+complete client for the public facing Crossref web services including
+metadata, whereas crminer focuses only on retrieving full text of
+articles. Similar interfaces to rcrossref are available in Ruby
 (\href{https://github.com/sckott/serrano}{serrano}) and Python

### 2019-08-23 manuscript.Rmd
acknowledgements maelle
fix #14 add what type of fulltext avail for publishers in table 1

@@
 Data Provider                      | Content Type                 | rOpenSci Package   | Documentation
----------------------------------- | ---------------------   | ------------------------ | -------------
+---------------------------------- | ---------------------------- | ------------------ | -------------
 Crossref                           | Metadata                     | rcrossref/crminer  | [^1]
@@
 Biodiversity Heritage Library      | Full text/Metadata           | rbhl               | [^3]
-Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
-Scopus (Elsevier)                  | Full text/Metadata      | fulltext                 | [^5]
-arXiv                              | Full text/Metadata      | aRxiv                    | [^6]
-Biomed Central (via Springer)      | Full text/Metadata      | fulltext                 | [^7]
-bioRxiv                            | Full text/Metadata      | fulltext                 | [^8]
-PMC/Pubmed (via Entrez)            | Full text/Metadata      | rentrez                  | [^9]
-Europe PMC                         | Full text/Metadata      | europepmc                | [^10]
+Public Library of Science (PLoS)   | Full text (pdf/xml)/Metadata | rplos              | [^4]
+Scopus (Elsevier)                  | Full text (pdf/xml)/Metadata | fulltext           | [^5]
+arXiv                              | Full text (pdf)/Metadata     | aRxiv              | [^6]
+Biomed Central (via Springer)      | Full text (pdf)/Metadata     | fulltext           | [^7]
+bioRxiv                            | Full text (pdf)/Metadata     | fulltext           | [^8]
+PMC/Pubmed (via Entrez)            | Full text (pdf/xml)/Metadata | rentrez            | [^9]
+Europe PMC                         | Full text (pdf/xml)/Metadata | europepmc          | [^10]
 Microsoft Academic Search          | Metadata                     | microdemic         | [^11]
@@
 SHERPA/RoMEO                       | Journal Level Metadata       | rromeo             | [^18]
-CORE                               | Full text/Metadata      | rcoreoa                  | [^19]
+CORE                               | Full text (pdf)/Metadata     | rcoreoa            | [^19]
 Dissemin                           | Metadata                     | dissemr            | [^20]

### 2019-08-23 manuscript.tex
fix #14 add what type of fulltext avail for publishers in table 1

@@
 Biodiversity Heritage Library & Full text/Metadata & rbhl & \footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
-Public Library of Science (PLoS) & Full text/altmetrics & rplos &
-\footnote{\url{http://api.plos.org/solr}}\tabularnewline
-Scopus (Elsevier) & Full text/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
-arXiv & Full text/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
-Biomed Central (via Springer) & Full text/Metadata & fulltext &
+Public Library of Science (PLoS) & Full text (pdf/xml)/Metadata & rplos
+& \footnote{\url{http://api.plos.org/solr}}\tabularnewline
+Scopus (Elsevier) & Full text (pdf/xml)/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
+arXiv & Full text (pdf)/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
+Biomed Central (via Springer) & Full text (pdf)/Metadata & fulltext &
 \footnote{\url{https://dev.springer.com/}}\tabularnewline
-bioRxiv & Full text/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
-PMC/Pubmed (via Entrez) & Full text/Metadata & rentrez & \footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
-Europe PMC & Full text/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
+bioRxiv & Full text (pdf)/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
+PMC/Pubmed (via Entrez) & Full text (pdf/xml)/Metadata & rentrez &
+\footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
+Europe PMC & Full text (pdf/xml)/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
 Microsoft Academic Search & Metadata & microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
@@
 SHERPA/RoMEO & Journal Level Metadata & rromeo & \footnote{\url{http://www.sherpa.ac.uk/romeo/apimanual.php?la=en\&fIDnum=\%7C\&mode=simple}}\tabularnewline
-CORE & Full text/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
+CORE & Full text (pdf)/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
 Dissemin & Metadata & dissemr & \footnote{\url{https://dissemin.readthedocs.io/en/latest/api.html}}\tabularnewline
@@
 
-Elsevier is one of the largest publishers. Most of their articles are
-not open access. However, they have a numbrer of advantages if you have
+Elsevier is one of the largest publishers. Scopus is one of their
+products, a searchable database of articles. Most of their articles are
+not open access. However, they have a number of advantages if you have
 access to their articles: they are one of the few publishers to provide
 machine readable XML (many publishers do have XML versions of articles,
-but do not provide it); they are one of the few (two) publishers part of
-Crossref's text and data mining program. The packages
+but do not provide it, as mentioned in the Introduction); they are one
+of the few (two) publishers that take part in Crossref's text and data
+mining program. The packages
 \href{https://github.com/ropensci/fulltext}{fulltext} and
@@
 arXiv and bioRxiv are preprint publishers, the former in existence for
-many years, and the latter new on the scene. You can access articles
-from these publishers through
+many years, and the latter new on the scene. Preprints are scholarly
+articles that are generally not peer-reviewed, but that for the most
+part will later be published in a different peer-review publication. You
+can access articles from these publishers through
 \href{https://github.com/ropensci/fulltext}{fulltext}. arXiv does
@@
 a ``versioned, publicly-editable catalog of research publications:
-journal articles, conference proceedings, pre-prints, blog posts''.
+journal articles, conference proceedings, preprints, blog posts''.
 Fatcat is currently does not have an R client, but is used inside of the
@@
 \url{https://github.com/ropensci/textmine}. Software can be found at
-\url{https://github.com/ropensci/xxx}, xxxx, all under MIT licenses.
+\url{https://github.com/ropensci}, all under MIT licenses.
 

### 2019-08-23 manuscript.Rmd
fix #17 dont need tidy actually

@@
 subsequent requests for the same article just use the cached version. Additionally,
-all errors in `ft_get()` are collected in a tidy data.frame in the output of the
+all errors in `ft_get()` are collected in a data.frame in the output of the
 function to help the user quickly determine what went wrong.

### 2019-08-23 manuscript.Rmd
fix #19 text in introduction about why programmatic is important

@@
 linking sequences to environments through text mining.
+
+Using programmatic tools for text mining is important for a variety of reasons.
+First, using programmatic tools makes it easier for yourself. You can re-run code
+to get updated results as data changes, and results can always be re-created.
+Second, code can be shared whereas GUI driving workflows can not; colleagues,
+reviewers and other researchers can reproduce and check your work. (FIXME: 
+re-work this; citations)
 

### 2019-08-23 manuscript.tex
fix #19 text in introduction about why programmatic is important

@@
 
+Using programmatic tools for text mining is important for a variety of
+reasons. First, using programmatic tools makes it easier for yourself.
+You can re-run code to get updated results as data changes, and results
+can always be re-created. Second, code can be shared whereas GUI driving
+workflows can not; colleagues, reviewers and other researchers can
+reproduce and check your work. (FIXME: re-work this; citations)
+
 Most recent text mining papers do not use programmatic approaches,
@@
 the cached version. Additionally, all errors in \texttt{ft\_get()} are
-collected in a tidy data.frame in the output of the function to help the
-user quickly determine what went wrong.
+collected in a data.frame in the output of the function to help the user
+quickly determine what went wrong.
 

### 2019-08-23 manuscript.tex
fix #20 move case studies further up

@@
 
-\hypertarget{data-sources}{%
-\section{Data sources}\label{data-sources}}
-
-There is increasing open access scientific literature content available
-online. However, only a small proportion of scientific journals provide
-access to their full text; whereas, most publishers provide open access
-to their metadata only (most often through Crossref; Table 1). The
-following is a synopsis of the major data sources and associated R
-tools.
-
-\newpage
-
-Table 1. Sources of scientific literature, their content type provided
-via web services, whether rOpenSci has an R packages for the service,
-and where to find the API documentation.
-
-\begin{longtable}[]{@{}llll@{}}
-\toprule
-Data Provider & Content Type & rOpenSci Package &
-Documentation\tabularnewline
-\midrule
-\endhead
-Crossref & Metadata & rcrossref/crminer & \footnote{\url{https://api.crossref.org}}\tabularnewline
-DataCite & Metadata & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
-Biodiversity Heritage Library & Full text/Metadata & rbhl & \footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
-Public Library of Science (PLoS) & Full text (pdf/xml)/Metadata & rplos
-& \footnote{\url{http://api.plos.org/solr}}\tabularnewline
-Scopus (Elsevier) & Full text (pdf/xml)/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
-arXiv & Full text (pdf)/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
-Biomed Central (via Springer) & Full text (pdf)/Metadata & fulltext &
-\footnote{\url{https://dev.springer.com/}}\tabularnewline
-bioRxiv & Full text (pdf)/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
-PMC/Pubmed (via Entrez) & Full text (pdf/xml)/Metadata & rentrez &
-\footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
-Europe PMC & Full text (pdf/xml)/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
-Microsoft Academic Search & Metadata & microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
-Directory of Open Access Journals & Metadata & jaod & \footnote{\url{https://doaj.org/api/v1/docs}}\tabularnewline
-JSTOR Data for Research & Full text & jstor & \footnote{\url{https://www.jstor.org/dfr/}}\tabularnewline
-ORCID & Metadata & rorcid & \footnote{\url{https://pub.orcid.org/}}\tabularnewline
-Wikimedia's Citoid & Citations & rcitoid & \footnote{\url{https://en.wikipedia.org/api/rest_v1/\#/Citation/getCitation}}\tabularnewline
-Open Citation Corpus & Citations & citecorp & \footnote{\url{http://opencitations.net/}}\tabularnewline
-Fatcat & Metadata & none & \footnote{\url{https://fatcat.wiki/}}\tabularnewline
-SHERPA/RoMEO & Journal Level Metadata & rromeo & \footnote{\url{http://www.sherpa.ac.uk/romeo/apimanual.php?la=en\&fIDnum=\%7C\&mode=simple}}\tabularnewline
-CORE & Full text (pdf)/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
-Dissemin & Metadata & dissemr & \footnote{\url{https://dissemin.readthedocs.io/en/latest/api.html}}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\hypertarget{crossref-and-datacite}{%
-\subsection{Crossref and Datacite}\label{crossref-and-datacite}}
-
-Crossref is a non-profit that creates (or ``mints'') Digital Object
-Identifiers (DOIs). In addition, they maintain metadata associated with
-each DOI. The metadata ranges from simple (including author, title,
-dates, DOI, type, publisher) to including number of citations to the
-article, as well as references in the article, and even abstracts. At
-the time of writing they hold 100 million DOIs.
-
-One can search by DOI or search citation data to get citations. In
-addition, Crossref has a text mining opt-in program for publishers. The
-result of this is that some publishers provide URLs for full text
-content of their articles. The majority of these links are pay-walled,
-while some are open access. Using any of the various tools for working
-with Crossref data, you can filter your search to get only articles with
-full text links, and further to get only articles with full text links
-that are open access.
-
-The main interfaces for Crossref in R are
-\href{https://github.com/ropensci/rcrossref}{rcrossref} and
-\href{https://github.com/ropensci/crminer}{crminer}. rcrossref is a
-complete client for the public facing Crossref web services including
-metadata, whereas crminer focuses only on retrieving full text of
-articles. Similar interfaces to rcrossref are available in Ruby
-(\href{https://github.com/sckott/serrano}{serrano}) and Python
-(\href{https://github.com/sckott/habanero}{habanero}).
-
-Datacite is similar to Crossref, but focuses on datasets instead of
-articles. The main interface for Datacite in R is
-\href{https://github.com/ropensci/rdatacite}{rdatacite}.
-
-\hypertarget{biodiversity-heritage-library}{%
-\subsection{Biodiversity Heritage
-Library}\label{biodiversity-heritage-library}}
-
-The Biodiversity Heritage Library (BHL) houses scans of biodiversity
-books, and provides web interfaces and APIs to query and fetch those
-data. They also provide text of the scanned pages. The main R interace
-to BHL is through \href{https://github.com/ropensci/rbhl}{rbhl}.
-
-\hypertarget{public-library-of-science}{%
-\subsection{Public Library of Science}\label{public-library-of-science}}
-
-The Public Library of Science (PLOS) is one of the largest open access
-only publishers. They as of this writing have published 2.1 million
-articles. One of the strongs advantages of PLOS is that they provide an
-API to their Solr instance, which is a very flexible way to search their
-articles. The main R interace to PLOS is through
-\href{https://github.com/ropensci/rplos}{rplos}.
-
-\hypertarget{elsevierscopus}{%
-\subsection{Elsevier/Scopus}\label{elsevierscopus}}
-
-Elsevier is one of the largest publishers. Scopus is one of their
-products, a searchable database of articles. Most of their articles are
-not open access. However, they have a number of advantages if you have
-access to their articles: they are one of the few publishers to provide
-machine readable XML (many publishers do have XML versions of articles,
-but do not provide it, as mentioned in the Introduction); they are one
-of the few (two) publishers that take part in Crossref's text and data
-mining program. The packages
-\href{https://github.com/ropensci/fulltext}{fulltext} and
-\href{https://github.com/ropensci/crminer}{crminer} can be used to
-access Elsevier articles through Crossref's TDM program. There's an
-interface to Scopus article search within
-\href{https://github.com/ropensci/fulltext}{fulltext}.
-
-\hypertarget{arxivbiorxiv}{%
-\subsection{arXiv/bioRxiv}\label{arxivbiorxiv}}
-
-arXiv and bioRxiv are preprint publishers, the former in existence for
-many years, and the latter new on the scene. Preprints are scholarly
-articles that are generally not peer-reviewed, but that for the most
-part will later be published in a different peer-review publication. You
-can access articles from these publishers through
-\href{https://github.com/ropensci/fulltext}{fulltext}. arXiv does
-provide a web API that we hook into; bioRxiv does not, but we can get
-you articles nonetheless.
-
-\hypertarget{pubmedpmceurope-pmc}{%
-\subsection{Pubmed/PMC/Europe PMC}\label{pubmedpmceurope-pmc}}
-
-Pubmed/PMC is a corpus/website of NIH funded research in the United
-States; while Europe PMC is an equivalent for the European Union. You
-can access articles from Pubmed/PMC through
-\href{https://github.com/ropensci/fulltext}{fulltext}, and for Europe
-PMC through \href{https://github.com/ropensci/europepmc}{europepmc}.
-
-\hypertarget{microsoft-academic-research}{%
-\subsection{Microsoft Academic
-Research}\label{microsoft-academic-research}}
-
-Microsoft Academic Research (MAR) is a search engine for research
-articles. You can use their GUI web interface to search, and they
-provide APIs for programmatic access. The R interface for MAR is
-\href{https://github.com/ropensci/microdemic}{microdemic}; and
-\href{https://github.com/ropensci/fulltext}{fulltext} hooks into
-\texttt{microdemic} as well for article search and abstract retrieval.
-
-\hypertarget{directory-of-open-access-journals}{%
-\subsection{Directory of Open Access
-Journals}\label{directory-of-open-access-journals}}
-
-Directory of Open Access Journals (DOAJ) maintains data on open access
-journals, as well as some portion of the articles in those journals.
-Thus, you can search for journals as well as articles with DOAJ. The R
-interface for DOAJ is \href{https://github.com/ropensci/jaod}{jaod}.
-
-\hypertarget{jstor}{%
-\subsection{JSTOR}\label{jstor}}
-
-JSTOR's Data for Research program gives institutions with access to
-JSTOR, access to full text of articles within JSTOR. There is no way
-however to make the interaction with JSTOR completely programmatic, thus
-making reproducible research very difficult. Nonetheless, there is an R
-package (\href{https://github.com/ropensci/jstor}{jstor}) for using data
-from JSTOR's Data for Research.
-
-\hypertarget{orcid}{%
-\subsection{ORCID}\label{orcid}}
-
-ORCID (\url{https://orcid.org/}) is an organization keeping track of
-identifiers and metadata for researchers around the world. Individuals
-can optionally maintain metadata on their scholarly works connected to
-their account with ORCID. Thus, across all of ORCID, a significant cache
-of metadata is accruing on scholarly works, their funding amounts,
-collaborators, etc., useful for bibliometrics research and more. The R
-interface for ORCID is
-\href{https://github.com/ropensci/rorcid}{rorcid}.
-
-\hypertarget{citoidopen-citation-corpus}{%
-\subsection{Citoid/Open Citation
-Corpus}\label{citoidopen-citation-corpus}}
-
-The Open Citation Corpus (\url{http://opencitations.net/}) holds records
-of which articles cite which other articles, allowing for all important
-research on the scholarly web of citation. Citation data has been very
-closely guarded until recently, but the largest publishers are still not
-contributing to the Open Citation Corpus. The R interface to the Open
-Citation Corpus is
-\href{https://github.com/ropenscilabs/rcitoid}{rcitoid}.
-
-\hypertarget{fatcat}{%
-\subsection{Fatcat}\label{fatcat}}
-
-Fatcat is a project from Ben Newbold of the Internet Archive Labs. It is
-a ``versioned, publicly-editable catalog of research publications:
-journal articles, conference proceedings, preprints, blog posts''.
-Fatcat is currently does not have an R client, but is used inside of the
-\href{https://github.com/ropensci/fulltext}{fulltext} package.
-
-\hypertarget{sherparomeo}{%
-\subsection{SHERPA/RoMEO}\label{sherparomeo}}
-
-SHERPA/RoMEO (\url{http://sherpa.mimas.ac.uk/romeo/index.php})
-aggregates and analyses publisher open access policies and provides
-summaries of self-archiving permissions and conditions of rights given
-to authors. The {[}rromeo{]}{[}{]} is an R interface to SHERPA/RoMEO.
-
-\hypertarget{core}{%
-\subsection{CORE}\label{core}}
-
-CORE (\url{https://core.ac.uk/}) touts itself as the world's largest
-collection of open access research articles, providing metadata on
-journals and articles, as well as access to the full text of articles.
-The \href{https://github.com/ropensci/rcoreoa}{rcoreoa} R package
-interfaces with the CORE API.
-
-\hypertarget{dissemin}{%
-\subsection{Dissemin}\label{dissemin}}
-
-Dissemin (\url{https://dissem.in/}) detects papers behind pay-walls and
-invites authors to upload them to an open repository. Dissemin provides
-metadata including links to open versions of articles. The
-{[}dissemr{]}{[}{]} R package interfaces with the Dissemin API.
-
-\hypertarget{fulltext-a-toolset-for-text-mining-in-r}{%
-\section{fulltext: a toolset for text mining in
-R}\label{fulltext-a-toolset-for-text-mining-in-r}}
-
-\href{https://github.com/ropensci/fulltext}{fulltext} is a general
-purpose R package for the data part of text mining: search for articles,
-get links to articles, get article abstracts, and fetch full text of
-articles. The \texttt{fulltext} package is always adding additional data
-sources as time allows (See Table 1). Starting from searching for
-articles, the outputs of search can be fed into a function to get links
-to those articles, or to get abstracts for those articles, or to fetch
-their full text. The following is a breakdown of the major distinct
-parts of \texttt{fulltext}.
-
-\hypertarget{search}{%
-\subsection{Search}\label{search}}
-
-\texttt{ft\_search()} provides search access to nine different data
-sources (PLOS, BMC, Crossref, Entrez, arXiv, bioRxiv, Europe PMC,
-Scopus, Microsoft Academic), creating a mostly unified interface to all
-data sources. The parts of each data source that are common are for the
-most part factored out into the parameters of the \texttt{ft\_search()}
-function: query term(s), pagination (number of results, result number to
-start at). In addition, we allow the user to pass on data source
-specific options to refine the search per data source.
-
-With \texttt{ft\_search()}, you can query any combination of the nine
-data sources at once. The returned object is a list, with access to
-results of each data source by its name (e.g., \texttt{\$plos}, or
-\texttt{\$crossref}). For each data source, the returned object does
-vary because the returned data from each data source widely varies; for
-the most part data.frame's are returned. For those data sources not
-queried, their slot is empty.
-
-One important aspect of the research result we highlight is the licenses
-in the returned data for each data source.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{x <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \KeywordTok{c}\NormalTok{(}\StringTok{"plos"}\NormalTok{, }\StringTok{"crossref"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-The results for this PLOS search have all CC-BY licenses
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{x}\OperatorTok{$}\NormalTok{plos}
-\CommentTok{#> Query: [ecology] }
-\CommentTok{#> Records found, returned: [47257, 10] }
-\CommentTok{#> License: [CC-BY] }
-\CommentTok{#>                              id}
-\CommentTok{#> 1  10.1371/journal.pone.0001248}
-\CommentTok{#> 2  10.1371/journal.pone.0059813}
-\CommentTok{#> 3  10.1371/journal.pone.0080763}
-\CommentTok{#> 4  10.1371/journal.pone.0155019}
-\CommentTok{#> 5  10.1371/journal.pone.0175014}
-\CommentTok{#> 6  10.1371/journal.pone.0150648}
-\CommentTok{#> 7  10.1371/journal.pone.0208370}
-\CommentTok{#> 8  10.1371/journal.pcbi.1003594}
-\CommentTok{#> 9  10.1371/journal.pone.0102437}
-\CommentTok{#> 10 10.1371/journal.pone.0166559}
-\end{Highlighting}
-\end{Shaded}
-
-Whereas the results for this Crossref search have mixed licenses
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{x}\OperatorTok{$}\NormalTok{crossref}
-\CommentTok{#> Query: [ecology] }
-\CommentTok{#> Records found, returned: [164657, 10] }
-\CommentTok{#> License: [variable, see individual records] }
-\CommentTok{#>    archive               container.title    created  deposited}
-\CommentTok{#> 1  Portico                       Ecology 2006-05-03 2018-08-04}
-\CommentTok{#> 2  Portico                       Ecology 2006-05-03 2018-08-04}
-\CommentTok{#> 3       NA                       Ecology 2006-05-03 2018-08-04}
-\CommentTok{#> 4       NA                       Ecology 2006-05-03 2018-08-04}
-\CommentTok{#> 5       NA                       Ecology 2006-05-03 2018-08-04}
-\CommentTok{#> 6       NA                       Ecology 2006-05-03 2018-08-04}
-\CommentTok{#> 7       NA                       Ecology 2006-05-09 2018-08-01}
-\CommentTok{#> 8  Portico                       Ecology 2017-04-26 2019-03-08}
-\CommentTok{#> 9       NA Trends in Ecology & Evolution 2002-07-25 2017-06-14}
-\CommentTok{#> 10      NA Journal of Industrial Ecology 2014-11-21 2017-06-23}
-\CommentTok{#> Variables not shown: published.print (chr), published.online (chr), doi}
-\CommentTok{#>      (chr), indexed (chr), issn (chr), issue (chr), issued (chr), member}
-\CommentTok{#>      (chr), page (chr), prefix (chr), publisher (chr), reference.count}
-\CommentTok{#>      (chr), score (chr), source (chr), title (chr), type (chr), url (chr),}
-\CommentTok{#>      volume (chr), author (list), link (list), license (list), subject}
-\CommentTok{#>      (chr), alternative.id (chr), subtitle (chr), reference (list)}
-\end{Highlighting}
-\end{Shaded}
-
-You can dig into the license field for each article, with URLs holding
-information on each license
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{vapply}\NormalTok{(x}\OperatorTok{$}\NormalTok{crossref}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{license, }\ControlFlowTok{function}\NormalTok{(w) w}\OperatorTok{$}\NormalTok{URL[}\DecValTok{1}\NormalTok{], }\StringTok{""}\NormalTok{)}
-\CommentTok{#>  [1] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
-\CommentTok{#>  [2] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
-\CommentTok{#>  [3] "http://doi.wiley.com/10.1002/tdm_license_1"  }
-\CommentTok{#>  [4] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
-\CommentTok{#>  [5] "http://doi.wiley.com/10.1002/tdm_license_1"  }
-\CommentTok{#>  [6] "http://doi.wiley.com/10.1002/tdm_license_1"  }
-\CommentTok{#>  [7] "http://doi.wiley.com/10.1002/tdm_license_1"  }
-\CommentTok{#>  [8] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
-\CommentTok{#>  [9] "http://www.elsevier.com/tdm/userlicense/1.0/"}
-\CommentTok{#> [10] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{links}{%
-\subsection{Links}\label{links}}
-
-\texttt{ft\_links()} provides two pathways to get links (URLs) for
-articles, with a choice of four different data sources (PLOS, BMC,
-Crossref, Entrez). First, you can use \texttt{ft\_search()}, then pass
-the output of that function to \texttt{ft\_links()}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{out <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"entrez"}\NormalTok{)}
-\KeywordTok{ft_links}\NormalTok{(out)}
-\CommentTok{#> <fulltext links>}
-\CommentTok{#> [Found] 6 }
-\CommentTok{#> [IDs] ID_30964001 ID_30962485 ID_30962432 ID_30952928 ID_30674747}
-\CommentTok{#>      ID_30674743 ...}
-\end{Highlighting}
-\end{Shaded}
-
-Second, you can pass DOIs directly to \texttt{ft\_links()}. Both end up
-at the same point, links for each article, if they could be found for
-the user selected data source.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# }\AlertTok{FIXME}
-\KeywordTok{ft_links}\NormalTok{(out}\OperatorTok{$}\NormalTok{entrez}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{doi)}
-\end{Highlighting}
-\end{Shaded}
-
-The biggest caveat with \texttt{ft\_links()} is that we can't gaurantee
-that the links will work. Link rot is one way in which the links may not
-work: link rot is when the URL does not point to the original content
-anymore, or fails altogether. Additionally, with Crossref, publishers
-can deposit URLs for articles, but they make change the URLs at some
-later date but not update the URLs with Crossref.
-
-\hypertarget{abstracts}{%
-\subsection{Abstracts}\label{abstracts}}
-
-\texttt{ft\_abstract()} provides access to article abstracts from four
-different data sources (PLOS, Scopus, Microsoft Academic Research,
-Crossref). The only way to use the function is to pass article
-identifiers, which are for the most DOIs.
-
-The advantage of abstracts over full text is that abstracts can often be
-retrieved even for paywalled articles. That is, you can have much
-broader coverage of the articles you're targeting relative to full text.
-
-If you are after abstracts, and you are already getting or already have
-full text, and if the articles are in XML format, then you can use
-\href{https://github.com/ropensci/pubchunks}{pubchunks} to extract out
-the abstracts.
-
-\hypertarget{fetch-full-text}{%
-\subsection{Fetch full text}\label{fetch-full-text}}
-
-\texttt{ft\_get()} fetchs full text of articles from many different data
-sources. From the DOIs that are passed in to the function, we detect the
-publisher, and there are specific plugins for certain publishers: AAAS,
-American Institute of Physics, American Society of Clinical Oncology,
-American Society for Microbiology, arXiv, bioRxiv, BiomedCentral,
-Copernicus, Crossref, Elife, Elsevier, Pubmed/PMC via NCBI's Entrez,
-Frontiers, IEEE, Informa, Instituto de Investigaciones Filologicas,
-American Medical Association, Microbiology Society, PeerJ, Pensoft,
-PLOS, PNAS, Royal Society of Chemistry, ScienceDirect, Scientific
-Societies, and Wiley.
-
-If there's no built-in plugin for the publisher already, we use the
-FTDOI API (\url{https://ftdoi.org}) to try to get the link for the full
-text of the article. If the FTDOI API doesn't bear fruit, we search
-Crossref for a link to the full text. If Crossref doesn't have any full
-text links, we give up.
-
-Since users can go through a lot of article requests, we cache
-successfully downloaded articles, and keep that knowledge consistent
-across R sessions; all subsequent requests for the same article just use
-the cached version. Additionally, all errors in \texttt{ft\_get()} are
-collected in a data.frame in the output of the function to help the user
-quickly determine what went wrong.
-
-\hypertarget{how-to-text-mine-from-r-three-case-studies}{%
-\section{How to text mine from R: Three case
-studies}\label{how-to-text-mine-from-r-three-case-studies}}
+\hypertarget{text-mining-in-r-three-case-studies}{%
+\section{Text mining in R: Three case
+studies}\label{text-mining-in-r-three-case-studies}}
 
@@
 
+\hypertarget{data-sources}{%
+\section{Data sources}\label{data-sources}}
+
+There is increasing open access scientific literature content available
+online. However, only a small proportion of scientific journals provide
+access to their full text; whereas, most publishers provide open access
+to their metadata only (most often through Crossref; Table 1). The
+following is a synopsis of the major data sources and associated R
+tools.
+
+\newpage
+
+Table 1. Sources of scientific literature, their content type provided
+via web services, whether rOpenSci has an R packages for the service,
+and where to find the API documentation.
+
+\begin{longtable}[]{@{}llll@{}}
+\toprule
+Data Provider & Content Type & rOpenSci Package &
+Documentation\tabularnewline
+\midrule
+\endhead
+Crossref & Metadata & rcrossref/crminer & \footnote{\url{https://api.crossref.org}}\tabularnewline
+DataCite & Metadata & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
+Biodiversity Heritage Library & Full text/Metadata & rbhl & \footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
+Public Library of Science (PLoS) & Full text (pdf/xml)/Metadata & rplos
+& \footnote{\url{http://api.plos.org/solr}}\tabularnewline
+Scopus (Elsevier) & Full text (pdf/xml)/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
+arXiv & Full text (pdf)/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
+Biomed Central (via Springer) & Full text (pdf)/Metadata & fulltext &
+\footnote{\url{https://dev.springer.com/}}\tabularnewline
+bioRxiv & Full text (pdf)/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
+PMC/Pubmed (via Entrez) & Full text (pdf/xml)/Metadata & rentrez &
+\footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
+Europe PMC & Full text (pdf/xml)/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
+Microsoft Academic Search & Metadata & microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
+Directory of Open Access Journals & Metadata & jaod & \footnote{\url{https://doaj.org/api/v1/docs}}\tabularnewline
+JSTOR Data for Research & Full text & jstor & \footnote{\url{https://www.jstor.org/dfr/}}\tabularnewline
+ORCID & Metadata & rorcid & \footnote{\url{https://pub.orcid.org/}}\tabularnewline
+Wikimedia's Citoid & Citations & rcitoid & \footnote{\url{https://en.wikipedia.org/api/rest_v1/\#/Citation/getCitation}}\tabularnewline
+Open Citation Corpus & Citations & citecorp & \footnote{\url{http://opencitations.net/}}\tabularnewline
+Fatcat & Metadata & none & \footnote{\url{https://fatcat.wiki/}}\tabularnewline
+SHERPA/RoMEO & Journal Level Metadata & rromeo & \footnote{\url{http://www.sherpa.ac.uk/romeo/apimanual.php?la=en\&fIDnum=\%7C\&mode=simple}}\tabularnewline
+CORE & Full text (pdf)/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
+Dissemin & Metadata & dissemr & \footnote{\url{https://dissemin.readthedocs.io/en/latest/api.html}}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{crossref-and-datacite}{%
+\subsection{Crossref and Datacite}\label{crossref-and-datacite}}
+
+Crossref is a non-profit that creates (or ``mints'') Digital Object
+Identifiers (DOIs). In addition, they maintain metadata associated with
+each DOI. The metadata ranges from simple (including author, title,
+dates, DOI, type, publisher) to including number of citations to the
+article, as well as references in the article, and even abstracts. At
+the time of writing they hold 100 million DOIs.
+
+One can search by DOI or search citation data to get citations. In
+addition, Crossref has a text mining opt-in program for publishers. The
+result of this is that some publishers provide URLs for full text
+content of their articles. The majority of these links are pay-walled,
+while some are open access. Using any of the various tools for working
+with Crossref data, you can filter your search to get only articles with
+full text links, and further to get only articles with full text links
+that are open access.
+
+The main interfaces for Crossref in R are
+\href{https://github.com/ropensci/rcrossref}{rcrossref} and
+\href{https://github.com/ropensci/crminer}{crminer}. rcrossref is a
+complete client for the public facing Crossref web services including
+metadata, whereas crminer focuses only on retrieving full text of
+articles. Similar interfaces to rcrossref are available in Ruby
+(\href{https://github.com/sckott/serrano}{serrano}) and Python
+(\href{https://github.com/sckott/habanero}{habanero}).
+
+Datacite is similar to Crossref, but focuses on datasets instead of
+articles. The main interface for Datacite in R is
+\href{https://github.com/ropensci/rdatacite}{rdatacite}.
+
+\hypertarget{biodiversity-heritage-library}{%
+\subsection{Biodiversity Heritage
+Library}\label{biodiversity-heritage-library}}
+
+The Biodiversity Heritage Library (BHL) houses scans of biodiversity
+books, and provides web interfaces and APIs to query and fetch those
+data. They also provide text of the scanned pages. The main R interace
+to BHL is through \href{https://github.com/ropensci/rbhl}{rbhl}.
+
+\hypertarget{public-library-of-science}{%
+\subsection{Public Library of Science}\label{public-library-of-science}}
+
+The Public Library of Science (PLOS) is one of the largest open access
+only publishers. They as of this writing have published 2.1 million
+articles. One of the strongs advantages of PLOS is that they provide an
+API to their Solr instance, which is a very flexible way to search their
+articles. The main R interace to PLOS is through
+\href{https://github.com/ropensci/rplos}{rplos}.
+
+\hypertarget{elsevierscopus}{%
+\subsection{Elsevier/Scopus}\label{elsevierscopus}}
+
+Elsevier is one of the largest publishers. Scopus is one of their
+products, a searchable database of articles. Most of their articles are
+not open access. However, they have a number of advantages if you have
+access to their articles: they are one of the few publishers to provide
+machine readable XML (many publishers do have XML versions of articles,
+but do not provide it, as mentioned in the Introduction); they are one
+of the few (two) publishers that take part in Crossref's text and data
+mining program. The packages
+\href{https://github.com/ropensci/fulltext}{fulltext} and
+\href{https://github.com/ropensci/crminer}{crminer} can be used to
+access Elsevier articles through Crossref's TDM program. There's an
+interface to Scopus article search within
+\href{https://github.com/ropensci/fulltext}{fulltext}.
+
+\hypertarget{arxivbiorxiv}{%
+\subsection{arXiv/bioRxiv}\label{arxivbiorxiv}}
+
+arXiv and bioRxiv are preprint publishers, the former in existence for
+many years, and the latter new on the scene. Preprints are scholarly
+articles that are generally not peer-reviewed, but that for the most
+part will later be published in a different peer-review publication. You
+can access articles from these publishers through
+\href{https://github.com/ropensci/fulltext}{fulltext}. arXiv does
+provide a web API that we hook into; bioRxiv does not, but we can get
+you articles nonetheless.
+
+\hypertarget{pubmedpmceurope-pmc}{%
+\subsection{Pubmed/PMC/Europe PMC}\label{pubmedpmceurope-pmc}}
+
+Pubmed/PMC is a corpus/website of NIH funded research in the United
+States; while Europe PMC is an equivalent for the European Union. You
+can access articles from Pubmed/PMC through
+\href{https://github.com/ropensci/fulltext}{fulltext}, and for Europe
+PMC through \href{https://github.com/ropensci/europepmc}{europepmc}.
+
+\hypertarget{microsoft-academic-research}{%
+\subsection{Microsoft Academic
+Research}\label{microsoft-academic-research}}
+
+Microsoft Academic Research (MAR) is a search engine for research
+articles. You can use their GUI web interface to search, and they
+provide APIs for programmatic access. The R interface for MAR is
+\href{https://github.com/ropensci/microdemic}{microdemic}; and
+\href{https://github.com/ropensci/fulltext}{fulltext} hooks into
+\texttt{microdemic} as well for article search and abstract retrieval.
+
+\hypertarget{directory-of-open-access-journals}{%
+\subsection{Directory of Open Access
+Journals}\label{directory-of-open-access-journals}}
+
+Directory of Open Access Journals (DOAJ) maintains data on open access
+journals, as well as some portion of the articles in those journals.
+Thus, you can search for journals as well as articles with DOAJ. The R
+interface for DOAJ is \href{https://github.com/ropensci/jaod}{jaod}.
+
+\hypertarget{jstor}{%
+\subsection{JSTOR}\label{jstor}}
+
+JSTOR's Data for Research program gives institutions with access to
+JSTOR, access to full text of articles within JSTOR. There is no way
+however to make the interaction with JSTOR completely programmatic, thus
+making reproducible research very difficult. Nonetheless, there is an R
+package (\href{https://github.com/ropensci/jstor}{jstor}) for using data
+from JSTOR's Data for Research.
+
+\hypertarget{orcid}{%
+\subsection{ORCID}\label{orcid}}
+
+ORCID (\url{https://orcid.org/}) is an organization keeping track of
+identifiers and metadata for researchers around the world. Individuals
+can optionally maintain metadata on their scholarly works connected to
+their account with ORCID. Thus, across all of ORCID, a significant cache
+of metadata is accruing on scholarly works, their funding amounts,
+collaborators, etc., useful for bibliometrics research and more. The R
+interface for ORCID is
+\href{https://github.com/ropensci/rorcid}{rorcid}.
+
+\hypertarget{citoidopen-citation-corpus}{%
+\subsection{Citoid/Open Citation
+Corpus}\label{citoidopen-citation-corpus}}
+
+The Open Citation Corpus (\url{http://opencitations.net/}) holds records
+of which articles cite which other articles, allowing for all important
+research on the scholarly web of citation. Citation data has been very
+closely guarded until recently, but the largest publishers are still not
+contributing to the Open Citation Corpus. The R interface to the Open
+Citation Corpus is
+\href{https://github.com/ropenscilabs/rcitoid}{rcitoid}.
+
+\hypertarget{fatcat}{%
+\subsection{Fatcat}\label{fatcat}}
+
+Fatcat is a project from Ben Newbold of the Internet Archive Labs. It is
+a ``versioned, publicly-editable catalog of research publications:
+journal articles, conference proceedings, preprints, blog posts''.
+Fatcat is currently does not have an R client, but is used inside of the
+\href{https://github.com/ropensci/fulltext}{fulltext} package.
+
+\hypertarget{sherparomeo}{%
+\subsection{SHERPA/RoMEO}\label{sherparomeo}}
+
+SHERPA/RoMEO (\url{http://sherpa.mimas.ac.uk/romeo/index.php})
+aggregates and analyses publisher open access policies and provides
+summaries of self-archiving permissions and conditions of rights given
+to authors. The {[}rromeo{]}{[}{]} is an R interface to SHERPA/RoMEO.
+
+\hypertarget{core}{%
+\subsection{CORE}\label{core}}
+
+CORE (\url{https://core.ac.uk/}) touts itself as the world's largest
+collection of open access research articles, providing metadata on
+journals and articles, as well as access to the full text of articles.
+The \href{https://github.com/ropensci/rcoreoa}{rcoreoa} R package
+interfaces with the CORE API.
+
+\hypertarget{dissemin}{%
+\subsection{Dissemin}\label{dissemin}}
+
+Dissemin (\url{https://dissem.in/}) detects papers behind pay-walls and
+invites authors to upload them to an open repository. Dissemin provides
+metadata including links to open versions of articles. The
+{[}dissemr{]}{[}{]} R package interfaces with the Dissemin API.
+
+\hypertarget{learn-more}{%
+\subsection{Learn more}\label{learn-more}}
+
+To learn how to use the above packages, follow the links to get to
+documentation, and how to report issues. The rOpenSci forum
+(\url{https://discuss.ropensci.org/}) is a good place to ask questions
+about these packages.
+
+\hypertarget{fulltext-a-toolset-for-text-mining-in-r}{%
+\section{fulltext: a toolset for text mining in
+R}\label{fulltext-a-toolset-for-text-mining-in-r}}
+
+\href{https://github.com/ropensci/fulltext}{fulltext} is a general
+purpose R package for the data part of text mining: search for articles,
+get links to articles, get article abstracts, and fetch full text of
+articles. The \texttt{fulltext} package is always adding additional data
+sources as time allows (See Table 1). Starting from searching for
+articles, the outputs of search can be fed into a function to get links
+to those articles, or to get abstracts for those articles, or to fetch
+their full text. The following is a breakdown of the major distinct
+parts of \texttt{fulltext}.
+
+\hypertarget{search}{%
+\subsection{Search}\label{search}}
+
+\texttt{ft\_search()} provides search access to nine different data
+sources (PLOS, BMC, Crossref, Entrez, arXiv, bioRxiv, Europe PMC,
+Scopus, Microsoft Academic), creating a mostly unified interface to all
+data sources. The parts of each data source that are common are for the
+most part factored out into the parameters of the \texttt{ft\_search()}
+function: query term(s), pagination (number of results, result number to
+start at). In addition, we allow the user to pass on data source
+specific options to refine the search per data source.
+
+With \texttt{ft\_search()}, you can query any combination of the nine
+data sources at once. The returned object is a list, with access to
+results of each data source by its name (e.g., \texttt{\$plos}, or
+\texttt{\$crossref}). For each data source, the returned object does
+vary because the returned data from each data source widely varies; for
+the most part data.frame's are returned. For those data sources not
+queried, their slot is empty.
+
+One important aspect of the research result we highlight is the licenses
+in the returned data for each data source.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \KeywordTok{c}\NormalTok{(}\StringTok{"plos"}\NormalTok{, }\StringTok{"crossref"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+The results for this PLOS search have all CC-BY licenses
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x}\OperatorTok{$}\NormalTok{plos}
+\CommentTok{#> Query: [ecology] }
+\CommentTok{#> Records found, returned: [47257, 10] }
+\CommentTok{#> License: [CC-BY] }
+\CommentTok{#>                              id}
+\CommentTok{#> 1  10.1371/journal.pone.0001248}
+\CommentTok{#> 2  10.1371/journal.pone.0059813}
+\CommentTok{#> 3  10.1371/journal.pone.0080763}
+\CommentTok{#> 4  10.1371/journal.pone.0155019}
+\CommentTok{#> 5  10.1371/journal.pone.0175014}
+\CommentTok{#> 6  10.1371/journal.pone.0150648}
+\CommentTok{#> 7  10.1371/journal.pone.0208370}
+\CommentTok{#> 8  10.1371/journal.pcbi.1003594}
+\CommentTok{#> 9  10.1371/journal.pone.0102437}
+\CommentTok{#> 10 10.1371/journal.pone.0166559}
+\end{Highlighting}
+\end{Shaded}
+
+Whereas the results for this Crossref search have mixed licenses
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x}\OperatorTok{$}\NormalTok{crossref}
+\CommentTok{#> Query: [ecology] }
+\CommentTok{#> Records found, returned: [164657, 10] }
+\CommentTok{#> License: [variable, see individual records] }
+\CommentTok{#>    archive               container.title    created  deposited}
+\CommentTok{#> 1  Portico                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 2  Portico                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 3       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 4       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 5       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 6       NA                       Ecology 2006-05-03 2018-08-04}
+\CommentTok{#> 7       NA                       Ecology 2006-05-09 2018-08-01}
+\CommentTok{#> 8  Portico                       Ecology 2017-04-26 2019-03-08}
+\CommentTok{#> 9       NA Trends in Ecology & Evolution 2002-07-25 2017-06-14}
+\CommentTok{#> 10      NA Journal of Industrial Ecology 2014-11-21 2017-06-23}
+\CommentTok{#> Variables not shown: published.print (chr), published.online (chr), doi}
+\CommentTok{#>      (chr), indexed (chr), issn (chr), issue (chr), issued (chr), member}
+\CommentTok{#>      (chr), page (chr), prefix (chr), publisher (chr), reference.count}
+\CommentTok{#>      (chr), score (chr), source (chr), title (chr), type (chr), url (chr),}
+\CommentTok{#>      volume (chr), author (list), link (list), license (list), subject}
+\CommentTok{#>      (chr), alternative.id (chr), subtitle (chr), reference (list)}
+\end{Highlighting}
+\end{Shaded}
+
+You can dig into the license field for each article, with URLs holding
+information on each license
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{vapply}\NormalTok{(x}\OperatorTok{$}\NormalTok{crossref}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{license, }\ControlFlowTok{function}\NormalTok{(w) w}\OperatorTok{$}\NormalTok{URL[}\DecValTok{1}\NormalTok{], }\StringTok{""}\NormalTok{)}
+\CommentTok{#>  [1] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [2] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [3] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [4] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [5] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [6] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [7] "http://doi.wiley.com/10.1002/tdm_license_1"  }
+\CommentTok{#>  [8] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\CommentTok{#>  [9] "http://www.elsevier.com/tdm/userlicense/1.0/"}
+\CommentTok{#> [10] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{links}{%
+\subsection{Links}\label{links}}
+
+\texttt{ft\_links()} provides two pathways to get links (URLs) for
+articles, with a choice of four different data sources (PLOS, BMC,
+Crossref, Entrez). First, you can use \texttt{ft\_search()}, then pass
+the output of that function to \texttt{ft\_links()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{out <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"entrez"}\NormalTok{)}
+\KeywordTok{ft_links}\NormalTok{(out)}
+\CommentTok{#> <fulltext links>}
+\CommentTok{#> [Found] 6 }
+\CommentTok{#> [IDs] ID_30964001 ID_30962485 ID_30962432 ID_30952928 ID_30674747}
+\CommentTok{#>      ID_30674743 ...}
+\end{Highlighting}
+\end{Shaded}
+
+Second, you can pass DOIs directly to \texttt{ft\_links()}. Both end up
+at the same point, links for each article, if they could be found for
+the user selected data source.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# }\AlertTok{FIXME}
+\KeywordTok{ft_links}\NormalTok{(out}\OperatorTok{$}\NormalTok{entrez}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{doi)}
+\end{Highlighting}
+\end{Shaded}
+
+The biggest caveat with \texttt{ft\_links()} is that we can't gaurantee
+that the links will work. Link rot is one way in which the links may not
+work: link rot is when the URL does not point to the original content
+anymore, or fails altogether. Additionally, with Crossref, publishers
+can deposit URLs for articles, but they make change the URLs at some
+later date but not update the URLs with Crossref.
+
+\hypertarget{abstracts}{%
+\subsection{Abstracts}\label{abstracts}}
+
+\texttt{ft\_abstract()} provides access to article abstracts from four
+different data sources (PLOS, Scopus, Microsoft Academic Research,
+Crossref). The only way to use the function is to pass article
+identifiers, which are for the most DOIs.
+
+The advantage of abstracts over full text is that abstracts can often be
+retrieved even for paywalled articles. That is, you can have much
+broader coverage of the articles you're targeting relative to full text.
+
+If you are after abstracts, and you are already getting or already have
+full text, and if the articles are in XML format, then you can use
+\href{https://github.com/ropensci/pubchunks}{pubchunks} to extract out
+the abstracts.
+
+\hypertarget{fetch-full-text}{%
+\subsection{Fetch full text}\label{fetch-full-text}}
+
+\texttt{ft\_get()} fetchs full text of articles from many different data
+sources. From the DOIs that are passed in to the function, we detect the
+publisher, and there are specific plugins for certain publishers: AAAS,
+American Institute of Physics, American Society of Clinical Oncology,
+American Society for Microbiology, arXiv, bioRxiv, BiomedCentral,
+Copernicus, Crossref, Elife, Elsevier, Pubmed/PMC via NCBI's Entrez,
+Frontiers, IEEE, Informa, Instituto de Investigaciones Filologicas,
+American Medical Association, Microbiology Society, PeerJ, Pensoft,
+PLOS, PNAS, Royal Society of Chemistry, ScienceDirect, Scientific
+Societies, and Wiley.
+
+If there's no built-in plugin for the publisher already, we use the
+FTDOI API (\url{https://ftdoi.org}) to try to get the link for the full
+text of the article. If the FTDOI API doesn't bear fruit, we search
+Crossref for a link to the full text. If Crossref doesn't have any full
+text links, we give up.
+
+Since users can go through a lot of article requests, we cache
+successfully downloaded articles, and keep that knowledge consistent
+across R sessions; all subsequent requests for the same article just use
+the cached version. Additionally, all errors in \texttt{ft\_get()} are
+collected in a data.frame in the output of the function to help the user
+quickly determine what went wrong.
+
 \hypertarget{future-directions}{%

### 2019-08-23 manuscript.tex
fix #18 link to info on where to go after getting full text

@@
 
+\emph{Further processing}
+
+From here, you'll likely want to use R packages for removing stopwords,
+tokenizing, and more. The Natural Language Processing CRAN Task View
+(\url{https://cloud.r-project.org/web/views/NaturalLanguageProcessing.html})
+is a good place to start if you're not sure what packages to use. Some
+packages worth trying in this space are
+\href{https://github.com/ropensci/antiword}{antiword},
+\href{https://github.com/ropensci/pdftools}{pdftools},
+\href{https://github.com/ropensci/rtika}{rtika},
+\href{https://github.com/ropensci/tokenizers}{tokenizers}, and
+\href{https://github.com/ropensci/textreuse}{textreuse}.
+
 \hypertarget{data-sources}{%

### 2019-08-23 manuscript.Rmd
fix #21 learn more about the pkgs at end of data sources section
fix #20 move case studies further up
fix #22 add section on papers that have used fulltext had to move use cases back down lower because really need to have talked about fulltext pkg first

@@
 
+## Learn more
+
+To learn how to use the above packages, follow the links to get to documentation,
+and how to report issues. The rOpenSci forum (<https://discuss.ropensci.org/>) is
+a good place to ask questions about these packages.
+
 # fulltext: a toolset for text mining in R
@@
 
-# How to text mine from R: Three case studies
+# Text mining in R: Use Cases
 
-## Case study 1: Citation mining
+## Use cases in the literature
+
+At least three papers have used the fulltext package. @Bauer2016 explored
+a way to think about the qualitative impact of citations rather than simple
+counts of citations; they did not use fulltext, but pointed it out as a good
+option for collecting papers. @Walsh2018 used fulltext and [roadoi][], an
+rOpenSci R package, to "map the literature" on nutritional interventions
+in cognitive health. @Piper2019 in a review of DNA metabarcoding for
+high-throughput insect surveillance used fulltext, along with [rentrez][]
+(an rOpenSci package), to collect articles for text mining.
+
+## Three case studies
+
+The following three case studies demonstrate in brief how one could use
+fulltext and related packages to do a few different text mining tasks.
+
+### Case study 1: Citation mining
 

### 2019-08-23 manuscript.tex
fix #22 add section on papers that have used fulltext had to move use cases back down lower because really need to have talked about fulltext pkg first

@@
 
-\hypertarget{text-mining-in-r-three-case-studies}{%
-\section{Text mining in R: Three case
-studies}\label{text-mining-in-r-three-case-studies}}
-
-\hypertarget{case-study-1-citation-mining}{%
-\subsection{Case study 1: Citation
-mining}\label{case-study-1-citation-mining}}
-
-In this example, xxxx
-
-\emph{Load libraries}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"rcrossref"}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{"rplos"}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{"rorcid"}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{"rcitoid"}\NormalTok{)}
-\KeywordTok{library}\NormalTok{(}\StringTok{"citecorp"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{rcrossref}
-
-Using \texttt{rcrossref} for Crossref data:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{x <-}\StringTok{ }\KeywordTok{cr_works}\NormalTok{(}\DataTypeTok{query=}\StringTok{"NSF"}\NormalTok{)}
-\KeywordTok{head}\NormalTok{(x}\OperatorTok{$}\NormalTok{data)}
-\CommentTok{#> # A tibble: 6 x 32}
-\CommentTok{#>   alternative.id container.title created deposited published.print doi  }
-\CommentTok{#>   <chr>          <chr>           <chr>   <chr>     <chr>           <chr>}
-\CommentTok{#> 1 S106352031630~ Applied and Co~ 2016-0~ 2019-02-~ 2018-03         10.1~}
-\CommentTok{#> 2 <NA>           Biogeosciences~ 2017-0~ 2017-07-~ <NA>            10.5~}
-\CommentTok{#> 3 <NA>           Global Biogeoc~ 2018-0~ 2019-01-~ 2018-10         10.1~}
-\CommentTok{#> 4 <NA>           IEEE Communica~ 2016-1~ 2017-12-~ 2017            10.1~}
-\CommentTok{#> 5 S002178241400~ Journal de Mat~ 2014-0~ 2018-10-~ 2014-10         10.1~}
-\CommentTok{#> 6 123            Light: Science~ 2019-0~ 2019-01-~ 2019-12         10.1~}
-\CommentTok{#> # ... with 26 more variables: indexed <chr>, issn <chr>, issue <chr>,}
-\CommentTok{#> #   issued <chr>, member <chr>, page <chr>, prefix <chr>, publisher <chr>,}
-\CommentTok{#> #   reference.count <chr>, score <chr>, ...}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{case-study-2-abstract-mining}{%
-\subsection{Case study 2: Abstract
-mining}\label{case-study-2-abstract-mining}}
-
-Sometimes you just need abstracts for your research question. The
-benefit of only needing abstracts, and not need full text, is that
-there's many more articles that will have abstracts available than have
-their full text available.
-
-As an example, let's say you xxxx
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"fulltext"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{xxxxx}
-
-Using \texttt{fulltext}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{res <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"crossref"}\NormalTok{,}
-  \DataTypeTok{crossrefopts =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{filter =} \KeywordTok{c}\NormalTok{(}\DataTypeTok{has_abstract =} \OtherTok{TRUE}\NormalTok{)))}
-\NormalTok{ids <-}\StringTok{ }\NormalTok{res}\OperatorTok{$}\NormalTok{crossref}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{doi}
-\NormalTok{out <-}\StringTok{ }\KeywordTok{ft_abstract}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ ids, }\DataTypeTok{from =} \StringTok{"crossref"}\NormalTok{)}
-\NormalTok{abstracts <-}\StringTok{ }\KeywordTok{vapply}\NormalTok{(out}\OperatorTok{$}\NormalTok{crossref, }\StringTok{"[["}\NormalTok{, }\StringTok{""}\NormalTok{, }\StringTok{"abstract"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Using \href{https://quanteda.io/}{quanteda}, read the abstracts into a
-corpus
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"quanteda"}\NormalTok{)}
-\NormalTok{corp <-}\StringTok{ }\KeywordTok{corpus}\NormalTok{(abstracts)}
-\KeywordTok{docvars}\NormalTok{(corp) <-}\StringTok{ }\NormalTok{ids}
-\end{Highlighting}
-\end{Shaded}
-
-Get a summary of the abstracts
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{summary}\NormalTok{(corp)}
-\CommentTok{#> Corpus consisting of 10 documents:}
-\CommentTok{#> }
-\CommentTok{#>    Text Types Tokens Sentences                    V1}
-\CommentTok{#>   text1   143    262        10   10.2458/v22i1.21112}
-\CommentTok{#>   text2   117    244         6   10.2458/v17i1.21696}
-\CommentTok{#>   text3    75    118         4   10.2458/v25i1.23119}
-\CommentTok{#>   text4     5      8         1    10.2458/v1i1.21154}
-\CommentTok{#>   text5   105    171         7   10.1155/2011/868426}
-\CommentTok{#>   text6   112    181         6   10.1155/2012/273413}
-\CommentTok{#>   text7   117    240         8 10.5194/we-13-91-2013}
-\CommentTok{#>   text8   140    245         9 10.5194/we-13-95-2013}
-\CommentTok{#>   text9   107    202         7   10.1155/2014/198707}
-\CommentTok{#>  text10   118    224         6   10.5402/2011/897578}
-\CommentTok{#> }
-\CommentTok{#> Source: /Users/sckott/github/ropensci/textmine/use-cases/* on x86_64 by sckott}
-\CommentTok{#> Created: Thu Apr 11 11:20:19 2019}
-\CommentTok{#> Notes:}
-\end{Highlighting}
-\end{Shaded}
-
-Use the \texttt{kwic()} function to see a word in context across the
-abstracts
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{kwic}\NormalTok{(corp, }\DataTypeTok{pattern =} \StringTok{"ecology"}\NormalTok{)}
-\CommentTok{#>                                                                         }
-\CommentTok{#>   [text1, 33] knowledge production within critical political | ecology |}
-\CommentTok{#>   [text1, 50]              in scientific articles on dryland | ecology |}
-\CommentTok{#>  [text1, 204]                 to equilibrium models in range | ecology |}
-\CommentTok{#>  [text1, 246]    communal areas.Keywords: Critical political | ecology |}
-\CommentTok{#>  [text1, 255]                 , scientific models, rangeland | ecology |}
-\CommentTok{#>    [text2, 5]                            < jats:p> Political | ecology |}
-\CommentTok{#>   [text2, 23]        manifestations of political economy and | ecology |}
-\CommentTok{#>   [text2, 45]                      I try to extend political | ecology |}
-\CommentTok{#>  [text2, 149]                   , in dialogue with political | ecology |}
-\CommentTok{#>  [text2, 177]            people and resources that political | ecology |}
-\CommentTok{#>  [text2, 229]      indigeneity scholars.Key words: political | ecology |}
-\CommentTok{#>   [text3, 71]                   an analysis from a political | ecology |}
-\CommentTok{#>  [text3, 114]                system, supermarkets, political | ecology |}
-\CommentTok{#>  [text6, 134]                was observed when allopatry and | ecology |}
-\CommentTok{#>  [text7, 167]             ecosystem should be considered for | ecology |}
-\CommentTok{#>  [text7, 185]                       the" four-color issue of | ecology |}
-\CommentTok{#>  [text7, 201]             step toward advancing knowledge in | ecology |}
-\CommentTok{#>  [text9, 195]         or for theoretical studies integrating | ecology |}
-\CommentTok{#>                                              }
-\CommentTok{#>  . This article is a                         }
-\CommentTok{#>  , and investigates the functions            }
-\CommentTok{#>  , and the fence-line photographs            }
-\CommentTok{#>  , fence-line photography, scientific        }
-\CommentTok{#>  , Southern Africa</                         }
-\CommentTok{#>  has expanded in multiple new                }
-\CommentTok{#>  in the" problem"                            }
-\CommentTok{#>  to engage with ethnic studies               }
-\CommentTok{#>  approaches to better understand the         }
-\CommentTok{#>  focuses on cannot be adequately             }
-\CommentTok{#>  , coloniality, Maidu,                       }
-\CommentTok{#>  standpoint allows a different interpretation}
-\CommentTok{#>  </ jats:p>                                  }
-\CommentTok{#>  act together, leading to                    }
-\CommentTok{#>  "? Here, I                                  }
-\CommentTok{#>  ", and propose that                         }
-\CommentTok{#>  and conservation biology. In                }
-\CommentTok{#>  and biogeography.</}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{case-study-3-full-text-mining}{%
-\subsection{Case study 3: Full text
-mining}\label{case-study-3-full-text-mining}}
-
-In this example, xxxx
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"fulltext"}\NormalTok{)}
-\CommentTok{# library("crminer")}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Search for articles}
-
-Search for the term \emph{ecology} in PLOS journals.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{(res1 <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \StringTok{'plos'}\NormalTok{))}
-\CommentTok{#> Query:}
-\CommentTok{#>   [ecology] }
-\CommentTok{#> Found:}
-\CommentTok{#>   [PLoS: 47272; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0] }
-\CommentTok{#> Returned:}
-\CommentTok{#>   [PLoS: 10; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0]}
-\end{Highlighting}
-\end{Shaded}
-
-Each publisher/search-engine has a slot with metadata and data
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{res1}\OperatorTok{$}\NormalTok{plos}
-\CommentTok{#> Query: [ecology] }
-\CommentTok{#> Records found, returned: [47272, 10] }
-\CommentTok{#> License: [CC-BY] }
-\CommentTok{#>                              id}
-\CommentTok{#> 1  10.1371/journal.pone.0001248}
-\CommentTok{#> 2  10.1371/journal.pone.0059813}
-\CommentTok{#> 3  10.1371/journal.pone.0080763}
-\CommentTok{#> 4  10.1371/journal.pone.0155019}
-\CommentTok{#> 5  10.1371/journal.pone.0175014}
-\CommentTok{#> 6  10.1371/journal.pone.0150648}
-\CommentTok{#> 7  10.1371/journal.pone.0208370}
-\CommentTok{#> 8  10.1371/journal.pcbi.1003594}
-\CommentTok{#> 9  10.1371/journal.pone.0102437}
-\CommentTok{#> 10 10.1371/journal.pone.0166559}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Get full text}
-
-Using the results from \texttt{ft\_search()} we can grab full text of
-some articles
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{(out <-}\StringTok{ }\KeywordTok{ft_get}\NormalTok{(res1))}
-\CommentTok{#> <fulltext text>}
-\CommentTok{#> [Docs] 10 }
-\CommentTok{#> [Source] ext - /Users/sckott/Library/Caches/R/fulltext }
-\CommentTok{#> [IDs] 10.1371/journal.pone.0001248 10.1371/journal.pone.0059813}
-\CommentTok{#>      10.1371/journal.pone.0080763 10.1371/journal.pone.0155019}
-\CommentTok{#>      10.1371/journal.pone.0175014 10.1371/journal.pone.0150648}
-\CommentTok{#>      10.1371/journal.pone.0208370 10.1371/journal.pcbi.1003594}
-\CommentTok{#>      10.1371/journal.pone.0102437 10.1371/journal.pone.0166559 ...}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Extract text from pdfs}
-
-Ideally for text mining you have access to XML or other text based
-formats. However, sometimes you only have access to PDFs. In this case
-you want to extract text from PDFs. \texttt{fulltext} can help with
-that.
-
-You can extract from any pdf from a file path, like:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{path <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"examples"}\NormalTok{, }\StringTok{"example1.pdf"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"fulltext"}\NormalTok{)}
-\KeywordTok{ft_extract}\NormalTok{(path)}
-\CommentTok{#> <document>/Library/Frameworks/R.framework/Versions/3.5/Resources/library/fulltext/examples/example1.pdf}
-\CommentTok{#>   Title: Suffering and mental health among older people living in nursing homes---a mixed-methods study}
-\CommentTok{#>   Producer: pdfTeX-1.40.10}
-\CommentTok{#>   Creation date: 2015-07-17}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Extract text chunks}
-
-Requires the \href{https://github.com/ropensci/pubchunks}{pubchunks}
-library. Here, we'll search for some PLOS articles, then get their full
-text, then extract various parts of each article with
-\texttt{pub\_chunks()}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pubchunks"}\NormalTok{)}
-\NormalTok{res <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"plos"}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{3}\NormalTok{)}
-\NormalTok{x <-}\StringTok{ }\KeywordTok{ft_get}\NormalTok{(res)}
-\NormalTok{x }\OperatorTok{%>%}\StringTok{ }\KeywordTok{ft_collect}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pub_chunks}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"doi"}\NormalTok{, }\StringTok{"history"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pub_tabularize}\NormalTok{()}
-\CommentTok{#> $plos}
-\CommentTok{#> $plos$`10.1371/journal.pone.0001248`}
-\CommentTok{#>                            doi history.received history.accepted}
-\CommentTok{#> 1 10.1371/journal.pone.0001248       2007-07-02       2007-11-06}
-\CommentTok{#>   .publisher}
-\CommentTok{#> 1       plos}
-\CommentTok{#> }
-\CommentTok{#> $plos$`10.1371/journal.pone.0059813`}
-\CommentTok{#>                            doi history.received history.accepted}
-\CommentTok{#> 1 10.1371/journal.pone.0059813       2012-09-16       2013-02-19}
-\CommentTok{#>   .publisher}
-\CommentTok{#> 1       plos}
-\CommentTok{#> }
-\CommentTok{#> $plos$`10.1371/journal.pone.0080763`}
-\CommentTok{#>                            doi history.received history.accepted}
-\CommentTok{#> 1 10.1371/journal.pone.0080763       2013-08-15       2013-10-16}
-\CommentTok{#>   .publisher}
-\CommentTok{#> 1       plos}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Further processing}
-
-From here, you'll likely want to use R packages for removing stopwords,
-tokenizing, and more. The Natural Language Processing CRAN Task View
-(\url{https://cloud.r-project.org/web/views/NaturalLanguageProcessing.html})
-is a good place to start if you're not sure what packages to use. Some
-packages worth trying in this space are
-\href{https://github.com/ropensci/antiword}{antiword},
-\href{https://github.com/ropensci/pdftools}{pdftools},
-\href{https://github.com/ropensci/rtika}{rtika},
-\href{https://github.com/ropensci/tokenizers}{tokenizers}, and
-\href{https://github.com/ropensci/textreuse}{textreuse}.
-
 \hypertarget{data-sources}{%
@@
 
+\hypertarget{text-mining-in-r-use-cases}{%
+\section{Text mining in R: Use Cases}\label{text-mining-in-r-use-cases}}
+
+\hypertarget{use-cases-in-the-literature}{%
+\subsection{Use cases in the
+literature}\label{use-cases-in-the-literature}}
+
+At least three papers have used the fulltext package. Bauer, Barbera \&
+Munzert (2016) explored a way to think about the qualitative impact of
+citations rather than simple counts of citations; they did not use
+fulltext, but pointed it out as a good option for collecting papers.
+Walsh \& Cherbuin (2018) used fulltext and
+\href{https://github.com/ropensci/roadoi}{roadoi}, an rOpenSci R
+package, to ``map the literature'' on nutritional interventions in
+cognitive health. Piper et al. (2019) in a review of DNA metabarcoding
+for high-throughput insect surveillance used fulltext, along with
+\href{https://github.com/ropensci/rentrez}{rentrez} (an rOpenSci
+package), to collect articles for text mining.
+
+\hypertarget{three-case-studies}{%
+\subsection{Three case studies}\label{three-case-studies}}
+
+The following three case studies demonstrate in brief how one could use
+fulltext and related packages to do a few different text mining tasks.
+
+\hypertarget{case-study-1-citation-mining}{%
+\subsubsection{Case study 1: Citation
+mining}\label{case-study-1-citation-mining}}
+
+In this example, xxxx
+
+\emph{Load libraries}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"rcrossref"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"rplos"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"rorcid"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"rcitoid"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"citecorp"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{rcrossref}
+
+Using \texttt{rcrossref} for Crossref data:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\KeywordTok{cr_works}\NormalTok{(}\DataTypeTok{query=}\StringTok{"NSF"}\NormalTok{)}
+\KeywordTok{head}\NormalTok{(x}\OperatorTok{$}\NormalTok{data)}
+\CommentTok{#> # A tibble: 6 x 32}
+\CommentTok{#>   alternative.id container.title created deposited published.print doi  }
+\CommentTok{#>   <chr>          <chr>           <chr>   <chr>     <chr>           <chr>}
+\CommentTok{#> 1 S106352031630~ Applied and Co~ 2016-0~ 2019-02-~ 2018-03         10.1~}
+\CommentTok{#> 2 <NA>           Biogeosciences~ 2017-0~ 2017-07-~ <NA>            10.5~}
+\CommentTok{#> 3 <NA>           Global Biogeoc~ 2018-0~ 2019-01-~ 2018-10         10.1~}
+\CommentTok{#> 4 <NA>           IEEE Communica~ 2016-1~ 2017-12-~ 2017            10.1~}
+\CommentTok{#> 5 S002178241400~ Journal de Mat~ 2014-0~ 2018-10-~ 2014-10         10.1~}
+\CommentTok{#> 6 123            Light: Science~ 2019-0~ 2019-01-~ 2019-12         10.1~}
+\CommentTok{#> # ... with 26 more variables: indexed <chr>, issn <chr>, issue <chr>,}
+\CommentTok{#> #   issued <chr>, member <chr>, page <chr>, prefix <chr>, publisher <chr>,}
+\CommentTok{#> #   reference.count <chr>, score <chr>, ...}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{case-study-2-abstract-mining}{%
+\subsubsection{Case study 2: Abstract
+mining}\label{case-study-2-abstract-mining}}
+
+Sometimes you just need abstracts for your research question. The
+benefit of only needing abstracts, and not need full text, is that
+there's many more articles that will have abstracts available than have
+their full text available.
+
+As an example, let's say you xxxx
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"fulltext"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{xxxxx}
+
+Using \texttt{fulltext}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"crossref"}\NormalTok{,}
+  \DataTypeTok{crossrefopts =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{filter =} \KeywordTok{c}\NormalTok{(}\DataTypeTok{has_abstract =} \OtherTok{TRUE}\NormalTok{)))}
+\NormalTok{ids <-}\StringTok{ }\NormalTok{res}\OperatorTok{$}\NormalTok{crossref}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{doi}
+\NormalTok{out <-}\StringTok{ }\KeywordTok{ft_abstract}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ ids, }\DataTypeTok{from =} \StringTok{"crossref"}\NormalTok{)}
+\NormalTok{abstracts <-}\StringTok{ }\KeywordTok{vapply}\NormalTok{(out}\OperatorTok{$}\NormalTok{crossref, }\StringTok{"[["}\NormalTok{, }\StringTok{""}\NormalTok{, }\StringTok{"abstract"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Using \href{https://quanteda.io/}{quanteda}, read the abstracts into a
+corpus
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"quanteda"}\NormalTok{)}
+\NormalTok{corp <-}\StringTok{ }\KeywordTok{corpus}\NormalTok{(abstracts)}
+\KeywordTok{docvars}\NormalTok{(corp) <-}\StringTok{ }\NormalTok{ids}
+\end{Highlighting}
+\end{Shaded}
+
+Get a summary of the abstracts
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{summary}\NormalTok{(corp)}
+\CommentTok{#> Corpus consisting of 10 documents:}
+\CommentTok{#> }
+\CommentTok{#>    Text Types Tokens Sentences                    V1}
+\CommentTok{#>   text1   143    262        10   10.2458/v22i1.21112}
+\CommentTok{#>   text2   117    244         6   10.2458/v17i1.21696}
+\CommentTok{#>   text3    75    118         4   10.2458/v25i1.23119}
+\CommentTok{#>   text4     5      8         1    10.2458/v1i1.21154}
+\CommentTok{#>   text5   105    171         7   10.1155/2011/868426}
+\CommentTok{#>   text6   112    181         6   10.1155/2012/273413}
+\CommentTok{#>   text7   117    240         8 10.5194/we-13-91-2013}
+\CommentTok{#>   text8   140    245         9 10.5194/we-13-95-2013}
+\CommentTok{#>   text9   107    202         7   10.1155/2014/198707}
+\CommentTok{#>  text10   118    224         6   10.5402/2011/897578}
+\CommentTok{#> }
+\CommentTok{#> Source: /Users/sckott/github/ropensci/textmine/use-cases/* on x86_64 by sckott}
+\CommentTok{#> Created: Thu Apr 11 11:20:19 2019}
+\CommentTok{#> Notes:}
+\end{Highlighting}
+\end{Shaded}
+
+Use the \texttt{kwic()} function to see a word in context across the
+abstracts
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{kwic}\NormalTok{(corp, }\DataTypeTok{pattern =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{window =} \DecValTok{2}\NormalTok{)}
+\CommentTok{#>                                                                }
+\CommentTok{#>   [text1, 33]  critical political | ecology | . This           }
+\CommentTok{#>   [text1, 50]          on dryland | ecology | , and            }
+\CommentTok{#>  [text1, 204]            in range | ecology | , and            }
+\CommentTok{#>  [text1, 246]  Critical political | ecology | , fence-line     }
+\CommentTok{#>  [text1, 255]         , rangeland | ecology | , Southern       }
+\CommentTok{#>    [text2, 5]         > Political | ecology | has expanded     }
+\CommentTok{#>   [text2, 23]         economy and | ecology | in the           }
+\CommentTok{#>   [text2, 45]    extend political | ecology | to engage        }
+\CommentTok{#>  [text2, 149]      with political | ecology | approaches to    }
+\CommentTok{#>  [text2, 177]      that political | ecology | focuses on       }
+\CommentTok{#>  [text2, 229]         : political | ecology | , coloniality    }
+\CommentTok{#>   [text3, 71]         a political | ecology | standpoint allows}
+\CommentTok{#>  [text3, 114]         , political | ecology | </               }
+\CommentTok{#>  [text6, 134]       allopatry and | ecology | act together     }
+\CommentTok{#>  [text7, 167]      considered for | ecology | "?               }
+\CommentTok{#>  [text7, 185]            issue of | ecology | ",               }
+\CommentTok{#>  [text7, 201]        knowledge in | ecology | and conservation }
+\CommentTok{#>  [text9, 195] studies integrating | ecology | and biogeography}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{case-study-3-full-text-mining}{%
+\subsubsection{Case study 3: Full text
+mining}\label{case-study-3-full-text-mining}}
+
+In this example, xxxx
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"fulltext"}\NormalTok{)}
+\CommentTok{# library("crminer")}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Search for articles}
+
+Search for the term \emph{ecology} in PLOS journals.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{(res1 <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \StringTok{'plos'}\NormalTok{))}
+\CommentTok{#> Query:}
+\CommentTok{#>   [ecology] }
+\CommentTok{#> Found:}
+\CommentTok{#>   [PLoS: 47272; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0] }
+\CommentTok{#> Returned:}
+\CommentTok{#>   [PLoS: 10; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0]}
+\end{Highlighting}
+\end{Shaded}
+
+Each publisher/search-engine has a slot with metadata and data
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{res1}\OperatorTok{$}\NormalTok{plos}
+\CommentTok{#> Query: [ecology] }
+\CommentTok{#> Records found, returned: [47272, 10] }
+\CommentTok{#> License: [CC-BY] }
+\CommentTok{#>                              id}
+\CommentTok{#> 1  10.1371/journal.pone.0001248}
+\CommentTok{#> 2  10.1371/journal.pone.0059813}
+\CommentTok{#> 3  10.1371/journal.pone.0080763}
+\CommentTok{#> 4  10.1371/journal.pone.0155019}
+\CommentTok{#> 5  10.1371/journal.pone.0175014}
+\CommentTok{#> 6  10.1371/journal.pone.0150648}
+\CommentTok{#> 7  10.1371/journal.pone.0208370}
+\CommentTok{#> 8  10.1371/journal.pcbi.1003594}
+\CommentTok{#> 9  10.1371/journal.pone.0102437}
+\CommentTok{#> 10 10.1371/journal.pone.0166559}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Get full text}
+
+Using the results from \texttt{ft\_search()} we can grab full text of
+some articles
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{(out <-}\StringTok{ }\KeywordTok{ft_get}\NormalTok{(res1))}
+\CommentTok{#> <fulltext text>}
+\CommentTok{#> [Docs] 10 }
+\CommentTok{#> [Source] ext - /Users/sckott/Library/Caches/R/fulltext }
+\CommentTok{#> [IDs] 10.1371/journal.pone.0001248 10.1371/journal.pone.0059813}
+\CommentTok{#>      10.1371/journal.pone.0080763 10.1371/journal.pone.0155019}
+\CommentTok{#>      10.1371/journal.pone.0175014 10.1371/journal.pone.0150648}
+\CommentTok{#>      10.1371/journal.pone.0208370 10.1371/journal.pcbi.1003594}
+\CommentTok{#>      10.1371/journal.pone.0102437 10.1371/journal.pone.0166559 ...}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Extract text from pdfs}
+
+Ideally for text mining you have access to XML or other text based
+formats. However, sometimes you only have access to PDFs. In this case
+you want to extract text from PDFs. \texttt{fulltext} can help with
+that.
+
+You can extract from any pdf from a file path, like:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{path <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"examples"}\NormalTok{, }\StringTok{"example1.pdf"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"fulltext"}\NormalTok{)}
+\KeywordTok{ft_extract}\NormalTok{(path)}
+\CommentTok{#> <document>/Library/Frameworks/R.framework/Versions/3.5/Resources/library/fulltext/examples/example1.pdf}
+\CommentTok{#>   Title: Suffering and mental health among older people living in nursing homes---a mixed-methods study}
+\CommentTok{#>   Producer: pdfTeX-1.40.10}
+\CommentTok{#>   Creation date: 2015-07-17}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Extract text chunks}
+
+Requires the \href{https://github.com/ropensci/pubchunks}{pubchunks}
+library. Here, we'll search for some PLOS articles, then get their full
+text, then extract various parts of each article with
+\texttt{pub\_chunks()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(}\StringTok{"pubchunks"}\NormalTok{)}
+\NormalTok{res <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"plos"}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{3}\NormalTok{)}
+\NormalTok{x <-}\StringTok{ }\KeywordTok{ft_get}\NormalTok{(res)}
+\NormalTok{x }\OperatorTok{%>%}\StringTok{ }\KeywordTok{ft_collect}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pub_chunks}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"doi"}\NormalTok{, }\StringTok{"history"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pub_tabularize}\NormalTok{()}
+\CommentTok{#> $plos}
+\CommentTok{#> $plos$`10.1371/journal.pone.0001248`}
+\CommentTok{#>                            doi history.received history.accepted}
+\CommentTok{#> 1 10.1371/journal.pone.0001248       2007-07-02       2007-11-06}
+\CommentTok{#>   .publisher}
+\CommentTok{#> 1       plos}
+\CommentTok{#> }
+\CommentTok{#> $plos$`10.1371/journal.pone.0059813`}
+\CommentTok{#>                            doi history.received history.accepted}
+\CommentTok{#> 1 10.1371/journal.pone.0059813       2012-09-16       2013-02-19}
+\CommentTok{#>   .publisher}
+\CommentTok{#> 1       plos}
+\CommentTok{#> }
+\CommentTok{#> $plos$`10.1371/journal.pone.0080763`}
+\CommentTok{#>                            doi history.received history.accepted}
+\CommentTok{#> 1 10.1371/journal.pone.0080763       2013-08-15       2013-10-16}
+\CommentTok{#>   .publisher}
+\CommentTok{#> 1       plos}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Further processing}
+
+From here, you'll likely want to use R packages for removing stopwords,
+tokenizing, and more. The Natural Language Processing CRAN Task View
+(\url{https://cloud.r-project.org/web/views/NaturalLanguageProcessing.html})
+is a good place to start if you're not sure what packages to use. Some
+packages worth trying in this space are
+\href{https://github.com/ropensci/antiword}{antiword},
+\href{https://github.com/ropensci/pdftools}{pdftools},
+\href{https://github.com/ropensci/rtika}{rtika},
+\href{https://github.com/ropensci/tokenizers}{tokenizers}, and
+\href{https://github.com/ropensci/textreuse}{textreuse}.
+
 \hypertarget{future-directions}{%
@@
 
+\leavevmode\hypertarget{ref-Bauer2016}{}%
+Bauer PC., Barbera P., Munzert S. 2016. The quality of citations:
+Towards quantifying qualitative impact in social science research.
+\emph{SSRN Electronic Journal}.
+
 \leavevmode\hypertarget{ref-Camerer2016}{}%
@@
 
+\leavevmode\hypertarget{ref-Piper2019}{}%
+Piper AM., Batovska J., Cogan NOI., Weiss J., Cunningham JP., Rodoni
+BC., Blacket MJ. 2019. Prospects and challenges of implementing DNA
+metabarcoding for high-throughput insect surveillance.
+\emph{GigaScience} 8.
+
 \leavevmode\hypertarget{ref-Piwowar2018}{}%
@@
 
+\leavevmode\hypertarget{ref-Walsh2018}{}%
+Walsh E., Cherbuin N. 2018. Mapping the literature on nutritional
+interventions in cognitive health: A data-driven approach.
+\emph{Nutrients} 11:38.
+
 \end{document}

### 2019-08-23 manuscript.Rmd

@@
 "versioned, publicly-editable catalog of research publications:  journal articles,
-conference proceedings, preprints, blog posts". Fatcat is currently does not have an
+conference proceedings, pre-prints, blog posts". Fatcat currently does not have an
 R client, but is used inside of the [fulltext][] package.

### 2019-08-23 manuscript.tex

@@
 
-There's more than 100 million research articles published (Crossref API:
-\url{https://github.com/CrossRef/rest-api-doc}), representing an
+There are more than 100 million research articles published (Crossref
+API: \url{https://github.com/CrossRef/rest-api-doc}), representing an
 enormous amount of knowledge. In addition to simply reading these

### 2019-08-23 use-cases/full-text-mining.Rmd

@@
 ```
+
+_Further processing_
+
+From here, you'll likely want to use R packages for removing stopwords, tokenizing, 
+and more. The Natural Language Processing CRAN Task View
+(https://cloud.r-project.org/web/views/NaturalLanguageProcessing.html) is a good place
+to start if you're not sure what packages to use. Some packages worth trying in this
+space are [antiword][], [pdftools][], [rtika][], [tokenizers][], and [textreuse][].
+
+[antiword]: https://github.com/ropensci/antiword
+[pdftools]: https://github.com/ropensci/pdftools
+[rtika]: https://github.com/ropensci/rtika
+[tokenizers]: https://github.com/ropensci/tokenizers
+[textreuse]: https://github.com/ropensci/textreuse

### 2019-08-23 manuscript.tex
remake pdf

@@
 a ``versioned, publicly-editable catalog of research publications:
-journal articles, conference proceedings, preprints, blog posts''.
-Fatcat is currently does not have an R client, but is used inside of the
+journal articles, conference proceedings, pre-prints, blog posts''.
+Fatcat currently does not have an R client, but is used inside of the
 \href{https://github.com/ropensci/fulltext}{fulltext} package.
